---
title: "Fermentation identification | v6"
output: html_notebook
---

This notebook requires the dataset produced in v6' `01 load.Rmd`. It:

-   isolates fermentation v non-fermentation periods
-   adds pulse cooling
-   creates and augments the pulse dataset

```{r libs, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(pracma)
library(RcppRoll)
detach("package:dplyr")
library(dplyr)

# Project path
path_root <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
path_data <- "/data/v6"
path_tank <- "/E1"
```

# Load required data

```{r}
(df <- readr::read_csv(
  paste0(path_root, path_data, path_tank, '/clean/df_clean.csv'), 
  col_types = cols(.default = "?", state = "i", abv = "d", mode_type = "i")))
```

```{r}

df %>% summary

```

# Identify and mark fermentation periods

Identify fermentation periods by eye

```{r eval=FALSE}

(p1 <- ggplot(data = df, aes(x = time)) +
  geom_line(aes(y = brix, colour = "brix"), colour = "steelblue"))

(p2 <- ggplot(data = df, aes(x = time)) +
  geom_line(aes(y = status, colour = "status"), colour = "tomato"))

gridExtra::grid.arrange(p1, p2, nrow = 2)

plotly::ggplotly(p1)
plotly::ggplotly(p2)

```

## Automatic fermentation curve detection

Looking at above historic brix data, the first step should be to detect the overall fermentation periods fenced in by

-   a starting point characterised by a steep rise in brix from a very low level to \~30brix
-   an end point characterised by a steep fall in brix into low levels

### Detect fermentation periods

Step 1) Focus on brix changes

Step 2) Identify fermentation periods

-   Identify fermentation start periods by simply marking all brix values above 15 brix (brix only ever goes beyond \~ +15 when a fermentation starts)
-   Next identify all potential end periods by marking all brix values below -100 brix. There will be many and only a few will be real end points
-   We identify the real end points in the next step

Step 3) Fill the NA's between ferment start periods and end periods

-   This step will close the gaps only between real start and real end points of a fermentation

Step 4) Remove the false end points

-   Now we can remove all false end points as they are not preceded by a "ferment" value

Step 5) Label the first fermentation value "start"

Step 6) Label each fermentation period by its fermentation start date

Step 7) Add a ferment boolean

```{r}

# Identify fermentation periods
df_brix <- df %>%
  select(time, brix, temp) %>%
  # filter(time > lubridate::ymd("2021-09-14"), time < "2021-10-20") %>%
  
  # 1) keep only brix changes.
  filter(brix != lag(brix)) %>% 

  # 2) Find fermentation start and end.
  mutate(ferment = case_when(
    brix > 15 ~ "ferment",
    lead(brix) < -100 ~ "end",
    TRUE ~ NA_character_
  )) %>% 

  # 3) Fill time points during start and end.
  fill(ferment) %>%

  # 4) Remove redundant "end" values.
  mutate(keep = if_else(ferment == "ferment" | (ferment == "end" & lag(ferment) == "ferment"), TRUE, FALSE)) %>%
  mutate(ferment = if_else(keep, ferment, NA_character_)) %>%
  select(-keep) %>%

  # 5) Rename first fermentation value to "start".
  mutate(ferment = if_else(ferment == "ferment" & is.na(lag(ferment)), "start", ferment)) %>%
  
  # 6) Name the fermentation period with their tank and start date.
  mutate(ferment_period = case_when(
    ferment == "start" ~ paste0(stringr::str_replace(path_tank, "/", ""), "-", format(time, "%y-%b-%d")),
    ferment == "end" ~ "end",
    TRUE ~ NA_character_))  %>%
  fill(ferment_period) %>%
  mutate(ferment_period = if_else(ferment_period == "end", lag(ferment_period), ferment_period)) %>%
  mutate(ferment_period = if_else(ferment_period == "end", NA_character_, ferment_period)) %>%

  # 7) Add a boolean ferment variable
  mutate(ferment_bool = if_else(!is.na(ferment), TRUE, FALSE))

# View(df_brix)
  
```

Check visually

```{r}

# Just checking ferment phase start and end.
starts <- df_brix %>% 
  mutate(check = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(check)

ends <- df_brix %>% 
  mutate(check = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(check)

(p <- df_brix %>% 
  ggplot() +
    geom_line(aes(x = time, y = brix)) + 
    geom_vline(aes(xintercept = starts), colour = 'cornflowerblue') +
    geom_vline(aes(xintercept = ends), colour = 'maroon') 
)

plotly::ggplotly(p)

```

### Defences

Check and remove faux fermentation periods

(check for defences not to wipe out sensible fermentation periods as was the case with tank F1)

#### Incorrect falloff

Above algo detected a false positive fermentation period for tank C1 that lingered constantly at around 20 brix before falling off.

The following procedure defends against that case by:

1.  Creating a rolling average with a window of 10
2.  Checking if rolling brix before drop off was above 10
3.  It should be around 0 or below, so if it's high it's defined as a false positive and the whole period will be removed

```{r}

# Detect and isolate false positive fermentation periods.
false_periods <- df_brix %>% 
  mutate(brix_rolling = RcppRoll::roll_mean(brix, 10, align = "right", fill = NA_real_)) %>% 
  mutate(remove_period = ifelse(ferment == "end" & lag(brix_rolling) > 10, TRUE, FALSE)) %>% 
  filter(remove_period) %>% 
  distinct(ferment_period) %>% 
  pull(ferment_period)

# Remove them.
df_brix <- df_brix %>% 
  mutate(
    ferment_bool = ifelse(ferment_period %in% false_periods, FALSE, ferment_bool),
    ferment_period = ifelse(ferment_bool, ferment_period, NA_character_),
    ferment = ifelse(ferment_bool, ferment, NA_character_),
  )

```

#### Too short

```{r}

# Single out all ferment_periods that are shorter than 3 days
short_periods <- df_brix %>%
  filter(ferment == "start" | ferment == "end") %>% 
  group_by(ferment_period, ferment, time) %>% 
  count() %>% 
  ungroup() %>%
  mutate(period_in_days = difftime(lag(time), time, units = "days") %>% as.double) %>% 
  filter(ferment == "start") %>% 
  filter(period_in_days > 0, period_in_days < 3) %>% 
  pull(ferment_period)

# Remove them.
df_brix <- df_brix %>%
  mutate(
    ferment_bool = ifelse(ferment_period %in% short_periods, FALSE, ferment_bool),
    ferment_period = ifelse(ferment_bool, ferment_period, NA_character_),
    ferment = ifelse(ferment_bool, ferment, NA_character_),
  )
  
```

### Single start or end points

```{r}

orphans <- df_brix %>%
  filter(ferment == "start" | ferment == "end") %>% 
  group_by(ferment_period) %>% 
  count() %>% 
  filter(n == 1) %>% 
  pull(ferment_period)
  
# Remove them.
df_brix <- df_brix %>%
  mutate(
    ferment_bool = ifelse(ferment_period %in% orphans, FALSE, ferment_bool),
    ferment_period = ifelse(ferment_bool, ferment_period, NA_character_),
    ferment = ifelse(ferment_bool, ferment, NA_character_),
  )


```

### Visual ferment period check:

```{r eval=FALSE}

# Check ferment start and end are detected correctly.
starts <- df_brix %>% 
  mutate(ferment = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(ferment)

ends <- df_brix %>% 
  mutate(ferment = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(ferment)

p <- df_brix %>% 
  ggplot() +
    geom_line(aes(x = time, y = brix)) + 
    geom_vline(aes(xintercept = starts), colour = 'cornflowerblue') +
    geom_vline(aes(xintercept = ends), colour = 'maroon')

plotly::ggplotly(p)

```

### Detect fermentation phases

#### Intro

Each fermentation curve is assumed to go through 3 phases characterised by four points:

-   Phase 1: pre-fermentation (flat curve at initial brix value of \~20/30)
-   Phase 2: fermentation (falling curve)
-   Phase 3: post-fermentation (flat curve ataround just sub-zero brix)

These are defined by four points and one extra points:

-   Point 1 (Phase 1 start): brix up to fermentation start levels (around 20/30 brix)
-   Point 2 (Phase 2 start): brix decreases notable
-   Point 3 (Phase 2 still): brix inflection - cell division decreases
-   Point 4 (Phase 3 start): brix mainly remains below 0 but still fluctuates
-   Point 5 (Phase 3 end): brix remains below 0 and steady and/or drops off

We identified Points (1) and (4) above. Points (2) and (3) are hard to find in the raw data, so we should try with a LOESS smoothed curve.

#### Detect pre-fermentation start (point 1) and post-fermentation end (point 5)

Straight forward as points and phase startes coincide with start and end:

```{r}

# Add start and end phases. 
df_brix <- df_brix %>% 
  mutate(ferment_phase = case_when(
    ferment == "start" ~ "Phase 1",
    ferment == "end" ~ "Phase 3",
    TRUE ~ NA_character_,
    )
  )

# Add points 1 and 4
df_brix <- df_brix %>% 
  mutate(ferment_point = case_when(
    ferment == "start" ~ "Point 1",
    ferment == "end" ~ "Point 5",
    TRUE ~ NA_character_,
    )
  )

```

#### Detect fermentation start (point 2) \| interactive

Let's try with the base LOESS function first, if that ain't working let's throw STL at it (time series loess):

Base LOESS doesn't work on the full data, but on particular fermentation periods, they do. Let's try with them:

```{r eval=FALSE}

# Focus in on just fermentation periods
df_brix %>% 
  group_by(ferment_period) %>% 
  count()

```

Now let's see what works for point (2):

```{r eval=FALSE}

# Create base data (based on specific periods)
df_test <- df_brix %>% 
  rowid_to_column("index") %>% 
  filter(ferment_period == "21-Nov-01")
  # filter(ferment_period == "21-Oct-02")
  # filter(ferment_period == "21-Oct-19")
  # filter(ferment_period == "21-Sep-14")
  # filter(ferment_period == "22-Aug-21")
  # filter(ferment_period == "22-Sep-09")
  # filter(ferment_period == "22-Sep-21")
  # filter(ferment_period == "22-Sep-29")
  
# Add LOESS smoother.
span <- 0.25
result <- loess(brix ~ index, data = df_test, span = span)
df_test$brix_smooth <- predict(result)

# Add gradient.
df_test <- df_test %>% 
  mutate(brix_deriv_1 = pracma::gradient(df_test$brix_smooth))

# Chart.
p1 <- df_test %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix)) +
    geom_line(aes(y = brix_smooth), colour = "tomato") 

p2 <- df_test %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1), colour = "blue")

gridExtra::grid.arrange(p1, p2, nrow = 2)

plotly::ggplotly(p1)
plotly::ggplotly(p2)

```

So after trying every above combination I think we should stick to simple and take the first row with brix_deriv_1 \< 0 as point(2)

Sometimes that's a bit too early, sometimes it's too late. We can conceive a more complex algo (e.g. taking into account raw brix values or gradients) but let's just keep it simple for now :)

#### Detect fermentation start and end points (points 2, 4) \| automatic

For each fermentation period calculate the LOESS smoothed brix curve, the smooth brix curve's slope (1st derivative) and the smoth brix curve's acceleration (2nd derivative).

We'll use the slope and acceleration to detect points 2 (curve start), just the slope to detect point 3 (curve inflection) and 4 (curve end).

We do so by using tidyr's nesting on`df_brix` which creates list-columns for each fermentation_period in order for loess and gradient to run on each nested period data.

Note, we're using r's base loess function with a span we set to **0.15**. 0.25 worked well on D1 but botched a fermentation on C1 which needed a tighter curve.

We're NOT using STL decomposition with a loess trend as base loess is just much simpler :))

```{r}

# Add smooth brix and gradient.
df_brix_smooth <- df_brix %>% 
  # Row number column required for loess calc
  rowid_to_column("index") %>% 

  # Get brix_smooth and brix_deriv_1
  # ---------------------------------

  # Nest by ferment_period (funny xyntax, but hey)
  tidyr::nest(data = c(-ferment_period, index, time, brix)) %>% 

  # for each ferment_period calculate the loess for their fitted values
  # and the gradient for the resulting brix_smooth values.
  mutate(
    m = purrr::map(data, loess, formula = brix ~ index, span = 0.15),
    brix_smooth = purrr::map(m, `[[`, "fitted"),
    brix_deriv_1 = purrr::map(brix_smooth, pracma::gradient)
  ) %>% 

  # remove the model variable
  select(-m) %>%

  # Unnest the data :)
  tidyr::unnest(cols = c(data, brix_smooth, brix_deriv_1)) %>% 

  # Get brix_deriv_1_smooth
  # ------------------------

  # Nest by ferment_period and get the gradient in
  tidyr::nest(data = c(-ferment_period, index, time, brix_smooth, brix_deriv_1)) %>% 

  # for each ferment_period calculate also smooth ouy the brix_deriv_1.
  mutate(
    m = purrr::map(data, loess, formula = brix_deriv_1 ~ index, span = 0.25),
    brix_deriv_1_smooth = purrr::map(m, `[[`, "fitted"),
    brix_deriv_2 = purrr::map(brix_deriv_1_smooth, pracma::gradient)
  ) %>% 

  # remove the model variable
  select(-m) %>%

  # Unnest the data :)
  tidyr::unnest(cols = c(data, brix_deriv_1_smooth, brix_deriv_2)) %>% 

  # Clean up
  select(-index) %>% 
  select(time, brix, brix_smooth, brix_deriv_1, brix_deriv_1_smooth, brix_deriv_2, ferment, ferment_bool, ferment_period, ferment_point, ferment_phase, temp) %>% 
  arrange(time)

```

Check visually:

```{r eval=FALSE}

# Check brix_smooth, brix_deriv_1 and brix_deriv_2 make sense.
df_brix_smooth %>% 
  filter(ferment_bool) %>% 
  # filter(ferment_period == '22-Sep-12') %>% # CHANGED
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix)) +
    geom_line(aes(y = brix_smooth), color = "tomato") +
    facet_wrap(~ferment_period, scales = "free", nrow = 2)

df_brix_smooth %>%
  filter(ferment_bool) %>%
  # filter(ferment_period == '22-Sep-12') %>% # CHANGED 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'steelblue') +
    facet_wrap(~ferment_period, scales = "free", nrow = 2)

df_brix_smooth %>%
  filter(ferment_bool) %>%
  # filter(ferment_period == '22-Sep-12') %>% # CHANGED 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru') +
    facet_wrap(~ferment_period, scales = "free", nrow = 2)


```

Add points 2 and 4 next (the start of phase 2 and the start of phase 3). Note, that there can be fermentation curves that skip phase 1 altogether and start with phase 2 straight away. Simple reason being, the smooth brix curve gradient is negative from the start.

Point 2 conditons - curve start, brix falls off:

1.  The brix curve needs to be falling (1st derivative / slope \< 0)
2.  The brix curve's fall needs to be accelerating (2nd derivative / acceleration \< 0)

Point 4 condition - curve end: brix must be below 0

```{r}

# Get the times for point 2: brix starts falling
ferment_curve_start_times <- df_brix_smooth %>% 
  filter(ferment_bool, brix_deriv_1 < 0, brix_deriv_2 < 0) %>% 
  group_by(ferment_period) %>% 
  mutate(first = first(brix_deriv_1)) %>% 
  ungroup() %>% 
  select(time, first) %>% 
  distinct(first, .keep_all = TRUE) %>% 
  pull(time)

# Get the times for point 4: brix below 0
ferment_curve_end_times <- df_brix_smooth %>% 
  filter(ferment_bool, brix_smooth < 0) %>% 
  group_by(ferment_period) %>% 
  mutate(first = first(brix_smooth)) %>% 
  ungroup() %>% 
  select(time, first) %>% 
  distinct(first, .keep_all = TRUE) %>% 
  pull(time)

# Add them to the `ferment_phase` var.
df_brix_smooth <- df_brix_smooth %>% 
  mutate(ferment_point = if_else(time %in% ferment_curve_start_times, "Point 2", ferment_point)) %>% 
  mutate(ferment_phase = if_else(time %in% ferment_curve_start_times, "Phase 2", ferment_phase)) %>% 
  mutate(ferment_point = if_else(time %in% ferment_curve_end_times, "Point 4", ferment_point)) %>% 
  mutate(ferment_phase = if_else(time %in% ferment_curve_end_times, "Phase 3", ferment_phase))

```

#### Detect cell division inflection point

Conditions:

1.  for each fermentation period
2.  find the minimum slope of a decreasing brix curve (slope \< 0)
3.  for a minimum brix value of 5

Note, condition (3) was added with tank E1, which had sharp slope offs just before the fermentation end point. This led to the inflection point being close or equal to point 4 (end). The minimum value of 5 was chosen arbitrarily founded on other curves.

Note: this requires all history and needs to be re-written for real time detections!

```{r}

# Get the times for point (2)
inflection_times <- df_brix_smooth %>% 
  fill(ferment_phase) %>%
  filter(ferment_phase == 'Phase 2') %>% 
  filter(ferment_bool, brix_deriv_1_smooth < 0, brix_smooth > 5) %>%
  group_by(ferment_period) %>%
  slice_min(brix_deriv_1_smooth) %>% 
  pull(time)

# Add them to the `ferment_phase` var.
df_brix_smooth <- df_brix_smooth %>%
  mutate(ferment_point = if_else(time %in% inflection_times, "Point 3", ferment_point))
  # Note, that for now, inflection point is still part of phase 2

```

#### Check points and phases

Check for all points over brix visually (note, that non-fermentation brix can be wildly off as it's potentially loess'ed across a loong period:

```{r}

# Check ferment start (point 1 - brix starts falling), end (point  - brix below 05) and the newly created points 2 and 4 are detected correctly.
starts <- df_brix_smooth %>% 
  mutate(check = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(check)

ends <- df_brix_smooth %>% 
  mutate(check = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(check)

point_2s <- df_brix_smooth %>% 
  mutate(check = ifelse(ferment_point == "Point 2", time, NA_Date_)) %>% 
  pull(check)

point_3s <- df_brix_smooth %>% 
  mutate(check = ifelse(ferment_point == "Point 3", time, NA_Date_)) %>% 
  pull(check)

point_4s <- df_brix_smooth %>% 
  mutate(check = ifelse(ferment_point == "Point 4", time, NA_Date_)) %>% 
  pull(check)

(p <- df_brix_smooth %>% 
  ggplot() +
    geom_line(aes(x = time, y = brix)) + 
    geom_line(aes(x = time, y = brix_smooth), colour = 'tomato') + 
    geom_vline(aes(xintercept = starts), colour = 'cornflowerblue') +
    geom_vline(aes(xintercept = ends), colour = 'maroon') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon')
)

plotly::ggplotly(p)

```

Check visually by fermentation period:

```{r}

# Check brix_smooth and brix_deriv_1 make sense.
p1 <- df_brix_smooth %>%
  # filter(!is.na(ferment_period)) %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix)) +
    geom_line(aes(y = brix_smooth), colour = 'tomato') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")

p2 <- df_brix_smooth %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")
  
p3 <- df_brix_smooth %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")
  
plotly::ggplotly(p1)
plotly::ggplotly(p2)
plotly::ggplotly(p3)

```

```{r}
df_brix_smooth %>% 
  group_by(ferment_period) %>% 
  count
```

#### Fill the fermentation phases

```{r}

# Fill phases
df_brix_smooth <- df_brix_smooth %>% 
  fill(ferment_phase) %>% mutate(ferment_phase = ifelse(ferment_bool, ferment_phase, NA_real_))

```

Finally create a period stamped fermentation phase variable...

```{r}

df_brix_smooth <- df_brix_smooth %>%
  mutate(
    ferment_period_phase = ifelse(
      !is.na(ferment_period) & !is.na(ferment_phase), 
      paste(ferment_period, ferment_phase), 
      NA_character_)
    )

```

# Write brix focussed data

This is the result of above wrangle and focuses on brix and the extra fermentation variables.

```{r}

df_brix_smooth %>% write_csv(paste0(path_root, path_data, path_tank, "/clean/df_brix_smooth.csv"), na = "")

```

# Augment original data

Next we augment the original dataset with the extra fermentation variables which requires us to jump through some filling loops - particularly complex for ferment :))

```{r}

df %>% 
  # Join all but the temp (which is fully available in the rhs)
  left_join(df_brix_smooth %>% select(-temp), by = c("time", "brix")) %>% 

  # Fill vars which are straight fillable
  fill(brix_smooth, brix_deriv_1, brix_deriv_1_smooth, ferment_bool) %>% 
  
  # Fill `ferment` which needs drastic filling gymnastics
  mutate(ferment2 = ferment) %>% 
  fill(ferment2) %>% 
  mutate(ferment3 = ifelse(ferment2 == "ferment", ferment2, ferment)) %>% 
  mutate(ferment3 = case_when(
    ferment_bool & ferment == "start" ~ "start",
    ferment_bool & ferment2 == "end" ~ "end",
    TRUE ~ "ferment"
  )) %>% 
  mutate(ferment = ifelse(ferment_bool, ferment3, NA_character_)) %>% 
  select(-c(ferment2, ferment3)) %>% 

  # Fill ferment_period and ferment_phase which need moderate filling gymnastics
  fill(ferment_period, ferment_phase, ferment_period_phase) %>% 
  mutate(ferment_period = ifelse(ferment_bool, ferment_period, NA_character_)) %>%
  mutate(ferment_phase = ifelse(ferment_bool, ferment_phase, NA_character_)) %>%
  mutate(ferment_period_phase = ifelse(ferment_bool, ferment_period_phase, NA_character_)) %>%
  
  # Fill ferment_point which needs a moderate+ amount of filling acrobatics
  mutate(ferment_point2 = ferment_point) %>% 
  fill(ferment_point2) %>% 
  mutate(ferment_point = ifelse(ferment_bool & ferment_point2 == "Point 5", "Point 5", ferment_point)) %>% 
  select(-ferment_point2) %>% 

  # Write
  write_csv(paste0(path_root, path_data, path_tank, "/clean/df_clean_ferment.csv"), na = "")

```

# Charts for the doc / presentation

We're taking an example fermentation and build a few charts

```{r eval=FALSE}

# Data.
df_sample <- df_brix_smooth %>%
  # filter(ferment_period=="D1-22-Sep-29")
  filter(ferment_period=="D1-21-Oct-02")

# Get waypoint vectors.
starts <- df_sample %>% 
  mutate(check = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(check)

ends <- df_sample %>% 
  mutate(check = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(check)

point_2s <- df_sample %>% 
  mutate(check = ifelse(ferment_point == "Point 2", time, NA_Date_)) %>% 
  pull(check)

point_3s <- df_sample %>% 
  mutate(check = ifelse(ferment_point == "Point 3", time, NA_Date_)) %>% 
  pull(check)

point_4s <- df_sample %>% 
  mutate(check = ifelse(ferment_point == "Point 4", time, NA_Date_)) %>% 
  pull(check)

# Graphing function (adding waypoints and facet).
add_points <- function(p) {
  p <- p +
    geom_vline(aes(xintercept = starts), colour = 'darkgreen') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    geom_vline(aes(xintercept = ends), colour = 'darkred') +
    facet_wrap(~ferment_period, scales = "free")
    
  print(p)
}

# Check brix_smooth and brix_deriv_1 make sense.
(p1_base <- df_sample %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix), alpha = 0.4, size = 0.25) +
    geom_line(aes(y = brix_smooth), colour = 'black', size = 0.75))

(p2_base <- df_sample %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'skyblue', size = 0.75))

(p3_base <- df_sample %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru', size = 0.75))

p1 <- df_sample %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix), alpha = 0.4, size = 0.25) +
    geom_line(aes(y = brix_smooth), colour = 'black', size = 0.75)

p1 <- add_points(p1)

p2 <- df_sample %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'skyblue', size = 0.75)

p2 <- add_points(p2)

p3 <- df_sample %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru', size = 0.75) 

p3 <- add_points(p3)
  
plotly::ggplotly(p1)
plotly::ggplotly(p2)
plotly::ggplotly(p3)



```
