---
title: "Fermentation identification | v6"
output: html_notebook
---

This notebook requires the dataset produced in  v6' `01 load.Rmd`. It:

- isolates fermentation v non-fermentation periods
- adds pulse cooling
- creates and augments the pulse dataset

```{r libs, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(pracma)
library(RcppRoll)
detach("package:dplyr")
library(dplyr)

# Project path
path_root <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
path_data <- "/data/v6"
path_tank <- "/C1"
```

# Load required data

```{r}
(df <- readr::read_csv(
  paste0(path_root, path_data, path_tank, '/clean/df_clean.csv'), 
  col_types = cols(.default = "?", state = "i", mode_type = "i")))
```

# Identify and mark fermentation periods

Identify fermentation periods by eye

```{r eval=FALSE}

(p1 <- ggplot(data = df, aes(x = time)) +
  geom_line(aes(y = brix, colour = "brix"), colour = "steelblue"))

(p2 <- ggplot(data = df, aes(x = time)) +
  geom_line(aes(y = status, colour = "status"), colour = "tomato"))

gridExtra::grid.arrange(p1, p2, nrow = 2)

plotly::ggplotly(p1)
plotly::ggplotly(p2)

```

## Automatic fermentation curve detection

Looking at above historic brix data, the first step should be to detect the overall fermentation periods fenced in by

- a starting point characterised by a steep rise in brix from a very low level to ~30brix 
- an end point characterised by a steep fall in brix into low levels

### Detect fermentation periods

Step 1) Focus on brix changes

Step 2) Identify fermentation periods

  - Identify fermentation start periods by simply mark all brix values above 15 brix (brix only ever goes beyond ~ +15 when a fermentation starts)
  - Next identify all potential end periods by marking all brix values below -100 brix. There will be many and only a few will be real end points
  - We identify the real end points in the next step
  
Step 3) Fill the NA's between ferment start periods and end periods

  - This step will close the gaps only between real start and real end points of a fermentation

Step 4) Remove the false end points

  - Now we can remove all false end points as they are not preceded by a "ferment" value

Step 5) Label the first fermentation value "start"

Step 6) Label each fermentation period by its fermentation start date 

Step 7) Add a ferment boolean

```{r}

# Identify fermentation periods
df_brix <- df %>%
  select(time, brix) %>%
  # filter(time > lubridate::ymd("2021-09-14"), time < "2021-10-20") %>%
  
  # 1) keep only brix changes.
  filter(brix != lag(brix)) %>% 

  # 2) Find fermentation start and end.
  mutate(ferment = case_when(
    brix > 15 ~ "ferment",
    lead(brix) < -100 ~ "end",
    TRUE ~ NA_character_
  )) %>% 

  # 3) Fill time points during start and end.
  fill(ferment) %>%

  # 4) Remove redundant "end" values.
  mutate(keep = if_else(ferment == "ferment" | (ferment == "end" & lag(ferment) == "ferment"), TRUE, FALSE)) %>%
  mutate(ferment = if_else(keep, ferment, NA_character_)) %>%
  select(-keep) %>%

  # 5) Rename first fermentation value to "start".
  mutate(ferment = if_else(ferment == "ferment" & is.na(lag(ferment)), "start", ferment)) %>%

  # 6) Name the fermentation period with their start date.
  mutate(ferment_period = case_when(
    ferment == "start" ~ format(time, "%y-%b-%d"),
    ferment == "end" ~ "end",
    TRUE ~ NA_character_))  %>%
  fill(ferment_period) %>%
  mutate(ferment_period = if_else(ferment_period == "end", lag(ferment_period), ferment_period)) %>%
  mutate(ferment_period = if_else(ferment_period == "end", NA_character_, ferment_period)) %>%

  # 7) Add a boolean ferment variable
  mutate(ferment_bool = if_else(!is.na(ferment), TRUE, FALSE))

# View(df_brix)
  
```

Above algo detected a false positive fermentation period for tank C1 that lingered constantly at around 20 brix before falling off. 

The following procedure defends against that case by:

1. Creating a rolling average with a window of 10 
2. Checking if rolling brix before drop off was above 10 
3. It should be around 0 or below, so if it's high it's defined as a false positive and the whole period will be removed

```{r}

# Detect and isolate false positive fermentation periods.
false_periods <- df_brix %>% 
  mutate(brix_rolling = RcppRoll::roll_mean(brix, 10, align = "right", fill = NA_real_)) %>% 
  mutate(remove_period = ifelse(ferment == "end" & lag(brix_rolling) > 10, TRUE, FALSE)) %>% 
  filter(remove_period) %>% 
  distinct(ferment_period) %>% 
  pull(ferment_period)

# Remove them.
df_brix <- df_brix %>% 
  mutate(
    ferment_bool = ifelse(ferment_period %in% false_periods, FALSE, ferment_bool),
    ferment_period = ifelse(ferment_bool, ferment_period, NA_character_),
    ferment = ifelse(ferment_bool, ferment, NA_character_),
  )

```

Check ferment periods visually:

```{r eval=FALSE}

# Check ferment start and end are detected correctly.
starts <- df_brix %>% 
  mutate(ferment = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(ferment)

ends <- df_brix %>% 
  mutate(ferment = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(ferment)

p <- df_brix %>% 
  ggplot() +
    geom_line(aes(x = time, y = brix)) + 
    geom_vline(aes(xintercept = starts), colour = 'cornflowerblue') +
    geom_vline(aes(xintercept = ends), colour = 'maroon')

plotly::ggplotly(p)

```

### Detect fermentation phases

#### Intro

Each fermentation curve is assumed to go through 3 phases characterised by four points:

- Phase 1: pre-fermentation (flat curve at initial brix value of ~20/30)
- Phase 2: fermentation (falling curve)
- Phase 3: post-fermentation (flat curve ataround just sub-zero brix)

These are defined by four points and one extra points:

- Point 1 (Phase 1 start): brix up to fermentation start levels (around 20/30 brix)
- Point 2 (Phase 2 start): brix decreases notable
- Point 3 (Phase 2 still): brix inflection - cell division decreases
- Point 4 (Phase 3 start): brix mainly remains below 0 but still fluctuates
- Point 5 (Phase 3 end): brix remains below 0 and steady and/or drops off

We identified Points (1) and (4) above. Points (2) and (3) are  hard to find in the raw data, so we should try with a LOESS smoothed curve.

#### Detect pre-fermentation start (point 1) and post-fermentation end (point 5)

Straight forward as points and phase startes coincide with start and end:

```{r}

# Add start and end phases. 
df_brix <- df_brix %>% 
  mutate(ferment_phase = case_when(
    ferment == "start" ~ "Phase 1",
    ferment == "end" ~ "Phase 3",
    TRUE ~ NA_character_,
    )
  )

# Add points 1 and 4
df_brix <- df_brix %>% 
  mutate(ferment_point = case_when(
    ferment == "start" ~ "Point 1",
    ferment == "end" ~ "Point 5",
    TRUE ~ NA_character_,
    )
  )

```


#### Detect fermentation start (point 2) | interactive

Let's try with the base LOESS function first, if that ain't working let's throw STL at it (time series loess):

Base LOESS doesn't work on the full data, but on particular fermentation periods, they do. Let's try with them:

```{r eval=FALSE}

# Focus in on just fermentation periods
df_brix %>% 
  group_by(ferment_period) %>% 
  count()

```

Now let's see what works for point (2):

```{r eval=FALSE}

# Create base data (based on specific periods)
df_test <-df_brix %>% 
  rowid_to_column("index") %>% 
  filter(ferment_period == "21-Nov-01")
  # filter(ferment_period == "21-Oct-02")
  # filter(ferment_period == "21-Oct-19")
  # filter(ferment_period == "21-Sep-14")
  # filter(ferment_period == "22-Aug-21")
  # filter(ferment_period == "22-Sep-09")
  # filter(ferment_period == "22-Sep-21")
  # filter(ferment_period == "22-Sep-29")
  
# Add LOESS smoother.
span <- 0.25
result <- loess(brix ~ index, data = df_test, span = span)
df_test$brix_smooth <- predict(result)

# Add gradient.
df_test <- df_test %>% 
  mutate(brix_deriv_1 = pracma::gradient(df_test$brix_smooth))

# Chart.
p1 <- df_test %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix)) +
    geom_line(aes(y = brix_smooth), colour = "tomato") 

p2 <- df_test %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1), colour = "blue")

gridExtra::grid.arrange(p1, p2, nrow = 2)

plotly::ggplotly(p1)
plotly::ggplotly(p2)

```

So after trying every above combination I think we should stick to simple and take the first row with brix_deriv_1 < 0 as point(2)

Sometimes that's a bit too early, sometimes it's too late. We can conceive a more complex algo (e.g. taking into account raw brix values or gradients) but let's just keep it simple for now :)

#### Detect fermentation start and end points (points 2, 4) | automatic

For each fermentation period calculate the LOESS smoothed brixed curve and the smooth brix curve's gradient. We'll use the smooth brix' gradient to detect points 2 (curve start), 3 (curve inflection) and 4 (curve end).

We do so by using tidyr's nesting on`df_brix` which creates list-columns for each fermentation_period in order for loess and gradient to run on each nested period data.

Note, we're using r's base loess function with a span we set to **0.15**. 0.25 worked well on D1 but botched a fermentation on C1 which needed a tighter curve. 

We're NOT using STL decomposition with a loess trend as this is just much simpler :))

ADD:

And we calculate the brix' 2nd derivative, which we use to help build point 2. Point 2 is the start of the fermentation which is conditioned by:

1. The brix curve needs tobe falling (1st derivative < 0)
2. The brix curve's fall needs to be accelerating (our eyes need to slide across a bump) (2nd derivative < 0)

```{r}

# Add smooth brix and gradient.
df_brix_smooth <- df_brix %>% 
  # Row number column required for loess calc
  rowid_to_column("index") %>% 

  # Get brix_smooth and brix_deriv_1
  # ---------------------------------

  # Nest by ferment_period (funny xyntax, but hey)
  tidyr::nest(data = c(-ferment_period, index, time, brix)) %>% 

  # for each ferment_period calculate the loess for their fitted values
  # and the gradient for the resulting brix_smooth values.
  mutate(
    m = purrr::map(data, loess, formula = brix ~ index, span = 0.15),
    brix_smooth = purrr::map(m, `[[`, "fitted"),
    brix_deriv_1 = purrr::map(brix_smooth, pracma::gradient)
  ) %>% 

  # remove the model variable
  select(-m) %>%

  # Unnest the data :)
  tidyr::unnest(cols = c(data, brix_smooth, brix_deriv_1)) %>% 

  # Get brix_deriv_1_smooth
  # ------------------------

  # Nest by ferment_period and get the gradient in
  tidyr::nest(data = c(-ferment_period, index, time, brix_smooth, brix_deriv_1)) %>% 

  # for each ferment_period calculate also smooth ouy the brix_deriv_1.
  mutate(
    m = purrr::map(data, loess, formula = brix_deriv_1 ~ index, span = 0.25),
    brix_deriv_1_smooth = purrr::map(m, `[[`, "fitted"),
    brix_deriv_2 = purrr::map(brix_deriv_1_smooth, pracma::gradient)
  ) %>% 

  # remove the model variable
  select(-m) %>%

  # Unnest the data :)
  tidyr::unnest(cols = c(data, brix_deriv_1_smooth, brix_deriv_2)) %>% 

  # Clean up
  select(-index) %>% 
  select(time, brix, brix_smooth, brix_deriv_1, brix_deriv_1_smooth, brix_deriv_2, ferment, ferment_bool, ferment_period, ferment_point, ferment_phase) %>% 
  arrange(time)

```

Check visually:

```{r eval=FALSE}

# Check brix_smooth and brix_deriv_1 make sense.
df_brix_smooth %>% 
  filter(ferment_bool) %>% 
  # filter(ferment_period == '22-Sep-12') %>% # CHANGED
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix)) +
    geom_line(aes(y = brix_smooth), color = "tomato") +
    facet_wrap(~ferment_period, scales = "free", nrow = 2)

df_brix_smooth %>%
  filter(ferment_bool) %>%
  # filter(ferment_period == '22-Sep-12') %>% # CHANGED 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'steelblue') +
    facet_wrap(~ferment_period, scales = "free", nrow = 2)

df_brix_smooth %>%
  filter(ferment_bool) %>%
  # filter(ferment_period == '22-Sep-12') %>% # CHANGED 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru') +
    facet_wrap(~ferment_period, scales = "free", nrow = 2)


```

Add points 2 and 4, the start of phase 2 and the start of phase 3 next. Note, that there can be fermentation curves that skip phase 1 altogether and start with phase 2 straight away. Simple reason being, the smooth brix curve gradient is negative from the start.

```{r}

# Get the times for point 2: brix starts falling
ferment_curve_start_times <- df_brix_smooth %>% 
  filter(ferment_bool, brix_deriv_1 < 0, brix_deriv_2 < 0) %>% 
  group_by(ferment_period) %>% 
  mutate(first = first(brix_deriv_1)) %>% 
  ungroup() %>% 
  select(time, first) %>% 
  distinct(first, .keep_all = TRUE) %>% 
  pull(time)

# Get the times for point (4): brix below 0
ferment_curve_end_times <- df_brix_smooth %>% 
  filter(ferment_bool, brix_smooth < 0) %>% 
  group_by(ferment_period) %>% 
  mutate(first = first(brix_smooth)) %>% 
  ungroup() %>% 
  select(time, first) %>% 
  distinct(first, .keep_all = TRUE) %>% 
  pull(time)

# Add them to the `ferment_phase` var.
df_brix_smooth <- df_brix_smooth %>% 
  mutate(ferment_point = if_else(time %in% ferment_curve_start_times, "Point 2", ferment_point)) %>% 
  mutate(ferment_phase = if_else(time %in% ferment_curve_start_times, "Phase 2", ferment_phase)) %>% 
  mutate(ferment_point = if_else(time %in% ferment_curve_end_times, "Point 4", ferment_point)) %>% 
  mutate(ferment_phase = if_else(time %in% ferment_curve_end_times, "Phase 3", ferment_phase))

```

#### Detect cell division inflection point

```{r}

# Get the times for point (2)
inflection_times <- df_brix_smooth %>% 
  filter(ferment_bool, brix_deriv_1 < 0) %>% 
  group_by(ferment_period) %>% 
  fill(ferment_phase) %>% 
  filter(ferment_phase == 'Phase 2') %>% 
  slice_min(brix_deriv_1_smooth) %>% 
  ungroup() %>% 
  pull(time)

# Add them to the `ferment_phase` var.
df_brix_smooth <- df_brix_smooth %>% 
  mutate(ferment_point = if_else(time %in% inflection_times, "Point 3", ferment_point)) 
  # Note, that for now, inflection point is still part of phase 2

```

#### Check points and phases

Check for all points over brix visually (note, that non-fermentation brix can be wildly off as it's potentially loess'ed across a loong period:

```{r}

# Check ferment start (point 1 - brix starts falling), end (point  - brix below 05) and the newly created points 2 and 4 are detected correctly.
starts <- df_brix_smooth %>% 
  mutate(check = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(check)

ends <- df_brix_smooth %>% 
  mutate(check = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(check)

point_2s <- df_brix_smooth %>% 
  mutate(check = ifelse(ferment_point == "Point 2", time, NA_Date_)) %>% 
  pull(check)

point_3s <- df_brix_smooth %>% 
  mutate(check = ifelse(ferment_point == "Point 3", time, NA_Date_)) %>% 
  pull(check)

point_4s <- df_brix_smooth %>% 
  mutate(check = ifelse(ferment_point == "Point 4", time, NA_Date_)) %>% 
  pull(check)

(p <- df_brix_smooth %>% 
  ggplot() +
    geom_line(aes(x = time, y = brix)) + 
    geom_line(aes(x = time, y = brix_smooth), colour = 'tomato') + 
    geom_vline(aes(xintercept = starts), colour = 'cornflowerblue') +
    geom_vline(aes(xintercept = ends), colour = 'maroon') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon')
)

plotly::ggplotly(p)

```

Check visually by fermentation period:

```{r}

# Check brix_smooth and brix_deriv_1 make sense.
p1 <- df_brix_smooth %>%
  # filter(!is.na(ferment_period)) %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix)) +
    geom_line(aes(y = brix_smooth), colour = 'tomato') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")

p2 <- df_brix_smooth %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")
  
p3 <- df_brix_smooth %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")
  
plotly::ggplotly(p1)
plotly::ggplotly(p2)
plotly::ggplotly(p3)

```

#### Fill the fermentation phases

```{r}

# Fill phases
df_brix_smooth <- df_brix_smooth %>% 
  fill(ferment_phase) %>% mutate(ferment_phase = ifelse(ferment_bool, ferment_phase, NA_real_))

```

Finally create a period stamped fermentation phase variable...

```{r}

df_brix_smooth <- df_brix_smooth %>%
  mutate(
    ferment_period_phase = ifelse(
      !is.na(ferment_period) & !is.na(ferment_phase), 
      paste(ferment_period, ferment_phase), 
      NA_character_)
    )
```

# Write brix focussed data

This is the result of above wrangle and focuses on brix and the extra fermentation variables.

```{r}

df_brix_smooth %>% write_csv(paste0(path_root, path_data, path_tank, "/clean/df_brix_smooth.csv"), na = "")

```

# Augment original data

Next we augment the original dataset with the extra fermentation variables which requires us to jump through some filling loops - particularly complex for ferment :))

```{r}

df %>% 
  left_join(df_brix_smooth, by = c("time", "brix")) %>% 

  # Fill vars which are straight fillable
  fill(brix_smooth, brix_deriv_1, brix_deriv_1_smooth, ferment_bool) %>% 
  
  # Fill `ferment` which needs drastic filling gymnastics
  mutate(ferment2 = ferment) %>% 
  fill(ferment2) %>% 
  mutate(ferment3 = ifelse(ferment2 == "ferment", ferment2, ferment)) %>% 
  mutate(ferment3 = case_when(
    ferment_bool & ferment == "start" ~ "start",
    ferment_bool & ferment2 == "end" ~ "end",
    TRUE ~ "ferment"
  )) %>% 
  mutate(ferment = ifelse(ferment_bool, ferment3, NA_character_)) %>% 
  select(-c(ferment2, ferment3)) %>% 

  # Fill ferment_period and ferment_phase which need moderate filling gymnastics
  fill(ferment_period, ferment_phase, ferment_period_phase) %>% 
  mutate(ferment_period = ifelse(ferment_bool, ferment_period, NA_character_)) %>%
  mutate(ferment_phase = ifelse(ferment_bool, ferment_phase, NA_character_)) %>%
  mutate(ferment_period_phase = ifelse(ferment_bool, ferment_period_phase, NA_character_)) %>%
  
  # Fill ferment_point which needs a moderate+ amount of filling acrobatics
  mutate(ferment_point2 = ferment_point) %>% 
  fill(ferment_point2) %>% 
  mutate(ferment_point = ifelse(ferment_bool & ferment_point2 == "Point 5", "Point 5", ferment_point)) %>% 
  select(-ferment_point2) %>% 

  # Write
  write_csv(paste0(path_root, path_data, path_tank, "/clean/df_clean_ferment.csv"), na = "")

```

