---
title: "Data prep v6 - part: manual fermentation period and phase detection"
output: html_notebook
---

This notebook is a legacy piece from the `02 prep.Rmd` that found fermentation periods and phases by eye / manually, which worked great for a single dataset to understand how they split up, distributed, and looked. 

However, this becomes tedious for many datasets and we want an automatic detection for ignition anyway so I took this out and keep it here just for historic purposes and in case we for some reaon need to fall back on it.

---

This notebook requires the dataset produced in  v6' `01 load.Rmd`. 

```{r libs, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(pracma)
detach("package:dplyr")
library(dplyr)

# Project path
path_root <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
path_data <- "/data/v6"
path_tank <- "/D1"
```

# Load required data

```{r}
(df <- readr::read_csv(
  paste0(path_root, path_data, path_tank, '/clean/df_clean.csv'), 
  col_types = cols(.default = "?", state = "i", mode_type = "i")))
```


# Identify and mark fermentation periods

Identify fermentation periods by eye

```{r}

(p1 <- ggplot(data = df, aes(x = time)) +
  geom_line(aes(y = brix, colour = "brix"), colour = "steelblue"))

(p2 <- ggplot(data = df, aes(x = time)) +
  geom_line(aes(y = status, colour = "status"), colour = "tomato"))

gridExtra::grid.arrange(p1, p2, nrow = 2)

plotly::ggplotly(p1)
plotly::ggplotly(p2)

```

## Manual fermentation curve detection

Next we identify the fermentation periods by eye. The result will be a `fermentation` variable identifying the full fermentation period plus a `fermentation_phase`variable identifying three phases for each fermentation describing the specific characteristics of the curve:

- Phase 1: pre-fermentation
- Phase 2: fermentation
- Phase 3: post-fermentation

These are defined by four "plot points":

- Point 1: brix up to fermentation start levels (around 20/30 brix)
- Point 2: brix decreases notable
- Point 3: brix mainly remains below 0 but still fluctuates
- Point 4: brix remains below 0 and steady and/or drops off

These are currently identified by eye! Automated plot point detection is possible. It would be a combination of raw data (points 1 and maybe 4) and smoothed data (points, 2, 3 and in some cases 4) but would be a bit of a project.

Here are the dates for tank D1:

2021
- Sep-14
  - 2021-09-14 13:23:29 (Point 1)
  - 2021-09-17 03:48:09 (Point 2)
  - 2021-09-23 18:43:32 (Point 3)
  - 2021-09-26 09:10:18 (Point 4)
- Oct-15
  - 2021-10-02 09:18:37 (Point 1)
  - 2021-10-05 19:49:58 (Point 2)
  - 2021-10-10 15:31:12 (Point 3)
  - 2021-10-13 09:36:14 (Point 4) quite erratic
- Oct-19
  - 2021-10-19 14:33:59 (Point 1)
  - 2021-10-22 12:08:37 (Point 2)
  - 2021-10-27 06:31:00 (Point 3)
  - 2021-10-30 13:10:28 (Point 4)
- Nov-01
  - 2021-11-01 11:10:29 (Point 1)
  - 2021-11-01 19:32:49 (Point 2)
  - 2021-11-04 21:54:27 (Point 3)
  - 2021-11-13 05:21:18 (Point 4) veery long and erratic phase 3

2022
- Aug-21
  - 2022-08-21 11:37:45 (Point 1)
  - 2022-08-22 07:32:01 (Point 2)
  - 2022-09-01 07:54:44 (Point 3)
  - 2022-09-07 07:28:28 (Point 4) no steady phase, just drop off
- Sep-09
  - 2022-09-09 07:18:13 (Point 1)
  - 2022-09-12 01:35:38 (Point 2)
  - 2022-09-16 22:50:10 (Point 3)
  - 2022-09-19 16:46:19 (Point 4) very very noisy phase 3, followed by drop off
- Sep-21
  - 2022-09-21 12:51:17 (Point 1)
  - 2022-09-24 00:47:58 (Point 2)
  - 2022-09-27 18:03:00 (Point 3)
  - 2022-09-28 20:11:54 (Point 4) drop off
- Sep-29
  - 2022-09-29 08:42:11 (Point 1)
  - 2022-10-03 08:33:32 (Point 2)
  - 2022-10-07 07:53:34 (Point 3)
  - 2022-10-09 19:45:56 (Point 4) drop off

```{r eval=FALSE}

# Set fermentation periods.
df <- df %>%
  mutate(ferment_period = case_when(
    time >= ymd_hms("2021-09-14 13:23:29") & time <= ymd_hms("2021-09-26 09:10:18") ~ "21-Sep-14",
    time >= ymd_hms("2021-10-02 09:18:37") & time <= ymd_hms("2021-10-13 09:36:14") ~ "21-Oct-15",
    time >= ymd_hms("2021-10-19 14:33:59") & time <= ymd_hms("2021-10-30 13:10:28") ~ "21-Oct-19",
    time >= ymd_hms("2021-11-01 11:10:29") & time <= ymd_hms("2021-11-13 05:21:18") ~ "21-Nov-01",
    time >= ymd_hms("2022-08-21 11:37:45") & time <= ymd_hms("2022-09-07 07:28:28") ~ "22-Aug-21",
    time >= ymd_hms("2022-09-09 07:18:13") & time <= ymd_hms("2022-09-19 16:46:19") ~ "22-Sep-09",
    time >= ymd_hms("2022-09-21 12:51:17") & time <= ymd_hms("2022-09-28 20:11:54") ~ "22-Sep-21",
    time >= ymd_hms("2022-09-29 08:42:11") & time <= ymd_hms("2022-10-09 19:45:56") ~ "22-Sep-29",
    TRUE ~ "no-ferment"
  ))

# Set fermentation period phases.
df <- df %>%
  mutate(ferment_phase = case_when(
    time >= ymd_hms("2021-09-14 13:23:29") & time < ymd_hms("2021-09-17 03:48:09") ~ "Phase 1",
    time >= ymd_hms("2021-09-17 03:48:09") & time < ymd_hms("2021-09-23 18:43:32") ~ "Phase 2",
    time >= ymd_hms("2021-09-23 18:43:32") & time <= ymd_hms("2021-09-26 09:10:18") ~ "Phase 3",
    time >= ymd_hms("2021-10-02 09:18:37") & time < ymd_hms("2021-10-05 19:49:58") ~ "Phase 1",
    time >= ymd_hms("2021-10-05 19:49:58") & time < ymd_hms("2021-10-10 15:31:12") ~ "Phase 2",
    time >= ymd_hms("2021-10-10 15:31:12") & time <= ymd_hms("2021-10-13 09:36:14") ~ "Phase 3",
    time >= ymd_hms("2021-10-19 14:33:59") & time < ymd_hms("2021-10-22 12:08:37") ~ "Phase 1",
    time >= ymd_hms("2021-10-22 12:08:37") & time < ymd_hms("2021-10-27 06:31:00") ~ "Phase 2",
    time >= ymd_hms("2021-10-27 06:31:00") & time <= ymd_hms("2021-10-30 13:10:28") ~ "Phase 3",
    time >= ymd_hms("2021-11-01 11:10:29") & time < ymd_hms("2021-11-01 19:32:49") ~ "Phase 1",
    time >= ymd_hms("2021-11-01 19:32:49") & time < ymd_hms("2021-11-04 21:54:27") ~ "Phase 2",
    time >= ymd_hms("2021-11-04 21:54:27") & time <= ymd_hms("2021-11-13 05:21:18") ~ "Phase 3",
    time >= ymd_hms("2022-08-21 11:37:45") & time < ymd_hms("2022-08-22 07:32:01") ~ "Phase 1",
    time >= ymd_hms("2022-08-22 07:32:01") & time < ymd_hms("2022-09-01 07:54:44") ~ "Phase 2",
    time >= ymd_hms("2022-09-01 07:54:44") & time <= ymd_hms("2022-09-07 07:28:28") ~ "Phase 3",
    time >= ymd_hms("2022-09-09 07:18:13") & time < ymd_hms("2022-09-12 01:35:38") ~ "Phase 1",
    time >= ymd_hms("2022-09-12 01:35:38") & time < ymd_hms("2022-09-16 22:50:10") ~ "Phase 2",
    time >= ymd_hms("2022-09-16 22:50:10") & time <= ymd_hms("2022-09-19 16:46:19") ~ "Phase 3",
    time >= ymd_hms("2022-09-21 12:51:17") & time < ymd_hms("2022-09-24 00:47:58") ~ "Phase 1",
    time >= ymd_hms("2022-09-24 00:47:58") & time < ymd_hms("2022-09-27 18:03:00") ~ "Phase 2",
    time >= ymd_hms("2022-09-27 18:03:00") & time <= ymd_hms("2022-09-28 20:11:54") ~ "Phase 3",
    time >= ymd_hms("2022-09-29 08:42:11") & time < ymd_hms("2022-10-03 08:33:32") ~ "Phase 1",
    time >= ymd_hms("2022-10-03 08:33:32") & time < ymd_hms("2022-10-07 07:53:34") ~ "Phase 2",
    time >= ymd_hms("2022-10-07 07:53:34") & time <= ymd_hms("2022-10-09 19:45:56") ~ "Phase 3",
    TRUE ~ "no-ferment"
  ))

# Set fermentation helper variables.
df <- df %>%
  mutate(
    ferment = ifelse(ferment_period == "no-ferment", "No fermentation", "Fermentation"),
    ferment_bool = ifelse(ferment_period == "no-ferment", FALSE, TRUE),
    ferment_phase_long = paste(ferment_period, ferment_phase)
  )

```

Let's 's look at the fermentation curves and phases:

```{r eval=FALSE}

p <- df %>% 
  filter(ferment_period != "no-ferment") %>% 
  ggplot() +
    geom_line(aes(x = time, y = brix, color = ferment_phase)) +
    facet_wrap(~ferment_period, scales = "free_x")

plotly::ggplotly(p)

```

Check if the new vars make sense:

- ferment_period
- ferment
- ferment_bool
- ferment_phase
- ferment_phase_long

```{r eval=FALSE}

df %>% 
  group_by(ferment, ferment_bool, ferment_period, ferment_phase, ferment_phase_long) %>% 
  count()

```
