---
title: "Pulse cooling identification | v6"
output: html_notebook
---

This notebook requires the dataset `df_clean_ferment.csv` produced in  v6' `02 ferment.Rmd` which is the main df plus the additional fermentation variables.

The notebook creates a df_pulse summarising pulse colling periods.

```{r libs, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
detach("package:dplyr")
library(dplyr)

# Project path
path_root <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
path_data <- "/data/v6"
path_tank <- "/F1"
```

# Load required data

```{r}
(df <- readr::read_csv(
  paste0(path_root, path_data, path_tank, '/clean/df_clean_ferment.csv'), 
  col_types = cols(
    .default = "?", 
    abv = "d",
    state = "i",
    mode_type = "i",
    ferment = "c",
    ferment_period = "c",
    ferment_point = "c",
    ferment_phase = "c",
    ferment_period_phase = "c"))
 )
```

# Build pulse cooling dataset

Here we want to focus in on the pulse periods. We do so by 

1. identify the pulses in the main data (resulting in `df_pulse_marked`)
2. build a pulse focused dataset (`df_pulse`)

## Check fermentation curves and statuses

```{r eval=FALSE}

# How does status look over time?
ggplot(df) +
  geom_line(aes(x = time, y = status))

# How many different statuses have we got?
df %>% 
  group_by(status_label) %>% 
  count()

# How many different statuses do we get for each fermentation period?
df %>% 
  group_by(ferment_period, status_label) %>% 
  count()

# Vis brix
df %>% 
  filter(ferment_bool) %>%   
    ggplot() +
      geom_line(aes(x = time, y = brix)) +
      facet_wrap(~ferment_period, scales = "free_x")

# Vis status
df %>% 
  filter(ferment_bool) %>%   
    ggplot() +
      geom_line(aes(x = time, y = status)) +
      facet_wrap(~ferment_period, scales = "free_x")

```

## Identify pulse periods

This is based on v3's `.../v4 - band change non-fermentation/02 create pulse data.Rmd` with some major changes and improvements (above all excluding the manual step!).

Steps 1) and 2) focus in on the status events only

Step 3) creates a time diff variable for between each status change

Step 4) creates the `pulse` variable based on 2/3 oscillation

Step 5) check for each pulse that it's not too long (threshold: 600)

Step 6) adds identifiers for start and end events to each pulse period

Step 7) expels all pulse periods that only have 2 pulses

Step 8) adds a unique identifier to each pulse period so we can group and analyse them individually

Step 9) just reshuffles the data a little

```{r}

df_pulse_marked <- df %>% 
  # 1) Keep status events only.
  filter(!is.na(status)) %>% 
  
  # 2) Keep only rows with status changes.
  mutate(status_change = ifelse(lag(status) == status, FALSE, TRUE)) %>% 
  filter(status_change) %>% 
  select(-status_change) %>% 
  
  # 3) Calculate time difference between status changes.
  mutate(diff = as.numeric(lead(time) - time)) %>% 
  
  # 4) Create the pulse var based on the status pulse patterns (2-3-2-...).
  mutate(pulse = ifelse(
    lag(status) != status & lead(status) != status
    & ((status == 3 | status == 2) & (lag(status) == 3 | lag(status) == 2) & (lead(status) == 3 | lead(status) == 2))
    , 'pulse', NA)) %>% 
  
  # 5) Untag pulses longer than 600 seconds.
  mutate(pulse = ifelse(diff > 600, NA, pulse)) %>% 
  
  # 6) Add start and end identifier to each pulse section.
  mutate(pulse = ifelse(is.na(lag(pulse)) & pulse == 'pulse' & lead(pulse) == 'pulse', 'start', pulse)) %>%
  mutate(pulse = ifelse(lag(pulse) == 'pulse' & pulse == 'pulse' & is.na(lead(pulse)), 'end', pulse)) %>% 
  
  # 7) Untag pulses with just two pulses (need two mutates as it has to work sequentially).
  mutate(pulse = ifelse(is.na(lag(pulse)) & is.na(lead(pulse, n = 2)), NA, pulse)) %>%
  mutate(pulse = ifelse(is.na(lag(pulse)) & is.na(lead(pulse)), NA, pulse)) %>%
  
  # 8) Add a unique identifier to each pulse period.
  mutate(pulse_group = ifelse(pulse == 'start', as.character(time), NA)) %>%
  fill(pulse_group) %>%
  mutate(pulse_group = ifelse(is.na(pulse), NA, pulse_group)) %>% 

  # 9) Reshuffle.
  select(time, temp, setpoint, status, brix, diff, pulse, pulse_group, everything())

```

Check the outcome...

```{r eval=FALSE}

df_pulse_marked
# view(df_pulse_marked)

```

Add the pulse_group var to the full dataset and write it...

```{r}

df_clean_ferment_pulse <- df %>% 
  left_join(df_pulse_marked %>% select(time, pulse, pulse_group), by = "time") %>% 
  mutate(pulse2 = pulse, pulse_group2 = pulse_group) %>% 
  fill(pulse2, pulse_group2) %>%
  mutate(
    pulse = ifelse(pulse2 == "pulse", pulse2, pulse),
    pulse_group = ifelse(pulse2 == "pulse", pulse_group2, pulse_group)
  ) %>% 
  select(-c(pulse2, pulse_group2))

df_clean_ferment_pulse %>% write_csv(paste0(path_root, path_data, path_tank, "/clean/df_clean_ferment_pulse.csv"), na = "")

```



## Create pulse dataset

This is based on `.../v4 - band change non-fermentation/03 analysis v4.Rmd`` with some changes - mainly recognising some new variables / variable names.

In a next step we produce a dataset that just focuses on the pulse cooling periods. Each row in this dataset is a pulse cooling period and each variable summarises a specific pulse period characteristic:

```{r}

# Number of modulations.
pulse_num <- df_pulse_marked %>% 
  group_by(pulse_group) %>% 
  count() %>% 
  mutate(pulse_group_start = lubridate::parse_date_time(pulse_group, orders = "Y-m-d H:M:S")) %>% 
  arrange(pulse_group_start) %>%
  ungroup()

# Modulation summary variables.
df_modulation <- df_pulse_marked %>% 
  filter(!is.na(pulse)) %>%
  mutate(
    status_diff_2 = ifelse(status == 2, diff, NA_real_),
    status_diff_3 = ifelse(status == 3, diff, NA_real_)
  ) %>% 
  group_by(pulse_group) %>% 
  summarise(
    pulse_dur_2_mean_sec = mean(status_diff_2, na.rm = TRUE), # avg non-cooling seconds
    pulse_dur_3_mean_sec = mean(status_diff_3, na.rm = TRUE), # avg cooling seconds
    pulse_dur_2 = sum(status_diff_2, na.rm = TRUE) / 60, # non-cooling in minutes
    pulse_dur_3 = sum(status_diff_3, na.rm = TRUE) / 60, # cooling in minutes
  ) %>% 
  mutate(
    cooling_percent = pulse_dur_3 / (pulse_dur_2 + pulse_dur_3), # cooling percent
  ) %>% 
  ungroup()

df_modulation

```


```{r}

# Main pulse period data.
df_pulse <- df_pulse_marked %>%
  filter(pulse == "start" | pulse == "end") %>% 
  select(time, ferment, ferment_bool, ferment_period, ferment_phase, ferment_period_phase, pulse, pulse_group, temp) %>% 
  mutate(temp_diff = ifelse(pulse == "start", as.numeric(lead(temp) - temp), NA))  %>% 
  mutate(pulse_duration = ifelse(pulse == "start", as.numeric(lead(time) - time), NA))  %>% 
  filter(pulse == "start") %>% 
  select(-pulse) %>%
  left_join(pulse_num %>% select(pulse_group, pulse_number = n), by = "pulse_group") %>% 
  mutate(pulse_frequency = pulse_number / pulse_duration) %>% 
  mutate(roc = temp_diff / (pulse_duration/60))  # temperature change per hour (pulse_duration is in minutes, hence /60 is in hours)

# Add modulation variables to main data. 
(df_pulse <- df_pulse %>%
  left_join(df_modulation, by = "pulse_group") %>% 
  select(
    time,
    ferment,
    ferment_bool,
    ferment_period,
    ferment_phase,
    ferment_period_phase,
    pulse_group,
    temp,
    temp_diff,
    pulse_duration,
    roc,
    pulse_dur_2_mean_sec,
    pulse_dur_3_mean_sec,
    cooling_percent,
    pulse_dur_2,
    pulse_dur_3,
    pulse_number,
    pulse_frequency,
  ))

```
# Write

```{r}

df_pulse %>% write_csv(paste0(path_root, path_data, path_tank, "/clean/df_pulse_periods.csv"), na = "")

```

