---
title: "Fermentation curve analysis | v6"
output: html_notebook
---

This notebook gets and analysis all fermentation curve data.

```{r libs, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
detach("package:dplyr")
library(dplyr)

# Project path
path_root <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
path_data <- "/data/v6"
path_tanks <- c("/C1","/D1","/E1","/F1")
```

# Load and prep

Load all available data - collating `df`'s

```{r load-data}

load_data <- function(dir) {
  print(dir)

  path <- paste0(path_root, path_data, dir, "/clean/df_brix_smooth.csv")

  df <- read_csv(path, 
          col_types = cols(
            .default = "?", 
            ferment = "c",
            ferment_period = "c",
            ferment_point = "c",
            ferment_phase = "c",
            ferment_period_phase = "c")
          )
  
  df <- df %>% mutate(tank = str_replace(dir, "/", ""))
  
  return(df)
}

df_list <- lapply(path_tanks, load_data)

df <- bind_rows(df_list) %>% as_tibble()

```
```{r}
df
```

Only keep fermentation curves (this is the same as df_brix_smooth_all_tanks.csv):

```{r filter-non-ferment}

df <- df %>% filter(ferment_bool)

```

Check and write

```{r check-data}

skimr::skim(df)

df %>% group_by(ferment_period) %>% count()

```

# Visualise

Visualise all 

```{r visualise}

# Get a vector of points to draw the ferment point lines.
starts <- df %>% 
  mutate(check = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(check)

ends <- df %>% 
  mutate(check = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(check)

point_2s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 2", time, NA_Date_)) %>% 
  pull(check)

point_3s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 3", time, NA_Date_)) %>% 
  pull(check)

point_4s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 4", time, NA_Date_)) %>% 
  pull(check)

# Plot
(p1 <- df %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix), color = "#cccccc") +
    geom_line(aes(y = brix_smooth), colour = 'hotpink', size = 0.75) +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))

(p2 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'cornflowerblue') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))
  
(p3 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))

(p4 <- df %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = temp), color = "#666666") +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))
  
plotly::ggplotly(p1)
plotly::ggplotly(p2)
plotly::ggplotly(p3)
plotly::ggplotly(p4)

```

Show fermentations brix and temp in batches and order:

```{r eval=FALSE}

# Get all ferment periods.
df %>% 
  group_by(ferment_period) %>% 
  count() %>% 
  pull(ferment_period)

# MANUAL! Go into VS code, order them and split them into their tanks
tank_c_ferments <- c("C1-21-Sep-03", "C1-22-Sep-12", "C1-22-Sep-29", "C1-21-Oct-01")
tank_d_ferments <- c("D1-21-Sep-14", "D1-21-Oct-02", "D1-21-Oct-19", "D1-21-Nov-01", "D1-22-Sep-09", "D1-22-Sep-21", "D1-22-Sep-29", "D1-22-Aug-21")
tank_e_ferments <- c("E1-21-Sep-09", "E1-21-Oct-29", "E1-21-Sep-28", "E1-22-Sep-10", "E1-22-Sep-22", "E1-22-Oct-03")
tank_f_ferments <- c("F1-21-Jun-03", "F1-22-Aug-26", "F1-21-Sep-11", "F1-21-Sep-30")

# ---

build_brix_temp_charts <- function(tank_ferments, interactive = FALSE) {
  for (fermentation in tank_ferments) {
    
    # Filter the data for each fermentation period.
    df_filtered <- df %>% 
      filter(ferment_period == fermentation)
    
    # Set the label points.
    starts <- df_filtered %>% 
      mutate(check = ifelse(ferment == "start", time, NA_Date_)) %>% 
      pull(check)
  
    ends <- df_filtered %>% 
      mutate(check = ifelse(ferment == "end", time, NA_Date_)) %>% 
      pull(check)
    
    point_2s <- df_filtered %>% 
      mutate(check = ifelse(ferment_point == "Point 2", time, NA_Date_)) %>% 
      pull(check)
    
    point_3s <- df_filtered %>% 
      mutate(check = ifelse(ferment_point == "Point 3", time, NA_Date_)) %>% 
      pull(check)
    
    point_4s <- df_filtered %>% 
      mutate(check = ifelse(ferment_point == "Point 4", time, NA_Date_)) %>% 
      pull(check)
    
    # Brix chart (also removing the x axis).
    p1 <- df_filtered %>%
      ggplot(aes(x = time)) +
        geom_hline(aes(yintercept = 0), alpha = 0.35) +
        geom_line(aes(y = brix), color = "#cccccc") +
        geom_line(aes(y = brix_smooth), colour = 'hotpink') +
        geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
        geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
        geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
        ylim(-10, NA) +
        labs(title = fermentation) +
        theme(plot.title = element_text(color="#999999",face="bold"),
              axis.ticks.x = element_blank(),
              axis.text.x = element_blank(),
              axis.title.x = element_blank())
    
    # Temp chart.
    p2 <- df_filtered %>%
      ggplot(aes(x = time)) +
        geom_line(aes(y = temp), color = "#999999") +
        geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
        geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
        geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
        theme(axis.title.x = element_blank())
  
    if (!interactive) {
      # Print static.
      suppressWarnings(gridExtra::grid.arrange(p1, p2, nrow = 2))
    } else {
      # Print interactive.
      pl1 <- plotly::ggplotly(p1)
      pl2 <- plotly::ggplotly(p2)
      
      print(plotly::subplot(pl1, pl2, nrows = 2))
    }
  }
}

# Show static charts.
build_brix_temp_charts(tank_c_ferments)
build_brix_temp_charts(tank_d_ferments)
build_brix_temp_charts(tank_e_ferments)
build_brix_temp_charts(tank_f_ferments)

# Show interactive charts.
build_brix_temp_charts(tank_d_ferments, interactive = TRUE)
build_brix_temp_charts(tank_c_ferments, interactive = TRUE)
build_brix_temp_charts(tank_e_ferments, interactive = TRUE)
build_brix_temp_charts(tank_f_ferments, interactive = TRUE)

```

# Profiling

Variables to profile:

- duration of fermentation phase
- duration of fermentation curve
- distance between start and inflection point
- distance between inflection point and end
- brix levels at each point
- temperatures at each point


```{r data-prep}

# Check how many periods we have in total.
df %>% distinct(ferment_period)

# Add duration between way points.
df_days <- df %>% 
  filter(!is.na(ferment_point)) %>% 
  mutate(diff_days_all = difftime(time, lag(time), units = "days") %>% as.numeric) %>% 
  select(tank, ferment_period, ferment_point, starts_with("diff"))

# Calculate durations from fermentation phase start.
df_days_full <- df_days %>% 
  group_by(ferment_period) %>% 
  mutate(diff_days_all = ifelse(row_number() == 1, 0, diff_days_all)) %>% 
  mutate(diff_from_phase_start = cumsum(diff_days_all)) %>% 
  ungroup()

# Calculate durations from fermentation curve start.
df_days_curve <- df_days_full %>% 
  filter(ferment_point != "Point 1", ferment_point != "Point 5") %>% 
  group_by(ferment_period) %>% 
  mutate(diff_days_curve = ifelse(ferment_point == "Point 2", 0, diff_days_all)) %>% 
  mutate(diff_from_curve_start = cumsum(diff_days_curve)) %>% 
  ungroup()

# Merge.
df_waypoints <- left_join(df_days_full, 
                          df_days_curve %>% 
                            select(ferment_period, ferment_point, diff_from_curve_start), 
                          by = c("ferment_period", "ferment_point"))

# Add brix level at each point.
df_waypoints <- left_join(df_waypoints, 
                          df %>%
                            filter(!is.na(ferment_point)) %>% 
                            select(ferment_period, ferment_point, brix, brix_smooth, temp), 
                          by = c("ferment_period", "ferment_point"))
 

df_waypoints
df_waypoints %>% skimr::skim()

```

Show the normal duration and brix values.

```{r normal-values}

# Density graph function:
show_freqs <- function(data, var, median_var, mean_var) {
  # Prep vars
  var = enquo(var)
  mean_var = enquo(mean_var)
  median_var = enquo(median_var)
  
# Make plot
p <- ggplot(data) +
  geom_histogram(aes(x = !!var), bins = 50, alpha = 0.2) +
  geom_density(aes(x = !!var, y = 0.5 * ..count..)) +
  geom_vline(aes(xintercept = !!mean_var)) +
  geom_vline(aes(xintercept = !!median_var), linetype = "dashed", alpha = 0.6) +
  geom_label(aes(x = !!mean_var, y = Inf, label = round(!!mean_var, 1)), vjust = 1.2, hjust = -0.2) +
  facet_wrap(~ferment_point_label, scales = "free") +
  theme(strip.text.x = element_text(size = 9, face = "bold", colour = "#666666"))
  
  # Showplot
  print(p)
}

# Plot data with mean and median data.
df_normals_per_line <- df_waypoints %>% 

  # Get facet labels ready.
  mutate(ferment_point_label = case_when(
    ferment_point == "Point 1" ~ "Point 1 | Phase start",
    ferment_point == "Point 2" ~ "Point 2 | Curve start",
    ferment_point == "Point 3" ~ "Point 3 | Inflection",
    ferment_point == "Point 4" ~ "Point 4 | Curve end",
    ferment_point == "Point 5" ~ "Point 5 | Phase end",
    TRUE ~ "NONE",
  )) %>% 

  # For each ferment_period...
  group_by(ferment_point) %>% 

  # ...get means and median values for each of the numeric columns.
  mutate(mean_diff_days_all = mean(diff_days_all)) %>% 
  mutate(mean_diff_from_phase_start = mean(diff_from_phase_start)) %>% 
  mutate(mean_diff_from_curve_start = mean(diff_from_curve_start)) %>% 
  mutate(mean_brix_smooth = mean(brix_smooth)) %>% 
  mutate(mean_temp = mean(temp)) %>% 

  mutate(median_diff_days_all = median(diff_days_all)) %>% 
  mutate(median_diff_from_phase_start = median(diff_from_phase_start)) %>% 
  mutate(median_diff_from_curve_start = median(diff_from_curve_start)) %>% 
  mutate(median_brix_smooth = median(brix_smooth)) %>% 
  mutate(median_temp = median(temp)) 

df_normals_per_line

# Visualise.
show_freqs(df_normals_per_line, diff_days_all, mean_diff_days_all, median_diff_days_all)
show_freqs(df_normals_per_line, diff_from_phase_start, mean_diff_from_phase_start, median_diff_from_phase_start)
show_freqs(df_normals_per_line, diff_from_curve_start, mean_diff_from_curve_start, median_diff_from_curve_start)
show_freqs(df_normals_per_line, brix_smooth, mean_brix_smooth, median_brix_smooth)
show_freqs(df_normals_per_line, temp, mean_temp, median_temp)

```

------------------

See how we can fence in the inflection points (DELETE)

```{r}
# JUST GET THE values we need
df_normals_per_line %>% 
  select(ferment_period, ferment_point_label, diff_from_curve_start) %>% 
  filter(ferment_point == "Point 3") %>% 
  arrange(diff_from_curve_start)

df_normals_per_line %>% 
  select(ferment_period, ferment_point_label, diff_from_curve_start) %>% 
  filter(ferment_point == "Point 3") %>% 
  summary()



```

------------------

Let's check the variance by checking SD of Brix, duration and temp for each point. 

```{r variance}

# Get the SD and the CV for each numeric variable grouped by waypoint.
df_normals_skimmed <- df_normals_per_line %>% 
  select(ferment_point, starts_with("diff"), brix_smooth, temp) %>% 
  group_by(ferment_point) %>% 
  skimr::skim() %>% 
  as_tibble() %>% 
  # Adding the coefficient of variation
  mutate(numeric.cv = numeric.sd / abs(numeric.mean) * 100)

# Focus dataset (we only need three vars).
df_normals_skimmed <- df_normals_skimmed %>% 
  select(skim_variable, ferment_point, numeric.mean, numeric.sd, numeric.cv) %>% 
  filter(skim_variable == "brix_smooth" | skim_variable == "diff_from_curve_start" | skim_variable == "temp")

df_normals_skimmed

# Plot mean.
df_normals_skimmed %>% 
  select(skim_variable, ferment_point, starts_with("numeric")) %>% 
  ggplot() +
    geom_bar(aes(x = ferment_point, y = numeric.mean), stat = "identity") +
    geom_label(aes(x = ferment_point, 
                   y = numeric.mean, label = round(numeric.mean, 1)),
             stat = "identity", nudge_y = 0.75) +
    facet_wrap(~skim_variable)

# Plot SD.
df_normals_skimmed %>% 
  select(skim_variable, ferment_point, starts_with("numeric")) %>% 
  ggplot() +
    geom_bar(aes(x = ferment_point, y = numeric.sd), stat = "identity") +
    geom_label(aes(x = ferment_point, 
                   y = numeric.sd, label = round(numeric.sd, 1)),
             stat = "identity", nudge_y = 0.75) +
    facet_wrap(~skim_variable)

# Plot CV.
df_normals_skimmed %>% 
  select(skim_variable, ferment_point, starts_with("numeric")) %>% 
  ggplot() +
    geom_bar(aes(x = ferment_point, y = numeric.cv), stat = "identity") +
    geom_label(aes(x = ferment_point, 
                   y = numeric.cv, label = scales::percent(numeric.cv / 100, accuracy = 1)),
               stat = "identity", nudge_y = 6) +
    facet_wrap(~skim_variable)

```

Variance is notable in parts:

- Brix variance across Point 1 and 2 (phase start and curve start) relatively stable. It's quite stable across Point 3 (inflection) with an SD of +/- 5 Brix (28% CV)
- The time in days between fermentation start and inflection can vary +/- 1.4 days which is notable given that the average is 2.8. The end date variability is with 2.6 days significantly higher in absolute terms but a little lower in relative terms as Point 4 is happening much later (6.4 days in average).
- Temperature at fermentation start and inflection point can vary at around 7.7째F and at fermentation end point around 4.2째F, which is low in CV as the temperatures are relatively high (between 60 and 80째F)


# Normal curve

First we produce a spline based on the average fermentation phase waypoints. Then we extract that continuous spline data which we can finally take to overlay all fermentation phase lines.

First, we produce the splines and extract that data:

```{r}

# Filter out just the normal values per point.
df_normals <- df_normals_per_line %>%
  ungroup() %>% 
  distinct(ferment_point, .keep_all = TRUE) %>% 
  select(starts_with(c("ferment_point", "mean", "median")))

# The curve will go straight throug to waypoint 5 if we don't add a 
# helper point close to point 5 with a brix just below 0 to mimic the other curves.
df_normal_helper <- df_normals %>% 
  filter(ferment_point == "Point 5") %>% 
  mutate(ferment_point = "Point 5 helper") %>% 
  mutate(mean_diff_from_phase_start = mean_diff_from_phase_start - 1) %>% 
  mutate(median_diff_from_phase_start = median_diff_from_phase_start - 1) %>% 
  mutate(mean_brix_smooth = -5)

df_normals_plot <- bind_rows(df_normals, df_normal_helper) %>% 
  arrange(factor(ferment_point, levels = c("Point 1", "Point 2", "Point 3", "Point 4", "Point 5 helper", "Point 5")))

# Plot the mean and median curves with a spline based on the points.
x_ticks <- seq(0, max(df_normals_plot$mean_diff_from_phase_start) + 1)

(p <- df_normals_plot %>% ggplot(aes(y = mean_brix_smooth)) +
  geom_hline(aes(yintercept = 0), alpha = 0.25) +
  geom_point(aes(x = mean_diff_from_phase_start), alpha = 0.5) +
  ggalt::geom_xspline(aes(x = mean_diff_from_phase_start), alpha = 0.5, linetype = "dashed") +
  geom_point(aes(x = median_diff_from_phase_start)) +
  ggalt::geom_xspline(aes(x = median_diff_from_phase_start)) +
  scale_x_continuous(breaks = x_ticks) +
  labs(x = "Days", y = "Brix"))

# Extract the spline data from the plot.
(df_mean_line <- ggplot_build(p)$data[[3]] %>% 
  as_tibble() %>% 
  select(day = x, brix_smooth = y))

(df_median_line <- ggplot_build(p)$data[[5]] %>% 
  as_tibble() %>% 
  select(day = x, brix_smooth = y))

# Check.
ggplot(df_median_line, aes(x = day, y = brix_smooth)) + geom_line()
ggplot(df_mean_line, aes(x = day, y = brix_smooth)) + geom_line()

```

Then we overlay all ferment phase lines with the mean and median lines.

```{r}

# Create data with a column that counts up days from phase start to end.
df_days <- df %>%
  group_by(ferment_period) %>% 
  mutate(
    day_diff = difftime(time, lag(time), units = "days"),
    day_diff = ifelse(is.na(day_diff), 0, day_diff),
    day = cumsum(day_diff),
  ) %>%
  ungroup() %>% 
  select(ferment_period, day, brix_smooth)

# Visualise lines with mean and median lines.
# No labels
ggplot() +
  # Core stuff
  geom_hline(aes(yintercept = 0), alpha= 0.2) +
  geom_line(data = df_days, aes(x = day, y = brix_smooth, group = ferment_period), alpha = 0.25, size = 0.25) +
  geom_line(data = df_median_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 1) +
  geom_line(data = df_mean_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 0.5, linetype = "dashed")

# Duration labels from phase start
ggplot() +
  # Core stuff
  geom_hline(aes(yintercept = 0), alpha= 0.2) +
  geom_line(data = df_days, aes(x = day, y = brix_smooth, group = ferment_period), alpha = 0.25, size = 0.25) +
  geom_line(data = df_median_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 1) +
  geom_line(data = df_mean_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 0.5, linetype = "dashed") +
  # Ornaments
  geom_vline(data = df_normals, aes(xintercept = median_diff_from_phase_start, colour = ferment_point)) +
  geom_label(data = df_normals, aes(label = round(median_diff_from_phase_start, 1), x = median_diff_from_phase_start), y = -50, 
                                    alpha = 0.85, label.r = unit(0, "lines")) +
  scale_color_manual(values = c("darkgreen", "darkseagreen", "steelblue", "salmon", "darkred")) +
  theme(legend.position = "none")

# Duration labels from fermentation start
ggplot() +
  # Core stuff
  geom_hline(aes(yintercept = 0), alpha= 0.2) +
  geom_line(data = df_days, aes(x = day, y = brix_smooth, group = ferment_period), alpha = 0.25, size = 0.25) +
  geom_line(data = df_median_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 1) +
  geom_line(data = df_mean_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 0.5, linetype = "dashed") +
  # Ornaments
  geom_vline(data = df_normals, aes(xintercept = median_diff_from_phase_start, colour = ferment_point)) +
  geom_label(data = df_normals, aes(label = round(median_diff_from_curve_start, 1), x = median_diff_from_phase_start), y = -50, 
                                    alpha = 0.85, label.r = unit(0, "lines")) +
  scale_color_manual(values = c("darkgreen", "darkseagreen", "steelblue", "salmon", "darkred")) +
  theme(legend.position = "none")

# Duration and Brix labels
ggplot() +
  # Core stuff
  geom_hline(aes(yintercept = 0), alpha= 0.2) +
  geom_line(data = df_days, aes(x = day, y = brix_smooth, group = ferment_period), alpha = 0.25, size = 0.25) +
  geom_line(data = df_median_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 1) +
  geom_line(data = df_mean_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 0.5, linetype = "dashed") +
  # Ornaments
  geom_vline(data = df_normals, aes(xintercept = median_diff_from_phase_start, colour = ferment_point)) +
  geom_label(data = df_normals, aes(label = round(median_diff_from_phase_start, 1), x = median_diff_from_phase_start), y = -50, 
                                    alpha = 0.85, label.r = unit(0, "lines")) +
  geom_label(data = df_normals %>% filter(ferment_point != "Point 5"), 
             aes(label = round(median_brix_smooth, 1), x = median_diff_from_phase_start, y = median_brix_smooth), 
             alpha = 0.85, label.r = unit(0, "lines")) +
  scale_color_manual(values = c("darkgreen", "darkseagreen", "steelblue", "salmon", "darkred")) +
  theme(legend.position = "none")

# Duration, Brix and temperature labels
ggplot() +
  # Core stuff
  geom_hline(aes(yintercept = 0), alpha= 0.2) +
  geom_line(data = df_days, aes(x = day, y = brix_smooth, group = ferment_period), alpha = 0.25, size = 0.25) +
  geom_line(data = df_median_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 1) +
  geom_line(data = df_mean_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 0.5, linetype = "dashed") +
  # Day labels
  geom_vline(data = df_normals, aes(xintercept = median_diff_from_phase_start, colour = ferment_point)) +
  geom_label(data = df_normals, aes(label = round(median_diff_from_phase_start, 1), x = median_diff_from_phase_start), y = -50, 
                                    alpha = 0.85, label.r = unit(0, "lines")) +
  # Brix labels
  geom_label(data = df_normals %>% filter(ferment_point != "Point 5"), 
             aes(label = round(median_brix_smooth, 1), x = median_diff_from_phase_start, y = median_brix_smooth), 
             alpha = 0.75, label.r = unit(0, "lines"), nudge_y = - 3) +
  # Temp labels
  geom_label(data = df_normals %>% filter(ferment_point != "Point 5"),
             aes(label = paste0(round(median_temp, 0), "째"), x = median_diff_from_phase_start, y = 35),
             alpha = 0.85, label.r = unit(0, "lines")) +
  # General style
  scale_color_manual(values = c("darkgreen", "darkseagreen", "steelblue", "salmon", "darkred")) +
  theme(legend.position = "none")

```

```{r}
df_normals
```


---

TODO:

- Show Chart 1 above first (explaining mean and meadian ideal curve)
- then show a new version of above aith just the duration labels (in the bottom)
- lastly show the full chart 2 above

After that make the point that variance is generally large by showing some of the histograms/densities from above

---

Here is a previous version using loess smoothing of all lines, which might be good to keep but of no real value as we've sorted it above...

```{r eval=FALSE}

# Get a row index to overlay all fermentation curves.
df_index <- df %>% 
  group_by(ferment_period) %>% 
  mutate(index = row_number()) %>% 
  ungroup()

write_csv(df_index, "test.csv")

# Visualise with line.
ggplot(df_index, aes(x = index, y = brix_smooth)) +
  geom_hline(aes(yintercept = 0), alpha = 0.2) +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0, ymax = -Inf), fill = "white", alpha = 0.05) +
  geom_line(aes(group = ferment_period), alpha = 0.2) +
  geom_smooth(se = FALSE, method = "loess", span = 0.2)

# Build plot to harvest its data for the single line below 
# (w/o hline as that cocks up the final data)..
p <- ggplot(df_index, aes(x = index, y = brix_smooth)) +
  geom_line(aes(group = ferment_period), alpha = 0.2) +
  geom_smooth(se = FALSE, method = "loess", span = 0.2)

# Get the loess average curve data.
df_ideal_curve <- ggplot_build(p)$data[[2]] %>% 
  as_tibble() %>% 
  select(x, y)

# Show just ideal curve
ggplot(df_ideal_curve, aes(x = x, y = y)) +
  geom_line()

```
