---
title: "Fermentation curve analysis | v6"
output: html_notebook
---

This notebook gets and analysis all fermentation curve data.

```{r libs, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
detach("package:dplyr")
library(dplyr)

# Project path
path_root <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
path_data <- "/data/v6"
path_tanks <- c("/C1","/D1","/E1","/F1")
```

# Load and prep

Load all available data - collating `df`'s

```{r load-data}

load_data <- function(dir) {
  print(dir)

  path <- paste0(path_root, path_data, dir, "/clean/df_brix_smooth.csv")

  df <- read_csv(path, 
          col_types = cols(
            .default = "?", 
            ferment = "c",
            ferment_period = "c",
            ferment_point = "c",
            ferment_phase = "c",
            ferment_period_phase = "c")
          )
  
  df <- df %>% mutate(tank = str_replace(dir, "/", ""))
  
  return(df)
}

df_list <- lapply(path_tanks, load_data)

df <- bind_rows(df_list) %>% as_tibble()

```

Only keep fermentation curves:

```{r filter-non-ferment}

df <- df %>% filter(ferment_bool)

```

Check

```{r check-data}

skimr::skim(df)

df %>% group_by(ferment_period) %>% count()

```

# Visualise

Visualise all 

```{r visualise}

# Get a vector of points to draw the ferment point lines.
starts <- df %>% 
  mutate(check = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(check)

ends <- df %>% 
  mutate(check = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(check)

point_2s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 2", time, NA_Date_)) %>% 
  pull(check)

point_3s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 3", time, NA_Date_)) %>% 
  pull(check)

point_4s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 4", time, NA_Date_)) %>% 
  pull(check)

# Plot
p1 <- df %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix)) +
    geom_line(aes(y = brix_smooth), colour = 'tomato') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")

p2 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")
  
p3 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")
  
plotly::ggplotly(p1)
plotly::ggplotly(p2)
plotly::ggplotly(p3)

```


# Mean curve

Get the ideal curve:

```{r}

# Get a row index to overlay all fermentation curves.
df_index <- df %>% 
  group_by(ferment_period) %>% 
  mutate(index = row_number()) %>% 
  ungroup()

write_csv(df_index, "test.csv")

# Visualise with line.
ggplot(df_index, aes(x = index, y = brix_smooth)) +
  geom_hline(aes(yintercept = 0), alpha = 0.2) +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0, ymax = -Inf), fill = "white", alpha = 0.05) +
  geom_line(aes(group = ferment_period), alpha = 0.2) +
  geom_smooth(se = FALSE, method = "loess", span = 0.2)

# Build plot to harvest its data for the single line below 
# (w/o hline as that cocks up the final data)..
p <- ggplot(df_index, aes(x = index, y = brix_smooth)) +
  geom_line(aes(group = ferment_period), alpha = 0.2) +
  geom_smooth(se = FALSE, method = "loess", span = 0.2)

# Get the loess average curve data.
df_ideal_curve <- ggplot_build(p)$data[[2]] %>% 
  as_tibble() %>% 
  select(x, y)

# Show just ideal curve
ggplot(df_ideal_curve, aes(x = x, y = y)) +
  geom_line()

```

This is ok, but one problem is the index - we'd rather have evenly spaced timepoints. The other problem is that loess does a good job in most areas (notably the start) but -probably due to the variable length time series distorts the more in moves towards the end.

So an alternative algorithm would need to: 

1. Evenly space the timepoints of each curve
2. Resample the data to have equal length
3. Take an average/median for each value

See also here for a [matlab answer](https://www.mathworks.com/matlabcentral/answers/1463604-plotting-an-average-of-multiple-time-series-of-different-length)

I tried with normalising the start date and the duration. But that didn't quite work out:

```{r}

# Rather than an index, let's get a time vector starting at time = 0.
df_time_diff <- df %>%
  group_by(ferment_period) %>% 
  mutate(
    time_diff = difftime(time, lag(time), units = "days"),
    time_diff = ifelse(is.na(time_diff), 0, time_diff),
    time_diff_cum = cumsum(time_diff),
    time_new = as.Date("2023-01-01") + as.difftime(time_diff_cum, units = "days")
  ) %>%
  ungroup() 

df_time_diff %>%
  select(time, time_diff, time_diff_cum, time_new)

df_time_diff %>%
  select(time, time_diff, time_diff_cum, time_new) %>% 
  summary()

# Visualise.
(p <- ggplot(df_time_diff, aes(x = time_new, y = brix_smooth)) +
  geom_line(aes(group = ferment_period), alpha = 0.2) +
  geom_smooth(se = FALSE, method = "loess", span = 0.2))

# Get the loess average curve data.
df_ideal_curve <- ggplot_build(p)$data[[2]] %>% 
  as_tibble() %>% 
  select(x, y)

# Show just ideal curve
ggplot(df_ideal_curve, aes(x = x, y = y)) +
  geom_line()


```

Problem here is the unevenly spaced time series for each curve I believe? If that's the case, a solution might be to convert to evenly spaced (uniform across all curves) event durations. 

Some more pointers [here](https://www.google.com/search?q=r+convert+unevenly+spaced+time+series+to+evenly+spaced+time+series&rlz=1C5CHFA_enDE796DE796&oq=r+convert+unevenly+spaced+time+series+to+evenly+spaced+time+series&aqs=chrome..69i57j69i64l3.16854j0j1&sourceid=chrome&ie=UTF-8)

TODO

# Profiling

Nice!

Now let's get the base data covering:

- duration of fermentation phase
- duration of fermentation curve
- distance between start and inflection point
- distance between inflection point and end
- brix levels at each point


```{r data-prep}

# Check how many periods we have in total.
df %>% distinct(ferment_period)

# Add duration between way points.
df_days <- df %>% 
  filter(!is.na(ferment_point)) %>% 
  mutate(diff_days_all = difftime(time, lag(time), units = "days") %>% as.numeric) %>% 
  select(tank, ferment_period, ferment_point, starts_with("diff"))

# Calculate durations from fermentation phase start.
df_days_full <- df_days %>% 
  group_by(ferment_period) %>% 
  mutate(diff_days_all = ifelse(row_number() == 1, 0, diff_days_all)) %>% 
  mutate(diff_from_phase_start = cumsum(diff_days_all)) %>% 
  ungroup()

# Calculate durations from fermentation curve start.
df_days_curve <- df_days_full %>% 
  filter(ferment_point != "Point 1", ferment_point != "Point 5") %>% 
  group_by(ferment_period) %>% 
  mutate(diff_days_curve = ifelse(ferment_point == "Point 2", 0, diff_days_all)) %>% 
  mutate(diff_from_curve_start = cumsum(diff_days_curve)) %>% 
  ungroup()

# Merge.
df_waypoints <- left_join(df_days_full, 
                          df_days_curve %>% 
                            select(ferment_period, ferment_point, diff_from_curve_start), 
                          by = c("ferment_period", "ferment_point"))

# Add brix level at each point.
df_waypoints <- left_join(df_waypoints, 
                          df %>%
                            filter(!is.na(ferment_point)) %>% 
                            select(ferment_period, ferment_point, brix, brix_smooth), 
                          by = c("ferment_period", "ferment_point"))
 

df_waypoints
df_waypoints %>% skimr::skim()

```

Show the normal duration and brix values.

```{r normal-values}

# Density graph function:
show_freqs <- function(data, var, mean_var, median_var) {
  # Prep vars
  var = enquo(var)
  mean_var = enquo(mean_var)
  median_var = enquo(median_var)
  
# Make plot
p <- ggplot(data) +
  geom_histogram(aes(x = !!var), bins = 50, alpha = 0.2) +
  geom_density(aes(x = !!var, y = 0.5 * ..count..)) +
  geom_vline(aes(xintercept = !!mean_var)) +
  geom_vline(aes(xintercept = !!median_var), linetype = "dashed", alpha = 0.6) +
  geom_label(aes(x = !!mean_var, y = Inf, label = round(!!mean_var, 1)), vjust = 1.2, hjust = -0.2) +
  facet_wrap(~ferment_point_label, scales = "free") +
  theme(strip.text.x = element_text(size = 9, face = "bold", colour = "#666666"))
  
  # Showplot
  print(p)
}

# Plot data with mean and median data.
df_plot <- df_waypoints %>% 
  # Get facet labels ready.
  mutate(ferment_point_label = case_when(
    ferment_point == "Point 1" ~ "Point 1 | Phase start",
    ferment_point == "Point 2" ~ "Point 2 | Curve start",
    ferment_point == "Point 3" ~ "Point 3 | Inflection",
    ferment_point == "Point 4" ~ "Point 4 | Curve end",
    ferment_point == "Point 5" ~ "Point 5 | Phase end",
    TRUE ~ "NONE",
  )) %>% 
  # For each ferment_period...
  group_by(ferment_point) %>% 
  # ...get means and median values for each of the numeric columns.
  mutate(mean_diff_days_all = mean(diff_days_all)) %>% 
  mutate(mean_diff_from_phase_start = mean(diff_from_phase_start)) %>% 
  mutate(mean_diff_from_curve_start = mean(diff_from_curve_start)) %>% 
  mutate(mean_brix_smooth = mean(brix_smooth)) %>% 
  mutate(median_diff_days_all = median(diff_days_all)) %>% 
  mutate(median_diff_from_phase_start = median(diff_from_phase_start)) %>% 
  mutate(median_diff_from_curve_start = median(diff_from_curve_start)) %>% 
  mutate(median_brix_smooth = median(brix_smooth))

# Visualise.
show_freqs(df_plot, diff_days_all, mean_diff_days_all, median_diff_days_all)
show_freqs(df_plot, diff_from_phase_start, mean_diff_from_phase_start, median_diff_from_phase_start)
show_freqs(df_plot, diff_from_curve_start, mean_diff_from_curve_start, median_diff_from_curve_start)
show_freqs(df_plot, brix_smooth, mean_brix_smooth, median_brix_smooth)

```















