---
title: "Fermentation curve analysis | v6"
output: html_notebook
---

This notebook gets and analysis all fermentation curve data.

```{r libs, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
detach("package:dplyr")
library(dplyr)

# Project path
path_root <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
path_data <- "/data/v6"
path_tanks <- c("/C1","/D1","/E1","/F1")
```

# Load and prep

Load all available data - collating `df`'s

```{r load-data}

load_data <- function(dir) {
  print(dir)

  path <- paste0(path_root, path_data, dir, "/clean/df_brix_smooth.csv")

  df <- read_csv(path, 
          col_types = cols(
            .default = "?", 
            ferment = "c",
            ferment_period = "c",
            ferment_point = "c",
            ferment_phase = "c",
            ferment_period_phase = "c")
          )
  
  df <- df %>% mutate(tank = str_replace(dir, "/", ""))
  
  return(df)
}

df_list <- lapply(path_tanks, load_data)

df <- bind_rows(df_list) %>% as_tibble()

```

Only keep fermentation curves:

```{r filter-non-ferment}

df <- df %>% filter(ferment_bool)

```

Check

```{r check-data}

skimr::skim(df)

df %>% group_by(ferment_period) %>% count()

```

# Visualise

Visualise all 

```{r visualise}

# Get a vector of points to draw the ferment point lines.
starts <- df %>% 
  mutate(check = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(check)

ends <- df %>% 
  mutate(check = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(check)

point_2s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 2", time, NA_Date_)) %>% 
  pull(check)

point_3s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 3", time, NA_Date_)) %>% 
  pull(check)

point_4s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 4", time, NA_Date_)) %>% 
  pull(check)

# Plot
p1 <- df %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix)) +
    geom_line(aes(y = brix_smooth), colour = 'tomato') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")

p2 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")
  
p3 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'khaki') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free")
  
plotly::ggplotly(p1)
plotly::ggplotly(p2)
plotly::ggplotly(p3)

```

# Profiling

Variables to profile:

- duration of fermentation phase
- duration of fermentation curve
- distance between start and inflection point
- distance between inflection point and end
- brix levels at each point


```{r data-prep}

# Check how many periods we have in total.
df %>% distinct(ferment_period)

# Add duration between way points.
df_days <- df %>% 
  filter(!is.na(ferment_point)) %>% 
  mutate(diff_days_all = difftime(time, lag(time), units = "days") %>% as.numeric) %>% 
  select(tank, ferment_period, ferment_point, starts_with("diff"))

# Calculate durations from fermentation phase start.
df_days_full <- df_days %>% 
  group_by(ferment_period) %>% 
  mutate(diff_days_all = ifelse(row_number() == 1, 0, diff_days_all)) %>% 
  mutate(diff_from_phase_start = cumsum(diff_days_all)) %>% 
  ungroup()

# Calculate durations from fermentation curve start.
df_days_curve <- df_days_full %>% 
  filter(ferment_point != "Point 1", ferment_point != "Point 5") %>% 
  group_by(ferment_period) %>% 
  mutate(diff_days_curve = ifelse(ferment_point == "Point 2", 0, diff_days_all)) %>% 
  mutate(diff_from_curve_start = cumsum(diff_days_curve)) %>% 
  ungroup()

# Merge.
df_waypoints <- left_join(df_days_full, 
                          df_days_curve %>% 
                            select(ferment_period, ferment_point, diff_from_curve_start), 
                          by = c("ferment_period", "ferment_point"))

# Add brix level at each point.
df_waypoints <- left_join(df_waypoints, 
                          df %>%
                            filter(!is.na(ferment_point)) %>% 
                            select(ferment_period, ferment_point, brix, brix_smooth), 
                          by = c("ferment_period", "ferment_point"))
 

df_waypoints
df_waypoints %>% skimr::skim()

```

Show the normal duration and brix values.

```{r normal-values}

# Density graph function:
show_freqs <- function(data, var, median_var, mean_var) {
  # Prep vars
  var = enquo(var)
  mean_var = enquo(mean_var)
  median_var = enquo(median_var)
  
# Make plot
p <- ggplot(data) +
  geom_histogram(aes(x = !!var), bins = 50, alpha = 0.2) +
  geom_density(aes(x = !!var, y = 0.5 * ..count..)) +
  geom_vline(aes(xintercept = !!mean_var)) +
  geom_vline(aes(xintercept = !!median_var), linetype = "dashed", alpha = 0.6) +
  geom_label(aes(x = !!mean_var, y = Inf, label = round(!!mean_var, 1)), vjust = 1.2, hjust = -0.2) +
  facet_wrap(~ferment_point_label, scales = "free") +
  theme(strip.text.x = element_text(size = 9, face = "bold", colour = "#666666"))
  
  # Showplot
  print(p)
}

# Plot data with mean and median data.
df_normals_per_line <- df_waypoints %>% 
  # Get facet labels ready.
  mutate(ferment_point_label = case_when(
    ferment_point == "Point 1" ~ "Point 1 | Phase start",
    ferment_point == "Point 2" ~ "Point 2 | Curve start",
    ferment_point == "Point 3" ~ "Point 3 | Inflection",
    ferment_point == "Point 4" ~ "Point 4 | Curve end",
    ferment_point == "Point 5" ~ "Point 5 | Phase end",
    TRUE ~ "NONE",
  )) %>% 
  # For each ferment_period...
  group_by(ferment_point) %>% 
  # ...get means and median values for each of the numeric columns.
  mutate(mean_diff_days_all = mean(diff_days_all)) %>% 
  mutate(mean_diff_from_phase_start = mean(diff_from_phase_start)) %>% 
  mutate(mean_diff_from_curve_start = mean(diff_from_curve_start)) %>% 
  mutate(mean_brix_smooth = mean(brix_smooth)) %>% 
  mutate(median_diff_days_all = median(diff_days_all)) %>% 
  mutate(median_diff_from_phase_start = median(diff_from_phase_start)) %>% 
  mutate(median_diff_from_curve_start = median(diff_from_curve_start)) %>% 
  mutate(median_brix_smooth = median(brix_smooth))

# Visualise.
show_freqs(df_normals_per_line, diff_days_all, mean_diff_days_all, median_diff_days_all)
show_freqs(df_normals_per_line, diff_from_phase_start, mean_diff_from_phase_start, median_diff_from_phase_start)
show_freqs(df_normals_per_line, diff_from_curve_start, mean_diff_from_curve_start, median_diff_from_curve_start)
show_freqs(df_normals_per_line, brix_smooth, mean_brix_smooth, median_brix_smooth)

```

# Normal curve

First we produce a spline based on the average fermentation phase waypoints. Then we extract that continuous spline data which we can then take to overlay all fermentation phase lines.

First, we produce the splines and extract that data:

```{r}

# Filter out just the normal values per point.
df_normals <- df_normals_per_line %>%
  ungroup() %>% 
  distinct(ferment_point, .keep_all = TRUE) %>% 
  select(starts_with(c("ferment_point", "mean", "median")))

# The curve will go straight throug to waypoint 5 if we don't add a 
# helper point close to point 5 with a brix just below 0 to mimic the other curves.
df_normal_helper <- df_normals %>% 
  filter(ferment_point == "Point 5") %>% 
  mutate(ferment_point = "Point 5 helper") %>% 
  mutate(mean_diff_from_phase_start = mean_diff_from_phase_start - 1) %>% 
  mutate(median_diff_from_phase_start = median_diff_from_phase_start - 1) %>% 
  mutate(mean_brix_smooth = -5)

df_normals_plot <- bind_rows(df_normals, df_normal_helper) %>% 
  arrange(factor(ferment_point, levels = c("Point 1", "Point 2", "Point 3", "Point 4", "Point 5 helper", "Point 5")))

# Plot the mean and median curves with a spline based on the points.
x_ticks <- seq(0, max(df_normals_plot$mean_diff_from_phase_start) + 1)

(p <- df_normals_plot %>% ggplot(aes(y = mean_brix_smooth)) +
  geom_hline(aes(yintercept = 0), alpha = 0.25) +
  geom_point(aes(x = mean_diff_from_phase_start), alpha = 0.5) +
  ggalt::geom_xspline(aes(x = mean_diff_from_phase_start), alpha = 0.5, linetype = "dashed") +
  geom_point(aes(x = median_diff_from_phase_start)) +
  ggalt::geom_xspline(aes(x = median_diff_from_phase_start)) +
  scale_x_continuous(breaks = x_ticks) +
  labs(x = "Days", y = "Brix"))

# Extract the spline data from the plot.
(df_mean_line <- ggplot_build(p)$data[[3]] %>% 
  as_tibble() %>% 
  select(day = x, brix_smooth = y))

(df_median_line <- ggplot_build(p)$data[[5]] %>% 
  as_tibble() %>% 
  select(day = x, brix_smooth = y))

# Check.
ggplot(df_median_line, aes(x = day, y = brix_smooth)) + geom_line()
ggplot(df_mean_line, aes(x = day, y = brix_smooth)) + geom_line()

```

Then we overlay all ferment phase lines with the mean and median lines.

```{r}

# Create data with a column that counts up days from phase start to end.
df_days <- df %>%
  group_by(ferment_period) %>% 
  mutate(
    day_diff = difftime(time, lag(time), units = "days"),
    day_diff = ifelse(is.na(day_diff), 0, day_diff),
    day = cumsum(day_diff),
  ) %>%
  ungroup() %>% 
  select(ferment_period, day, brix_smooth)

# Visualise lines with mean and median lines.
ggplot() +
  geom_line(data = df_days, aes(x = day, y = brix_smooth, group = ferment_period), alpha = 0.25, size = 0.25) +
  geom_line(data = df_median_line, aes(x = day, y = brix_smooth), colour = "steelblue", size = 1) +
  geom_line(data = df_mean_line, aes(x = day, y = brix_smooth), size = 0.5, linetype = "dashed")
    


```

TODO: Add waypoint lines and labels

---

Here is a previous version using loess smoothing of all lines, which is maybe good to keep but of no real value as we've sorted it above...

```{r eval=FALSE}

# Get a row index to overlay all fermentation curves.
df_index <- df %>% 
  group_by(ferment_period) %>% 
  mutate(index = row_number()) %>% 
  ungroup()

write_csv(df_index, "test.csv")

# Visualise with line.
ggplot(df_index, aes(x = index, y = brix_smooth)) +
  geom_hline(aes(yintercept = 0), alpha = 0.2) +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0, ymax = -Inf), fill = "white", alpha = 0.05) +
  geom_line(aes(group = ferment_period), alpha = 0.2) +
  geom_smooth(se = FALSE, method = "loess", span = 0.2)

# Build plot to harvest its data for the single line below 
# (w/o hline as that cocks up the final data)..
p <- ggplot(df_index, aes(x = index, y = brix_smooth)) +
  geom_line(aes(group = ferment_period), alpha = 0.2) +
  geom_smooth(se = FALSE, method = "loess", span = 0.2)

# Get the loess average curve data.
df_ideal_curve <- ggplot_build(p)$data[[2]] %>% 
  as_tibble() %>% 
  select(x, y)

# Show just ideal curve
ggplot(df_ideal_curve, aes(x = x, y = y)) +
  geom_line()

```
