---
title: "Inflection point refine"
format: html
editor: visual
---

## Objective

We need to find an algorithm detecting the inflection point on real time data.

### Current implementation

So far inflection point detection requires the full curve data at least from between fermentation start to fermentation end. The algo uses the full data to find the global first derivative minimum as a detection:

1.  find the minimum slope value of a decreasing brix curve (slope \< 0)
2.  \...of all brix values above 5 (to remove real low ones)

### Simple refined implementation

To start with we should implement a simple solution based on **derivatives** and **heuristics**. If this is not working we can think about a more complex solution also producing quantitative measures of uncertainty (bootstrapped linear regression or non-linear regression).

1.  take the smoothed curve (maybe heavily smoothed?)
2.  (optional) aggregate over a day (or another time period) to increase smoothing
3.  check for 2nd derivative to change sign from negative to positive ([explainer](https://www.khanacademy.org/math/ap-calculus-ab/ab-diff-analytical-applications-new/ab-5-6a/v/inflection-points))
4.  potentially add heuristics:
    -   minimum time distance to fermentation start
    -   brix range

To be tested with the existing curves.

## Prep and initial thoughts

What do we need to refine the inflection?

-   Brix and time data of all curves (time maybe indexed as periods?)
-   1st and 2nd derivative
-   A way to simulate real time brix

Next test what inflection points the 2nd derivative brings about

### Libs

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(slider)
detach("package:dplyr")
library(dplyr)

# Project path
path_root <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
path_data <- "/data/v6"
path_tank <- "/all"
```

### Load data

```{r}
df <- readr::read_csv(paste0(path_root, path_data, path_tank, "/df_brix_smooth_all_tanks.csv"))
df
```

Here's a visual overview

```{r}
# Get a vector of points to draw the ferment point lines.
starts <- df %>% 
  mutate(check = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(check)

ends <- df %>% 
  mutate(check = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(check)

point_2s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 2", time, NA_Date_)) %>% 
  pull(check)

point_3s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 3", time, NA_Date_)) %>% 
  pull(check)

point_4s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 4", time, NA_Date_)) %>% 
  pull(check)

# Plot
(p1 <- df %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix), color = "#cccccc") +
    geom_line(aes(y = brix_smooth), colour = 'hotpink', size = 0.75) +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))

(p2 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'cornflowerblue') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))
  
(p3 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))

(p4 <- df %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = temp), color = "#666666") +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))
  
plotly::ggplotly(p1)
plotly::ggplotly(p2)
plotly::ggplotly(p3)
plotly::ggplotly(p4)
```

### Single dataset

```{r}
(test <- df %>% 
  filter(ferment_period == "D1-21-Oct-02"))
```

Just look at 2nd derivative

```{r}

test %>% ggplot() +
  geom_line(aes(x = time, y = brix_smooth))

test %>% ggplot() +
  geom_line(aes(x = time, y = brix_deriv_1_smooth))

test %>% ggplot() +
  geom_line(aes(x = time, y = brix_deriv_2))
  

```

Ok! Next, let's detect negative to positive sign changes of the 2nd derivative

```{r}

# Creating a new column to store potential inflections:
test %>% 
  mutate(inflection = ifelse(lag(brix_deriv_2) < 0 & brix_deriv_2 > 0, "Inflection", ""))

# Write to file to have a better look
test %>% 
  mutate(inflection = ifelse(lag(brix_deriv_2) < 0 & brix_deriv_2 > 0, "Inflection", "")) %>% 
  select(time, brix, brix_deriv_2, ferment_point, inflection) %>% 
  write_csv("test.csv")

```

It'd be good to only calculate them after ferment start and before ferment end. But that would probably ideally be implemented once we have the other waypoint calcs set up to be real time (so we can check which phase we're in).

Then it's up for discussion if we

1.  just take the first and call it a day or
2.  show each inflection point and let the viewer decide if it;s a real one.

To start with, option (2) is my clear favourite as it allows us / the user / the system to learn.

Then we'd want to calculate these for all curves and have a looksies.

First, we add 2nd derivative based inflection points as new variable:

```{r}

df <- df %>% 
  mutate(inflection = ifelse(lag(brix_deriv_2) < 0 & brix_deriv_2 > 0, "Inflection", ""))
```

Second we visualise and have a look:

```{r visual}
# Get a vector of points to draw the ferment point lines.
starts <- df %>% 
  mutate(check = ifelse(ferment == "start", time, NA_Date_)) %>% 
  pull(check)

ends <- df %>% 
  mutate(check = ifelse(ferment == "end", time, NA_Date_)) %>% 
  pull(check)

point_2s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 2", time, NA_Date_)) %>% 
  pull(check)

point_3s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 3", time, NA_Date_)) %>% 
  pull(check)

point_2nd_deriv_inflection <- df %>% 
  mutate(check = ifelse(inflection == "Inflection", time, NA_Date_)) %>% 
  pull(check)

point_4s <- df %>% 
  mutate(check = ifelse(ferment_point == "Point 4", time, NA_Date_)) %>% 
  pull(check)

# Plot
(p1 <- df %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix), color = "#cccccc") +
    geom_line(aes(y = brix_smooth), colour = 'hotpink', size = 0.75) +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_2nd_deriv_inflection), colour = 'black', linetype = 'dotted') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    ylim(c(-5, 30)) +
    facet_wrap(~ferment_period, scales = "free"))

(p2 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'cornflowerblue') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_2nd_deriv_inflection), colour = 'black', linetype = 'dotted') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))
  
(p3 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_2), colour = 'peru') +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_2nd_deriv_inflection), colour = 'black', linetype = 'dotted') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))

(p4 <- df %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = temp), color = "#666666") +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_2nd_deriv_inflection), colour = 'black', linetype = 'dotted') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    facet_wrap(~ferment_period, scales = "free"))
  
plotly::ggplotly(p1)
plotly::ggplotly(p2)
plotly::ggplotly(p3)
plotly::ggplotly(p4)



```

So that's looking good. There are many dashed inflection point lines, but each *globally detected inflection point* (Point 3) is hit by a *2nd deriv inflection point*. This is good.

I am not really too worried about the ones coming after as a workflow (at least to start) would be to flag the inflection point up and then let a human eye determine if it is or is not **the** inflection point. Once that human eye has approved an inflection, they can turn off the detection algorithm for this fermentation.

However, we should defend against the early *2nd deriv inflection points* as these are often before the brix drop off and as such are rather unlikely to define inflection and should be removed.

As such, the next challenge is brix drop off detection.

## Drop off detection / Inflection defense

Couple of options:

1.  Identify brix drop
2.  Identify 1st derivate drops
3.  Use change point analysis (which requires the global timeseries) (so no)

Looking at the brix and its 1st and 2nd derivatives there isn't really a stark benefit the derivatives would bring to the detection. It seems the actual brix drop would be used best to identify its own drop :)

There are a few graph-knee detection algos, but some like the [kneedle](https://towardsdatascience.com/detecting-knee-elbow-points-in-a-graph-d13fc517a63c) approach require the global data, [others](https://stackoverflow.com/a/47681057/3219033) are basically a fancy derivative calculation.

So one route to

-   either the drop off detection
-   or the inflection defense

is to check for the slope to be of a certain minimum negative value and (less so) check for brix to be below a certain level

Drop off detection algo: if we're **past ferment start** and the slope is at least **negative x** == drop off point

Early Inflection defense: if there's a 2nd deirvative inflection and the slope is at least **negative x** == inflection

### Slopes at inflection point

Let's check the slope at inflection points:

```{r}
df %>% names

(check <- df %>% 
  mutate(inflection = ifelse(lag(brix_deriv_2) < 0 & brix_deriv_2 > 0, "Inflection", "")) %>% 
  filter(ferment_phase == "Phase 2") %>% 
  filter(inflection == "Inflection") %>% 
  select(time, inflection, brix_smooth, brix_deriv_1_smooth))

check %>% summary
  
```

Let's visualise that (depends on code block `visual` above):

```{r}

p1 <- df %>%
  # filter(ferment_period_phase == "D1-21-Oct-02") %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix), color = "#cccccc") +
    geom_line(aes(y = brix_smooth), colour = 'hotpink', size = 0.75) +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_2nd_deriv_inflection), colour = 'black', linetype = 'dotted') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    # geom_label(aes(x = !!mean_var, y = Inf, label = round(!!mean_var, 1)), vjust = 1.2, hjust = -0.2) +
    ylim(c(-5, 30)) +
    facet_wrap(~ferment_period, scales = "free")

df %>%
  select(time, brix_deriv_1_smooth, ferment_period, inflection) %>% 
  mutate(brix_slope_window = slider::slide_dbl(brix_deriv_1_smooth, mean, .before = 10))
  

infl_data <- df %>%
# df %>%
  select(time, brix_deriv_1_smooth, ferment_period, inflection) %>% 
  
  # Add a rolling 1st derivative average:
  mutate(brix_slope_window = slider::slide_dbl(brix_deriv_1_smooth, mean, .before = 20)) %>% 
  
  filter(inflection == "Inflection") %>% 
  group_by(ferment_period) %>% 
  mutate(index = row_number()) %>% 
  ungroup()
    

p1 + 
  geom_text(data = infl_data, 
            aes(x = time, 
                y = index * 4, 
                # The actual 1st derivative values
                # label = round(brix_deriv_1_smooth, 2)), 
                # The rolling average 1st derivative values
                label = round(brix_slope_window, 2)),
            vjust = 1.2, hjust = -0.2)

```

The rolling window doesn't change much ☹️

### Rethinking derivatives

Looking at above I wonder why the slope values are so different when the brix slopes themselves look pretty similar? I mean come to think of it, shouldn't the slope value of an irregular time series be normalised over time?

Let's try and calculate 1st derivatives normalised by time:

```{r}

df_deriv_norm <- df %>% 
  # select(time, brix_smooth) %>% 
  mutate(time_diff = as.numeric(interval(lag(time, 1), time), units = "days")) %>% 
  mutate(brix_deriv_1_norm = (brix_smooth - lag(brix_smooth)) / time_diff)

```

And now let's visualise:

```{r}

p1 <- df_deriv_norm %>%
  # filter(ferment_period_phase == "D1-21-Oct-02") %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix), color = "#cccccc") +
    geom_line(aes(y = brix_smooth), colour = 'hotpink', size = 0.75) +
    geom_vline(aes(xintercept = point_2s), colour = 'darkseagreen') +
    geom_vline(aes(xintercept = point_3s), colour = 'steelblue') +
    geom_vline(aes(xintercept = point_2nd_deriv_inflection), colour = 'black', linetype = 'dotted') +
    geom_vline(aes(xintercept = point_4s), colour = 'salmon') +
    ylim(c(-5, 30)) +
    facet_wrap(~ferment_period, scales = "free")


infl_data <- df_deriv_norm %>%
# df %>%
  # select(time, brix_deriv_1_smooth, ferment_period, inflection) %>% 
  select(time, brix_deriv_1_norm, ferment_period, inflection) %>%
  
  # Add a rolling avg for the NORMALISED 1st derivative:
  mutate(brix_deriv_1_norm_rolling = slide_dbl(brix_deriv_1_norm, mean, .before = 20)) %>% 
  
  filter(inflection == "Inflection") %>% 
  group_by(ferment_period) %>% 
  mutate(index = row_number()) %>% 
  ungroup()
    

p1 + 
  geom_text(data = infl_data, 
            aes(x = time, 
                y = index * 4, 
                # Actual derivative values
                # label = round(brix_deriv_1_norm, 2)),
                # Actual derivative values
                label = round(brix_deriv_1_norm_rolling, 2)),
            vjust = 1.2, hjust = -0.2)

```

Let's also check smooth brix vs non-normalised vs normalised 1st derivative:

```{r}

(p1 <- df_deriv_norm %>%
  # filter(ferment_period_phase == "D1-21-Oct-02") %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix), color = "#cccccc") +
    geom_line(aes(y = brix_smooth), colour = 'hotpink', size = 0.75) +
    ylim(c(-5, 30)) +
    facet_wrap(~ferment_period, scales = "free"))

(p2 <- df_deriv_norm %>%
  # filter(ferment_period_phase == "D1-21-Oct-02") %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), colour = 'steelblue', size = 0.75) +
    facet_wrap(~ferment_period, scales = "free"))

(p3 <- df_deriv_norm %>%
  # filter(ferment_period_phase == "D1-21-Oct-02") %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_norm), colour = 'grey', size = 0.75) +
    ylim(c(-20, 20)) +
    facet_wrap(~ferment_period, scales = "free"))

```

**Summary:** the time normalised slope indeed clarifies the picture as it highlights slope magnitudes at specific time points. We won't require to extend the concept to the 2nd derivative (acceleration) as the current use of the 2nd derivative is focussed on identifying a slope sign change which the non-normalised 2nd derivative covers.

### Inflection refine algo

So, after looking at the normalised and the rolling normalised, the following could work:

-   Calculate normalised slope (NS)
-   Calculate normalised with rolling avg (window = 20) (RNS)
-   Show each inflection with an RNS \>= 7 with a high visual weight (covering all current series but *E1-22-Sep-10*)
-   Show each inflection with an RNS \>= 5 but \< 7 with a low visual weight (this will cover all)

### Brix drop off algo

The time normalised slope can be used to calculate a brix drop off. An algo could check for any NS \< -10 or maybe looking into RNS values? Let's compare NS versus RNS values:

```{r}

df_deriv_norm <- df_deriv_norm %>% 
  mutate(brix_deriv_1_norm_rolling = slide_dbl(brix_deriv_1_norm, mean, .before = 20))

```

Check curves

```{r}
(p1 <- df_deriv_norm %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix), color = "#cccccc") +
    geom_line(aes(y = brix_smooth), colour = 'hotpink', size = 0.75) +
    ylim(c(-5, 30)) +
    geom_hline(yintercept = 0, color="#aaaaaa") +
    facet_wrap(~ferment_period, scales = "free"))

(p2 <- df_deriv_norm %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_norm), colour = 'skyblue', size = 0.75) +
    ylim(c(-20, 5)) +
    geom_hline(yintercept = 0, color="#aaaaaa") +
    facet_wrap(~ferment_period, scales = "free"))

(p3 <- df_deriv_norm %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_norm_rolling), colour = 'steelblue', size = 0.75) +
    ylim(c(-20, 5)) +
    geom_hline(yintercept = 0, color="#aaaaaa") +
    facet_wrap(~ferment_period, scales = "free"))

```

NS should be just fine. We should do

-   NS \> 10
-   since ferment start
-   Drop off NOT required (as in other values should not depend on it)

## Summary and next steps

### Inflection definition

This notebook developed a new algo allowing to calculate the **inflection point in real time** based on the 2nd derivative changing sign from - to +.

There can't be a single inflection point as the smoothed yet still partly noisy brix data will generate multiple inflections. However, we can isolate the most likely inflections by additionally defending against minor inflections in that we consult the **time-normalised slope** underlining inflection points that happen after steep negative slopes.

We bolster this defense by not just looking at the previous slope value but a **window** of 20 previous slope values.

### Brix drop off definition

We can further use this time-normalised slope to estimate a **brix drop off** point. As the fermentation start is rather sensitive, the brix drop off would be an additional waypoint indicating the brix to fall off significantly. This should be an informational waypoint rather than a crucial one as it might not be detected.

### Next steps

The **next notebook** should calculate each waypoint from a real-time perspective. Some of them can just be copied from `02 ferment prep.Rmd`, others would be refined from this very notebook.
