---
title: "Waypoint refactor"
format: html
editor: visual
---

## Objective

After (1) calculating waypoints in `02 ferment prep.Rmd` and (2) refining waypoints in `07 waypoint refine.qmd` this notebook's task is to cleanly create waypoint calculation functions to work in real time.

```{r libs}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(pracma)
library(RcppRoll)
detach("package:dplyr")
library(dplyr)

# Project path
path_root <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
path_data <- "/data/v6"
path_tank <- "/D1"
```

## Data

```{r}
(df <- readr::read_csv(
  paste0(path_root, path_data, path_tank, '/clean/df_clean.csv'), 
  col_types = cols(.default = "?", state = "i", abv = "d", mode_type = "i")))
```

```{r eval=FALSE}

(p1 <- ggplot(data = df, aes(x = time)) +
  geom_line(aes(y = brix, colour = "brix"), colour = "steelblue"))

plotly::ggplotly(p1)

```

## Data prep

The data output of this data prep section would be assumed to be handled on the Ignition side. In other words: all algos will expect the output of these data prep steps.

```{r}

(df_brix <- df %>%
  select(time, brix, temp) %>%
  filter(brix != lag(brix)))
  
```

## Tools

This should include:

-   Brix smooth
-   1st deriv / slope
-   1st deriv / Slope smooth
-   2nd deriv / acceleration

Further mechanics for later:

-   Add a variable uniquely labeling fermentation period and tank
-   Add a boolean ferment variable

## Fermentation start
