---
title: "Waypoint refactor"
format: html
editor: visual
---

## Objective

After (1) calculating waypoints in `02 ferment prep.Rmd` and (2) refining waypoints in `07 waypoint refine.qmd` this notebook's task is to cleanly create waypoint calculation functions to work in real time.

```{r libs}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(pracma)
library(RcppRoll)
library(scales)
detach("package:dplyr")
library(dplyr)

# Project path
path_root <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
path_data <- "/data/v6"
path_tank <- "/D1"
```

## Data

```{r}
(df <- readr::read_csv(
  paste0(path_root, path_data, path_tank, '/clean/df_clean.csv'), 
  col_types = cols(.default = "?", state = "i", abv = "d", mode_type = "i")))
```

```{r eval=FALSE}

(p1 <- ggplot(data = df, aes(x = time)) +
  geom_line(aes(y = brix, colour = "brix"), colour = "steelblue"))

plotly::ggplotly(p1)

```

## Data prep

The data output of this data prep section would be assumed to be handled on the Ignition side. In other words: all algos will expect the output of these data prep steps.

```{r}

(df <- df %>%
  select(time, brix, temp) %>%
  filter(brix != lag(brix)))
  
```

## Process

### Steps

-   Function checking each brix value (or each x brix values) coming in doing the following:
-   Check the phase:
    -   **Non-fermentation**
    -   **Fermentation**
        -   **Lag phase** (from phase start to ferment start

        -   **Exponential phase** (from ferment start to inflection). This can **optionally** further split into

            -   flat phase (before brix drop off)
            -   steep phase (after brix drop off)

        -   **Stationary phase** (from inflection to ferment end)

        -   **Post fermentation phase** (from ferment end to phase end / press off)
-   Decide what to do on fermentation boolean:
    -   If its a **non-fermentation** phase:

        -   set fermentation boolean to false
        -   return

    -   If it's a **fermentation** phase:

        -   get data from start of fermentation phase to latest value
        -   calculate tool metrics
        -   check for new waypoint:
            -   run algos for each waypoint possible in that phase
            -   run defenses for each phase (advanced)
            -   if a defense triggers update previous data (advanced)
        -   If new waypoint:
            -   mark new waypoint

            -   start new phase

            -   rinse and repeat

### Description and questions

Clarify how ignition can handle these updates. Here's what would be required:

Starting at non-fermentation for each tank, one procedure (`brix_watch`) would need to check each new / incoming brix value to see if fermentation has started (simple check for brix \> \~15).

`brix_watch` would then trigger the function `brix_handle` handling all further procedures during fermentation.

`brix_handle` would manipulate a fermentation dataset for each new brix value or a set window of brix values. It would require timestamp and raw brix values coming from the sensor as well as all newly created variables (brix smooth, derivatives, waypoints, phases).

**`brix_handle` input**: required data from fermentation start (sensor data: timestamp, brix, mutated data: fermentation boolean, smooth brix, derivatives, waypoints, phases)

**`brix_handle` procedure** (sub-routines running in `brix_handle)`:

1.  Calculate and set smooth brix and its derivatives (based on all brix values from fermentation start)
2.  Check for and set new waypoints and update fermentation phase
3.  Run defenses against incoming values to continue or abort fermentation phase
4.  Check for a fermentation end and return from this function to the higher order function checking for a new fermentation start.

**`brix_handle` output:**

1.  updated data from fermentation start
2.  fermentation boolean to control `brix_watch`

`brix_handle` would **terminate** once it detects the fermentation end or a false fermentation.

**Questions**:

-   Can we handle changes on the brix as described above? As in can we trigger a function (`brix_watch` here) on fermentation start once the brix value is past X as described above?
-   Ideally `brix_handle` would write to the same state it feeds from. In other words, we would need to pass the data to `brix_handle` it previously altered. So new data comes in, the function calculates for example the fermentation phase, which we would need for the next brix value. Would that work?
-   Note, that as a consequence to above point `brix_handle` would have the power to alter any previously set variables. For example 2 days into the fermentation it turns out it wasn't a fermentation at all. `brix_watch` would then be able to overwrite all changes to state and label the previous 2 days as non-fermentation.

## Smooth brix and derivatives

This should include:

-   Brix smooth
-   1st deriv / slope
-   1st deriv / Slope smooth
-   2nd deriv / acceleration

Further mechanics for later:

-   Add a variable uniquely labeling fermentation period and tank
-   Add a boolean ferment variable

### Helper functions

```{r}
#' Building out a brix and smooth brix chart.
chart_smooth_brix <- function(df) {
  
  p1 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix), color = "#cccccc") +
    geom_line(aes(y = brix_smooth), color = "hotpink", size = 1)
  print(p1)
  plotly::ggplotly(p1)
}

#' Building out a brix and smooth brix chart.
chart_smooth_slope <- function(df) {
  
  p1 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1), color = "#cccccc") +
    geom_line(aes(y = brix_deriv_1_smooth), color = "steelblue", size = 1)
  print(p1)
  plotly::ggplotly(p1)
}

#' Building out a time normalised slope chart
chart_norm_slope <- function(df) {
  
  p1 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_smooth), color = "steelblue", size = 1)

  p2 <- df %>% 
  ggplot(aes(x = time)) +
    geom_line(aes(y = brix_deriv_1_norm), color = "skyblue", size = 1)
  
  print(p1)
  print(p2)
  plotly::ggplotly(p1)
  plotly::ggplotly(p2)
}
```

We need to scale the span by the length of the data. Should be sth like:

\<= 250 -\> 2 // 500 -\> 1 // 1000 -\> 0.5 // \>= 2000 -\> 0.25

```{r}
#' Scale function with clamp mechanic.
#' 
#' @param value The value to scale
#' @param domain The tuple vector for the input domain
#' @param range The tuple vector for the output range
rescale_clamp <- function(value, domain, range) {
  if (value < min(domain)) return(range[1])
  if (value > max(domain)) return(range[2])
  out <- scales::rescale(value, from = domain, to = range)
  return(round(out,2))
}
```

### Interactive set up

```{r}

# Prep (not in function).
df_test <- df %>% 
  filter(time > lubridate::ymd_hms("2021-09-14 13:47:02")) %>% 
  filter(time < lubridate::ymd_hms("2021-09-20 00:00:00"))

# ------------------------------------------------------------------------------

# Calculate span.
length <- nrow(df_test)
span <- rescale_clamp(length, c(250, 2000), c(2, 0.25))

print(paste(length, ":", span))

# Get index for loess.
df_test <- df_test %>% rowid_to_column("index_")

# Add brix smooth.
result <- loess(brix ~ index_, data = df_test, span = span, na.action = na.exclude)
df_test$brix_smooth <- predict(result)

# Add slope.
df_test <- df_test %>% mutate(brix_deriv_1 = pracma::gradient(df_test$brix_smooth))

# Add slope smooth
result <- loess(brix_deriv_1 ~ index_, data = df_test, span = span, na.action = na.exclude)
df_test$brix_deriv_1_smooth <- predict(result)

# Add time normalised slope
df_test <- df_test %>% 
  mutate(time_diff = as.numeric(interval(lag(time, 1), time), units = "days")) %>% 
  mutate(brix_deriv_1_norm = (brix_smooth - lag(brix_smooth)) / time_diff)

# ------------------------------------------------------------------------------

# Chart it up.
chart_smooth_brix(df_test)
chart_smooth_slope(df_test)
chart_norm_slope(df_test)



```

### Smoother function

```{r}
#' Smoother function
#' 
#' @param data Input df expected to have `time` and `brix` cols (and no `index_` col)
#' @returns A df with augmented brix smooth, slope and acceleration cols
smooth_brix <- function(data) {
  # Base data.
  df <- data

  # Calculate span based on length. Span will be clamped betweeen 
  # 2 (for low # of rows) and 0.25 (for high # of rows) based on visual checks.
  length <- nrow(df)
  span <- rescale_clamp(length, c(250, 700), c(2, 0.25))
  
  # Debug span.
  print(paste(length, ":", span))
  
  # Get index for loess.
  df <- df %>% rowid_to_column("index_")
  
  # Add brix smooth.
  result <- loess(brix ~ index_, data = df, span = span, na.action = na.exclude)
  df$brix_smooth <- predict(result)
  
  # Add slope.
  df <- df %>% mutate(brix_deriv_1 = pracma::gradient(df$brix_smooth))
  
  # Add slope smooth
  result <- loess(brix_deriv_1 ~ index_, data = df, span = span, na.action = na.exclude)
  df$brix_deriv_1_smooth <- predict(result)
  
  # Add time normalised slope
  df <- df %>% 
    mutate(time_diff = as.numeric(interval(lag(time, 1), time), units = "days")) %>% 
    mutate(brix_deriv_1_norm = (brix_smooth - lag(brix_smooth)) / time_diff)

  # Clean up.
  df %>% select(-index_)

  return(df)
}
```

Test function

```{r}
test <- df %>% 
  filter(time > lubridate::ymd_hms("2021-09-14 13:47:02")) %>% 
  filter(time < lubridate::ymd_hms("2021-09-30 00:00:00")) %>%
  smooth_brix()

test
nrow(test)

chart_smooth_brix(test)

```

TODO: Clarify how ignition can handle these updates (see \[Description and questionsDescription and questions\] above)

Starting each tank at non-fermentation, one procedure would need to check each new / incoming brix value to see if it's fermentation start

From this point onwards a function needs to be called for incoming brix values (either for each new value or for a set window of new brix values).

This function would:

1.  Calculate smooth brix and its derivatives based on all brix values from fermentation start
2.  Check
