---
title: "Data v4 analysis - create profile data"
output: html_notebook
---

This requires the `df-clean.csv` created at `01 load raw data.Rmd`.

The transformation builds on the `explore.Rmd` file of v3.

```{r libs, echo=FALSE}
library(tidyverse)
library(ggplot2)
detach("package:dplyr")
library(dplyr)

# Project path
root_path <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
data_path <-  "/data/v4"
```


# Load

Read state as double otherwise it'll be read as logical.

```{r}

(df <- read_csv(paste0(root_path, data_path,  "/df-clean.csv"), col_types = cols(.default = "?")))

```

# Profile

Let's check each variable

```{r}

df %>% skimr::skim()

plot_col <- function(col) {
  column <- enquo(col)
  
  p <- df %>% 
    group_by(!!column) %>% 
    count() %>% 
    ggplot() +
      geom_bar(aes(x = !!column, y = n), stat = "identity")

  print(p)
}

plot_col(status)
plot_col(setpoint)
```

# Visualise

Let's look at the time series:

```{r vis-initial}

(p1 <- ggplot(df, aes(x = time)) +
  geom_line(aes(y = temp, colour = "temp")) +
  geom_line(aes(y = setpoint, colour = "setpoint")) +
  scale_colour_manual("", breaks = c("temp", "setpoint"), values = c("grey", "pink")) +
  labs(y = "temp and setpoint"))

(p2 <- ggplot(data = df, aes(x = time)) +
  geom_line(aes(y = status, colour = "lightskyblue")))

gridExtra::grid.arrange(p1, p2, nrow = 2)

plotly::ggplotly(p1)
plotly::ggplotly(p2)

```

Let's visualise them all in one visual

```{r}

# Get scaled variable's min/max.
df %>% select(status) %>% summary()

# Build the scales to show `status` in the same graph as the temps
# (to_scale_max - to_scale_min) * (value - value_min) / (value_max)
linear_scale <- function(x) {
  ((17-1)*(x-0)/(130-0)) + 1
}

invert_scale <- function(x) {
  ((130-0)*(x-1)/(17-1)) + 0
}

(p <- ggplot(df, aes(x = time)) +
  # Specificaly set `label` to also show the non-scaled status.
  geom_line(aes(y = invert_scale(status), colour = "status", label = status), size = 0.25) +
  geom_line(aes(y = setpoint, colour = "setpoint")) +
  geom_line(aes(y = temp, colour = "temp")) +
  scale_colour_manual("", breaks = c("temp", "setpoint", "status"), values = c("orange", "lightskyblue", "grey")) +
  scale_y_continuous(
    name = "Temperatures",
    sec.axis = sec_axis(trans = ~linear_scale(.), name = "Status")
  ) +
   scale_x_datetime(date_labels="%b %y", date_breaks  = "1 month"))

plotly::ggplotly(p)

```

Check individual months

```{r}
makeChart <- function(mth) {
p <- ggplot(df %>% filter(month == mth), aes(x = time)) +
  geom_line(aes(y = invert_scale(status), colour = "status", label = status), size = 0.25) +
  geom_line(aes(y = setpoint, colour = "setpoint")) +
  geom_line(aes(y = temp, colour = "temp")) +
  scale_colour_manual("", breaks = c("temp", "setpoint", "status"), values = c("orange", "lightskyblue", "grey")) +
  scale_y_continuous(
    name = "Temperatures",
    sec.axis = sec_axis(trans = ~linear_scale(.), name = "State")
  ) +
  labs(title = mth)
  
print(p)
plotly::ggplotly(p)

}

makeChart("Mar")
makeChart("Apr")
makeChart("May")
```

# Isolate pulse cooling sections

## Build and test algo

1) We create a diff time in seconds between rows (next minus current)
2) Check if the `status` values is pulsing between 2 and 3

This will result in most rows being correctly identified. However, it can be that events other than a status change happen during pulse cooling, like - obviously as that's hoped for - a temperature change. In this case, the status remains the same over a couple of events. To catch these two, we also:

3) Check if a `status` does not change BUT the time diff is small (e.g. < 180)

As this change happens on the full `pulse` vector, the mutate function is not aware of a change in `pulse` while doing it, so we need a couple of these mutates to cover multiple pulse cooling events at the same time. Hacky, but it works.

4) Check for each pulse that it's not too long (threshold: 600)

5) Identify start and end events for each pulse period

6) Add a unique identifier to each pulse period so we can group and analyse them individually

```{r}
df %>%
  # filter(time > lubridate::ymd("2021-10-06"), time < lubridate::ymd("2021-10-08")) %>%

  # 1) Add time diff variable to each row.
  mutate(diff = as.numeric(lead(time) - time)) %>% 
  
  # 2) Create `pulse` variable by checking if status is oscillating.
  mutate(pulse = ifelse(
    lag(status) != status & lead(status) != status
    & ((status == 3 | status == 2) & (lag(status) == 3 | lag(status) == 2) & (lead(status) == 3 | lead(status) == 2))
    , 'pulse', NA)) %>%

  # 3) Check if non-status events slipped in.
  mutate(pulse = ifelse(lag(pulse) == 'pulse' & is.na(pulse), ifelse(diff < 180, 'pulse', NA), pulse)) %>%
  mutate(pulse = ifelse(lag(pulse) == 'pulse' & is.na(pulse), ifelse(diff < 180, 'pulse', NA), pulse)) %>%
  mutate(pulse = ifelse(lag(pulse) == 'pulse' & is.na(pulse), ifelse(diff < 180, 'pulse', NA), pulse)) %>%
  mutate(pulse = ifelse(lag(pulse) == 'pulse' & is.na(pulse), ifelse(diff < 180, 'pulse', NA), pulse)) %>%

  # 4) Check if diff's are kept at bay
  mutate(pulse = ifelse(diff > 600, NA, pulse)) %>%
  
  # 5) Add start and end identifier to each pulse section.
  mutate(pulse = ifelse(is.na(lag(pulse)) & pulse == 'pulse' & lead(pulse) == 'pulse', 'start', pulse)) %>%
  mutate(pulse = ifelse(lag(pulse) == 'pulse' & pulse == 'pulse' & is.na(lead(pulse)), 'end', pulse)) %>%

  # 6) Add a unique identifier to each pulse period.
  mutate(pulse_group = ifelse(pulse == 'start', as.character(time), NA)) %>%
  fill(pulse_group) %>%
  mutate(pulse_group = ifelse(is.na(pulse), NA, pulse_group)) %>%

  # 7) Write for a manual chop...
  write_csv(paste0(root_path, data_path, "/pulse/pulse-check-01.csv"), na = "")

```

One thing this doesn't clean away is two short sections of the same `status`, like for example: 2,2 and then 3,3. There's surely a way to figure that out automatically (and in a next iteration we should) but for now, we save above, go through the `pulse`s and correct them manually (-> add an `issue` column and add an `"x"` for each issue row).

Here we read it in:

```{r}
df_man <- read_csv(paste0(root_path, data_path, "/pulse/pulse-manual-correction.csv"), col_types = cols(pulse = "c", pulse_group = "c", issue = "c"))
```

Some more cleaning and re-shuffling:

```{r}

(df_man <- df_man %>% 
  filter(is.na(issue)) %>% 
  select(-issue) %>% 
  select(time, temp, setpoint, status, diff, pulse, pulse_group, everything()))

```

Last thing to remove is very short pulses with only a single pulse:

```{r}

# Test
df_man %>% 
  mutate(rmv = ifelse(pulse == "start" & is.na(lead(pulse, n = 2)), TRUE, ifelse(pulse == "pulse" & is.na(lead(pulse)), TRUE, NA))) %>% 
  filter(is.na(rmv))

# Do
df_pulse <- df_man %>% 
  mutate(rmv = ifelse(pulse == "start" & is.na(lead(pulse, n = 2)), TRUE, ifelse(pulse == "pulse" & is.na(lead(pulse)), TRUE, NA))) %>% 
  filter(is.na(rmv)) %>% 
  select(-rmv)
  
```


Check the outcome:

```{r}
df_pulse %>% view()
```

# Save

```{r}
df_pulse %>% 
  write_csv(paste0(root_path, data_path, "/pulse/df-pulse.csv", na = ""))
```

