---
title: "Load data v2"
output: html_notebook
---

This file is loading in and prepping the new datasets - after we had multiple sources for `sp` and `state` in the v1 version.

```{r libs, include=FALSE}
library(tidyverse)
library(ggplot2)
detach("package:dplyr")
library(dplyr)
```

# Check a single month

Let's load June as an example dataset.

```{r load}

jun_1_orig <- readxl::read_xlsx("../../data/v2/D1/Jun.xlsx", sheet = "Dataset 1")
jun_2_orig <- readxl::read_xlsx("../../data/v2/D1/Jun.xlsx", sheet = "Dataset 2")

```

Let's do a base reshape (removing Brix, renaming cols and removing full NA's)

```{r base-reshape}

(jun_1_orig <- jun_1_orig %>% 
  select(c(1,2,3)) %>% 
  rename(time = 1, temp = 2, state = 3) %>% 
  filter(!is.na(temp) | !is.na(state)))

(jun_2_orig <- jun_2_orig %>% 
  rename(time = 1, sp = 2))



```

Let's look at it

```{r vis-initial, eval=FALSE}

jun_1_orig %>% 
  ggplot() +
    geom_line(aes(x = time, y = temp))

```

There's a funny datapoint at the end of each month (NOTE: NOT ANYMORE AS WE'VE UPDATED THE DATA. OLD DATA IN `D1 orig` FOLDER) and dataset (1 and 2) around the date-time `07/02/2022 - 08:15:00` (can be earlier can be later). 

I'll remove it for now.

```{r check, eval=FALSE}

jun_1 <- jun_1_orig %>% 
  head(., -1)

jun_2 <- jun_2_orig %>% 
  head(., -1)

```

Looking better:

```{r vis-initial-2, eval=FALSE}

jun_1 %>% 
  ggplot() +
    geom_line(aes(x = time, y = temp))

```

This might all make sense now...

```{r vis-plotly, eval=FALSE}

p1 <- ggplot() +
  geom_line(data = jun_1, aes(x = time, y = temp, colour = "temp")) +
  geom_step(data = jun_2, aes(x = time, y = sp, colour = "sp"), direction = "hv") + # note the step func changes at the next datapoint!
  scale_colour_manual("", breaks = c("temp", "sp"), values = c("grey", "pink")) +
  labs(y = "temp and setpoint")


p2 <- ggplot() +
  geom_line(data = jun_1, aes(x = time, y = state, colour = "state")) +
  scale_colour_manual("", breaks = c("state"), values = c("orange"))

gridExtra::grid.arrange(p1, p2, nrow = 2)

plotly::ggplotly(p1)
plotly::ggplotly(p2)

```

# All files

Manual reshape: I removed the estranged 2022-07-02 entry from all months/datasets.

Now, let's load all files:

```{r laod-all}

# Get file names > remove unwanted files > get file_paths
files <- list.files(path = '../../data/v2/D1')

files_clean <- tibble(file_name = files) %>% 
  filter(!str_starts(file_name, '~')) %>% # remove open books
  pull(file_name)

file_paths <- sapply(files_clean, function(file) { paste0('../../data/v2/D1/', file)  }, USE.NAMES = FALSE)

# Read in all files.
df_temp_list <- lapply(file_paths, function (x) { readxl::read_xlsx(x, sheet = "Dataset 1") })
df_sp_list <- lapply(file_paths, function (x) { readxl::read_xlsx(x, sheet = "Dataset 2") })

# Remove list entries with empty tibbles.
df_sp_list <- purrr::discard(df_sp_list, function(x) nrow(x) == 0)

# Join list items.
df_temp_orig <- bind_rows(df_temp_list) %>% as_tibble()
df_sp_orig <- bind_rows(df_sp_list) %>% as_tibble()

names(df_temp_orig)
names(df_sp_orig)

```

# Reshape and dupes

```{r reshape}

df_temp <- df_temp_orig %>% 
  select(c(1,2,3)) %>% 
  rename(time = 1, temp = 2, state = 3) %>% 
  filter(!is.na(temp) | !is.na(state))

df_sp <- df_sp_orig %>% 
  rename(time = 1, sp = 2)

names(df_temp)
names(df_sp)

```

```{r add-date-vars}

(df_temp <- df_temp %>% 
  mutate(
    wday = lubridate::wday(time, label = TRUE),
    day = lubridate::mday(time),
    month = lubridate::month(time, label = TRUE),
    year = lubridate::year(time)
  ))

(df_sp <- df_sp %>% 
  mutate(
    wday = lubridate::wday(time, label = TRUE),
    day = lubridate::mday(time),
    month = lubridate::month(time, label = TRUE),
    year = lubridate::year(time)
  ))

```


No time duplicates (which is all that counts):

```{r check_dupes, eval=FALSE}

df_temp %>% 
  group_by(time) %>% 
  count() %>% 
  arrange(desc(n))

df_sp %>% 
  group_by(time) %>% 
  count() %>% 
  arrange(desc(n))

```

# Check visuals

Let's look at all 

```{r vis, eval=FALSE}

p1 <- ggplot() +
  geom_line(data = df_temp, aes(x = time, y = temp, colour = "temp")) +
  geom_step(data = df_sp, aes(x = time, y = sp, colour = "sp"), direction = "hv") + # note the step func changes at the next datapoint!
  scale_colour_manual("", breaks = c("temp", "sp"), values = c("grey", "pink")) +
  labs(y = "temp and setpoint")


p2 <- ggplot() +
  geom_line(data = df_temp, aes(x = time, y = state, colour = "state")) +
  scale_colour_manual("", breaks = c("state"), values = c("orange"))

gridExtra::grid.arrange(p1, p2, nrow = 2)

plotly::ggplotly(p1)
plotly::ggplotly(p2)

```


plotly has issues with the step function


```{r vis-2, eval=FALSE}

ggplot() +
  geom_step(data = df_sp, aes(x = time, y = sp, colour = "sp"), direction = "hv") + # note the step func changes at the next datapoint!
  scale_colour_manual("", breaks = c("temp", "sp"), values = c("grey", "pink"))


```

# Merge

Let's merge the data. Maybe start with January

```{r merge-check, eval=FALSE}

df_temp_jan <- df_temp %>% 
  filter(month == "Jan")

df_sp_jan <- df_sp %>% 
  filter(month == "Jan")


bind_rows(df_temp_jan, df_sp_jan) %>% view()

bind_rows(df_temp_jan, df_sp_jan) %>% 
  arrange(desc(time)) %>% 
  view()

bind_rows(df_temp_jan, df_sp_jan) %>% 
  arrange(desc(time)) %>% 
  fill(temp) %>% 
  view()

bind_rows(df_temp_jan, df_sp_jan) %>% 
  arrange(desc(time)) %>% 
  fill(temp, state, sp) %>% 
  view()

```

That all looks good, let's merge them all and fill the variable gaps...

(Note, if the first date value is missing, all other values are missing up to the first variable value)

```{r merge}

(df <- bind_rows(df_temp, df_sp) %>% 
  select(time, temp, sp, state, everything()) %>% 
  arrange(time) %>% 
  fill(temp, state, sp))
```

# Visualise all data

Let's look at all data:

```{r vis-df}

(p1 <- ggplot(df, aes(x = time)) +
  geom_line(aes(y = temp, colour = "temp")) +
  geom_line(aes(y = sp, colour = "sp")) +
  scale_colour_manual("", breaks = c("temp", "sp"), values = c("grey", "pink")) +
  labs(y = "temp and setpoint"))

(p2 <- ggplot(data = df_temp, aes(x = time)) +
  geom_line(aes(y = state, colour = "state")) +
  scale_colour_manual("", breaks = c("state"), values = c("orange")))

gridExtra::grid.arrange(p1, p2, nrow = 2)

plotly::ggplotly(p1)
plotly::ggplotly(p2)

```

Seems to make all sense. A few pulse cooling sessions from October...

What next? Move to the `explore.Rmd` file 

# Write 

Writing file to load it in the Rmd to be sent to Caymus - see [here](https://stackoverflow.com/questions/64089594/my-rmd-file-keeps-getting-error-message-saying-that-you-are-passing-a-function#comment113343838_64089594)

```{r}

write_csv(df, '../../data/v2/output/D1-2021.csv', na = "")

```

