---
title: "Data source check June"
output: html_notebook
html_notebook:
    code_folding: hide
---

As part of an analysis of the association betwen temperature (`temp`), setpoint (`sp`) and state (`state`) data we've extracted the same data from 3 sources, which differed in setpoint as well as state data. 

We're now trying to understand which source is accurate.

This document highlights the differences between the datasets, solely looking at June 2021 data for tank D1.

## The three data sources

The variables  `temp`, `sp` and `state` are being produced from the three datasets

- **Extraction 1** (15 minute data - extracted from ignition before Jan 11)
- **Extraction 2** (1-2 hour data - extracted from ignition, via the interface on Jan 20)
- **Raw Signal** data from the database

We're only looking at the time between _Jun 8 to Jun 28_.

All data sources (plus the raw signal data for `sp` and `state` for the full year) can be found in the file `june-compare.xslx`, but here are the datasets in the order listed above for reference:

```{r}

(extr_01 <- readxl::read_xlsx("../data/june compare/june-compare.xlsx", sheet = "Extracted 1"))
(extr_02 <- readxl::read_xlsx("../data/june compare/june-compare.xlsx", sheet = "Extracted 2"))
(updt_raw <- readxl::read_xlsx("../data/june compare/june-compare.xlsx", sheet = "Raw signal"))

```


Note, that the extraction process produced interpolated figures for `sp` and `state`. In detail:
 
- **Extracted 1** includes interpolation for `sp` and `state`
- **Extracted 2** includes interpolation for `sp` only
- The **Raw Signal data** does not include any interpolation for neither `sp` nor `state`.

**The key question is which `sp` and `state` values are accurate?**

The following is a comparison of all values for the  June data:

## Temp check:

Result: all 3 datasets produce the same result. _Extracted 2_ looks different as it has a lower resolution (not quarter hourly data but 1-2 hour slots).


```{r}

# Visual
# extr_01 %>% ggplot() + geom_line(aes(x = t_stamp, y = eng))
# extr_02 %>% ggplot() + geom_line(aes(x = t_stamp, y = temp))
# updt_raw %>% ggplot() + geom_line(aes(x = t_stamp, y = temp))

# Data
check_tmp <- extr_01 %>% 
  select(t_stamp, temp_01 = temp) %>% 
  left_join(updt_raw %>% 
              select(t_stamp, temp_raw = temp), 
            by = "t_stamp")

# Check correlation
# cor(check_tmp %>% select(-1))

# Composed visual.
p1 <- check_tmp %>% ggplot(aes(x = t_stamp)) +
  geom_line(aes(y = temp_01, colour = "Extracted 1")) +
  geom_line(aes(y = temp_raw, colour = "Raw signal")) +
  scale_colour_manual("", breaks = c("Extracted 1", "Raw signal"), values = c("orange", "skyblue")) +
  labs(title = "Extracted 1 and Raw signal")

p2 <- extr_02 %>% ggplot() + 
  geom_line(aes(x = t_stamp, y = temp)) +
  labs(title = "Extracted 2")

plotly::ggplotly(p1, width = 800)
plotly::ggplotly(p2, width = 675)

```

## `sp` == setpoint data check:

Result: there's a discrepancy of 6.5 hours between the **Extracted 1** and the **Raw Signal** data.

```{r}

# Visual
# extr_01 %>% ggplot() + geom_line(aes(x = t_stamp, y = sp))
# extr_02 %>% ggplot() + geom_line(aes(x = t_stamp, y = sp))
# extr_02 %>% ggplot() + geom_line(aes(x = t_stamp, y = sp_extracted))
# updt_raw %>% ggplot() + geom_line(aes(x = t_stamp, y = sp_final))

# Data
(check_tmp <- extr_01 %>% 
  select(t_stamp, sp_01 = sp) %>% 
  left_join(updt_raw %>% 
              select(t_stamp, sp_raw = sp), 
            by = "t_stamp"))

# Check correlation
# cor(check_tmp %>% select(-1))

# Composed visual.
p1 <- check_tmp %>% ggplot(aes(x = t_stamp)) +
  geom_line(aes(y = sp_01, colour = "Extracted 1")) +
  geom_line(aes(y = sp_raw, colour = "Raw signal")) +
  scale_colour_manual("", breaks = c("Extracted 1", "Raw signal"), values = c("orange", "skyblue")) +
  labs(title = "Extracted 1 and Raw signal")

plotly::ggplotly(p1, width = 800)

p2 <- extr_02 %>% ggplot() + 
  geom_line(aes(x = t_stamp, y = sp)) +
  labs(title = "Extracted 2")

plotly::ggplotly(p2, width = 675)

```

## `state` check:

Result: 

- On Feb 17 there's a discrepancy of 7:15 hours between the **Extracted 1** and the **Raw Signal** data.
- On Feb 26 there's a discrepancy of 7:00 hours between the **Extracted 1** and the **Raw Signal** data.
- The **Extracted 2** data has an extra step from 1 to 2 on Feb 19
- The **Extracted 2** data's steps happen a few days after the other two datasets.

```{r}

# Visual
extr_01 %>% ggplot() + 
  geom_line(aes(x = t_stamp, y = state)) + 
  labs(title = "Extracted 1") + 
  scale_y_continuous(limits = c(1,3))

extr_02 %>% ggplot() + 
  geom_line(aes(x = t_stamp, y = state)) + 
  labs(title = "Extracted 2") + 
  scale_y_continuous(limits = c(1,3))

updt_raw %>% ggplot() + 
  geom_line(aes(x = t_stamp, y = state)) + 
  labs(title = "Raw signal") + 
  scale_y_continuous(limits = c(1,3))

# Data
(check_tmp <- extr_01 %>% 
  select(t_stamp, state_01 = state) %>% 
  left_join(updt_raw %>% 
              select(t_stamp, state_raw = state), 
            by = "t_stamp"))

# Check correlation
# cor(check_tmp %>% select(-1))

# Composed visual.
p1 <- check_tmp %>% ggplot(aes(x = t_stamp)) +
  geom_line(aes(y = state_01, colour = "Extracted 1")) +
  geom_line(aes(y = state_raw, colour = "Raw signal")) +
  scale_colour_manual("", breaks = c("Extracted 1", "Raw signal"), values = c("orange", "skyblue")) +
  labs(title = "Extracted 1 and Raw signal")

plotly::ggplotly(p1, width = 800)

p2 <- extr_02 %>% ggplot() + 
  geom_line(aes(x = t_stamp, y = state)) +
  labs(title = "Extracted 2")

plotly::ggplotly(p2, width = 675)


```




















