---
title: "Explore"
output: html_notebook
---

This script required the loaded and part-recoded frames from `load.Rmd` - either `df_d1` or `df_f2`

```{r eval=FALSE}
library(tidyverse)
library(ggplot2)
library(plotly)
detach("package:dplyr")
library(dplyr)
library(lubridate)

```

Load data (from previous script `recode.Rmd`)

```{r}

(df <- df_d1_recoded)
# (df <- df_f2_recoded)

```

```{r eval=FALSE}
df %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = temp, colour = "temp")) +
    geom_line(aes(x = t_stamp, y = sp_final, colour = "setpoint")) +
    scale_colour_manual("", breaks = c("temp", "setpoint"), values = c("grey", "pink")) +
    ylab("temp and setpoint") +
    facet_wrap(vars(month), scales = "free_x")

df %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state_final, colour = "status")) +
    scale_colour_manual("", breaks = c("status", "final"), values = c("orange", "skyblue")) +
    ylab("state") +
    facet_wrap(vars(month), scales = "free_x")

```

Check all months across all variables...

All state and sp variables:

```{r eval=FALSE}

f <- function(mth) {
  p1 <- df %>% 
    filter(month == mth) %>% 
    ggplot() +
      geom_line(aes(x = t_stamp, y = temp, colour = "tp")) +
      geom_line(aes(x = t_stamp, y = sp_extracted, colour = "spe"), linetype = "longdash") +
      geom_line(aes(x = t_stamp, y = sp_actual, colour = "spa"), linetype = "longdash") +
      geom_line(aes(x = t_stamp, y = sp_final, colour = "sp")) +
      scale_colour_manual("", breaks = c("tp", "sp", "spa", "spe"), values = c("grey", "pink", "brown", "skyblue")) +
      scale_y_continuous(limits = c(0,150)) +
      labs(y = "temp and setpoint", title = mth)
  
  p2 <- df %>% 
    filter(month == mth) %>% 
    ggplot() +
      geom_line(aes(x = t_stamp, y = state_extracted, colour = "ste"), linetype = "longdash") +
      geom_line(aes(x = t_stamp, y = state_actual, colour = "sta"), linetype = "longdash") +
      geom_line(aes(x = t_stamp, y = state_final, colour = "st")) +
      scale_colour_manual("", breaks = c("st", "sta", "ste"), values = c("orange", "brown", "skyblue")) +
      scale_y_continuous(breaks = c(1,2,3), limits = c(1,3)) +
      ylab("state")
  
  gridExtra::grid.arrange(p1, p2, nrow = 2)
}

for (mth in c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")) {
  f(mth)
}

```

Only the final state and sp variables

```{r eval=FALSE}

f <- function(mth) {
  p1 <- df %>% 
    filter(month == mth) %>% 
    ggplot() +
      geom_line(aes(x = t_stamp, y = temp, colour = "tp")) +
      # geom_line(aes(x = t_stamp, y = sp_extracted, colour = "spe"), linetype = "longdash") +
      # geom_line(aes(x = t_stamp, y = sp_actual, colour = "spa"), linetype = "longdash") +
      geom_line(aes(x = t_stamp, y = sp_final, colour = "sp")) +
      scale_colour_manual("", breaks = c("tp", "sp", "spa", "spe"), values = c("grey", "pink", "brown", "skyblue")) +
      scale_y_continuous(limits = c(0,150)) +
      labs(y = "temp and setpoint", title = mth)
  
  p2 <- df %>% 
    filter(month == mth) %>% 
    ggplot() +
      # geom_line(aes(x = t_stamp, y = state_extracted, colour = "ste"), linetype = "longdash") +
      # geom_line(aes(x = t_stamp, y = state_actual, colour = "sta"), linetype = "longdash") +
      geom_line(aes(x = t_stamp, y = state_final, colour = "st")) +
      scale_colour_manual("", breaks = c("st", "sta", "ste"), values = c("orange", "brown", "skyblue")) +
      scale_y_continuous(breaks = c(1,2,3), limits = c(1,3)) +
      ylab("state")
  
  gridExtra::grid.arrange(p1, p2, nrow = 2)
}

for (mth in c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")) {
  f(mth)
}

```


```{r}

p1 <- df %>% 
  filter(t_stamp > ymd("2021-02-08"), t_stamp < ymd("2021-02-13")) %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = temp, colour = "tp")) +
    geom_line(aes(x = t_stamp, y = sp_extracted, colour = "spe"), linetype = "longdash") +
    # geom_line(aes(x = t_stamp, y = sp_actual, colour = "spa"), linetype = "longdash") +
    geom_line(aes(x = t_stamp, y = sp_final, colour = "sp")) +
    scale_colour_manual("", breaks = c("tp", "sp", "spa", "spe"), values = c("grey", "pink", "brown", "skyblue")) +
    scale_y_continuous(limits = c(0,150)) +
    labs(y = "temp and setpoint", title = "Feb")

p2 <- df %>% 
  filter(t_stamp > ymd("2021-02-08"), t_stamp < ymd("2021-02-13")) %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state_extracted, colour = "ste"), linetype = "longdash") +
    # geom_line(aes(x = t_stamp, y = state_actual, colour = "sta"), linetype = "longdash") +
    geom_line(aes(x = t_stamp, y = state_final, colour = "st")) +
    scale_colour_manual("", breaks = c("st", "sta", "ste"), values = c("orange", "brown", "skyblue")) +
    scale_y_continuous(breaks = c(1,2,3), limits = c(1,3)) +
    ylab("state")

gridExtra::grid.arrange(p1, p2, nrow = 2)

```

# Data source check June

After a chat with Nic we tried to figure out 
- how to reverse engineer the extraction algo (basically how to turn the interpolations into step functions)
- what the story with this tank is

As an inventory of values, I'm comparing the variables `temp`, `sp` and `state` here for 4 datasets 

- Excel original
- Nic sent
- R extracted recoded
- R final

from the ~3 sources Ignition (Rick), Ignition (Nic), raw readings from db (Shawn).

Only looking at the signalbetween Jun 08 to Jun 25.

First,I extract the _recoded_ and _raw readings_ data:

```{r}
# Swap out state, sp variables you need
june_focussed <- df %>% 
  filter(t_stamp >= lubridate::ymd("2021-06-08"), 
         t_stamp < lubridate::ymd("2021-06-28")) %>% 
  select(t_stamp, temp, sp_extracted, state_extracted)

write_csv(june_focussed, "../data/june compare/r_extracted.csv")

june_focussed <- df %>% 
  filter(t_stamp >= lubridate::ymd("2021-06-08"), 
         t_stamp < lubridate::ymd("2021-06-28")) %>% 
  select(t_stamp, temp, sp_final, state_final)
  
write_csv(june_focussed, "../data/june compare/r_final.csv")
  
```

I compile it all in xls and get the 4 frames back in here:

```{r}

extr_org <- readxl::read_xlsx("../data/june compare/june-compare-original.xlsx", sheet = "Extracted original")
extr_nic <- readxl::read_xlsx("../data/june compare/june-compare-original.xlsx", sheet = "Extracted Nic")
extr_rec <- readxl::read_xlsx("../data/june compare/june-compare-original.xlsx", sheet = "Extracted recoded")
updt_raw <- readxl::read_xlsx("../data/june compare/june-compare-original.xlsx", sheet = "Updated with raw readings")

```

we have three data sources:

1. The originally extracted ignition data (Rick)
2. Raw orignal data from the database (Shawn)
3. Newly extracted ignition data just for June (Nic)

```{r}
extr_org
updt_raw
extr_nic
```

 I cleaned the originally extracted data but did not amend the other two datasets.

- The **originally extracted data** includes interpolation for `sp` and `state`
- The **Nic extracted data** includes interpolation for `sp` only
- The **raw signal data** does not include any interpolation.

The question is - which values to take?

The following is a comparison of all values for the  June data:

## Temp check:

Result: congruent

```{r}

# Visual
# extr_org %>% ggplot() + geom_line(aes(x = t_stamp, y = eng))
# extr_nic %>% ggplot() + geom_line(aes(x = t_stamp, y = eng))
# extr_rec %>% ggplot() + geom_line(aes(x = t_stamp, y = temp))
# updt_raw %>% ggplot() + geom_line(aes(x = t_stamp, y = temp))

# Data
(check_tmp <- extr_org %>% 
  select(t_stamp, temp_org = eng) %>% 
  left_join(extr_rec %>% 
              select(t_stamp, temp_rec = temp), 
            by = "t_stamp") %>% 
  left_join(updt_raw %>% 
              select(t_stamp, temp_raw = temp), 
            by = "t_stamp"))

# Check correlation
cor(check_tmp %>% select(-1))

# Composed visual.
p1 <- check_tmp %>% ggplot(aes(x = t_stamp)) +
  # geom_line(aes(y = temp_org, colour = "original")) +
  geom_line(aes(y = temp_rec, colour = "extracted")) +
  geom_line(aes(y = temp_raw, colour = "raw")) +
  scale_colour_manual("", breaks = c("original", "extracted", "raw"), values = c("grey", "orange", "skyblue"))

p2 <- extr_nic %>% ggplot() + geom_line(aes(x = t_stamp, y = eng))

plotly::ggplotly(p1)
plotly::ggplotly(p2)
```

## `sp` == setpoint data check:

Result: There's a discrepancy of 6.5 hours between the original ignition data and the raw signal data.

```{r}

# Visual
# extr_org %>% ggplot() + geom_line(aes(x = t_stamp, y = sp))
# extr_nic %>% ggplot() + geom_line(aes(x = t_stamp, y = sp))
# extr_rec %>% ggplot() + geom_line(aes(x = t_stamp, y = sp_extracted))
# updt_raw %>% ggplot() + geom_line(aes(x = t_stamp, y = sp_final))

# Data
(check_tmp <- extr_org %>% 
  select(t_stamp, sp_org = sp) %>% 
  left_join(extr_rec %>% 
              select(t_stamp, sp_rec = sp_extracted), 
            by = "t_stamp") %>% 
  left_join(updt_raw %>% 
              select(t_stamp, sp_raw = sp_final), 
            by = "t_stamp"))

# Check correlation
cor(check_tmp %>% select(-1))

# Composed visual.
p1 <- check_tmp %>% ggplot(aes(x = t_stamp)) +
  # geom_line(aes(y = sp_org, colour = "original")) +
  geom_line(aes(y = sp_rec, colour = "extracted")) +
  geom_line(aes(y = sp_raw, colour = "raw")) +
  scale_colour_manual("", breaks = c("original", "extracted", "raw"), values = c("grey", "orange", "skyblue"))

plotly::ggplotly(p1)

p2 <- extr_nic %>% ggplot() + geom_line(aes(x = t_stamp, y = sp))

plotly::ggplotly(p2)

```

## `state` check:

Result: 
- On Feb 17 there's a discrepancy of 7:15 hours between the original ignition data and the raw signal data.
- On Feb 26 there's a discrepancy of 7:00 hours between the original ignition data and the raw signal data.
- Nic's data has an extra step from 1 to 2 on Feb 19
- Nic's data steps from 2 to 3 on Feb 24 - the originally extracted and the raw signal data steps up on Feb 26 

```{r}

# Visual
# extr_org %>% ggplot() + geom_line(aes(x = t_stamp, y = state)) + labs(title = "Original extraction") + scale_y_continuous(limits = c(1,3))
extr_rec %>% ggplot() + geom_line(aes(x = t_stamp, y = state_extracted)) + labs(title = "Recoded extraction") + scale_y_continuous(limits = c(1,3))
extr_nic %>% ggplot() + geom_line(aes(x = t_stamp, y = state)) + labs(title = "Nic extraction") + scale_y_continuous(limits = c(1,3))
updt_raw %>% ggplot() + geom_line(aes(x = t_stamp, y = state_final)) + labs(title = "Raw signal") + scale_y_continuous(limits = c(1,3))

# Data
(check_tmp <- extr_org %>% 
  select(t_stamp, state_org = state) %>% 
  left_join(extr_rec %>% 
              select(t_stamp, state_rec = state_extracted), 
            by = "t_stamp") %>% 
  left_join(updt_raw %>% 
              select(t_stamp, state_raw = state_final), 
            by = "t_stamp"))

# Check correlation
cor(check_tmp %>% select(-1))

# Composed visual.
p1 <- check_tmp %>% ggplot(aes(x = t_stamp)) +
  # geom_line(aes(y = state_org, colour = "original")) +
  geom_line(aes(y = state_rec, colour = "extracted")) +
  geom_line(aes(y = state_raw, colour = "raw")) +
  scale_colour_manual("", breaks = c("original", "extracted", "raw"), values = c("grey", "orange", "skyblue"))

plotly::ggplotly(p1)

p2 <- extr_nic %>% ggplot() + geom_line(aes(x = t_stamp, y = state))

plotly::ggplotly(p2)

```