---
title: "Special D1 months data join issue"
output: html_notebook
---

# D1 tank data fixes

Problems found in the data were explored here and manuallly changed for the raw xls data (the files that have all variables in `data/raw/D1`). The main problem was:

"one specific question that came up was for the D1 August data, which shows April dates instead. I assume August just got labelled wrong, but we’re missing the 31st"

Rick from Caymus then send over the unjoined data (one file per months) which I checked here.

## November

I mistakenly did November first, which turned out to be good as it had some duplicates to be cleaned out:

```{r}

st <- readxl::read_xlsx("../data/raw/D1 unjoined/NovState.xlsx")
sp <- readxl::read_xlsx("../data/raw/D1 unjoined/NovSPPV.xlsx")
p <- readxl::read_xlsx("../data/raw/D1 unjoined/NovPulse.xlsx")
b <- readxl::read_xlsx("../data/raw/D1 unjoined/NovBRIX.xlsx")

nov <- st %>% 
  left_join(sp, by = "t_stamp") %>% 
  left_join(p, by = "t_stamp") %>% 
  left_join(b, by = "t_stamp")

skimr::skim(nov)

```


Add some new dat var's

```{r}

nov <- nov %>% 
  mutate(
    year = lubridate::year(t_stamp),
    month = lubridate::month(t_stamp, label = TRUE),
    day = lubridate::wday(t_stamp, label = FALSE),
    wday = lubridate::wday(t_stamp, label = TRUE)
  )

nov %>% 
  filter(is.na(sp))

```


The missings are all December values. Let's remove them..

```{r}
nov <- nov %>% 
  filter(month == 'Nov')

View(nov)
```

Let's check for duplicates:

```{r}

nov %>% 
  select(t_stamp, state, sp, eng, pulsing, brix) %>% 
  distinct()

nov %>% 
  # group_by(t_stamp, state, sp, eng, pulsing, brix) %>% 
  group_by(t_stamp) %>%
  count() %>% 
  filter(n > 1)

readr::write_csv(nov, "nov.csv") # to check in excel

```

I checked in excel and it turns out that for each of the above quarter hours (there 2021-11-0 1am-1:45am) we have 8 rows, which differ in `eng` and `brix`. As all values are equally close to that quarter hour's average of the respective variable, 

I will take a random row:

```{r}

nov <- nov %>% 
  distinct(t_stamp, .keep_all = T)

readr::write_csv(nov, "nov_clean.csv") # to feed back into the main load/explore script...

```

## April

Check April (which previously had all August cases)

```{r}

st <- readxl::read_xlsx("../data/raw/D1 unjoined/AprState.xlsx")
sp <- readxl::read_xlsx("../data/raw/D1 unjoined/AprSPPV.xlsx")
b <- readxl::read_xlsx("../data/raw/D1 unjoined/AprBRIX.xlsx")

apr <- st %>% 
  left_join(sp, by = "t_stamp") %>% 
  left_join(b, by = "t_stamp")

skimr::skim(apr)

```


April data looking good. Export for the main load/explore:

```{r}
apr %>% readr::write_csv("apr_final.csv")
```


## August

Check August (which previously was not available)

```{r}

st <- readxl::read_xlsx("../data/raw/D1 unjoined/AugState.xlsx")
sp <- readxl::read_xlsx("../data/raw/D1 unjoined/AugSPPV.xlsx")
b <- readxl::read_xlsx("../data/raw/D1 unjoined/AugBRIX.xlsx")

aug <- st %>% 
  left_join(sp, by = "t_stamp") %>% 
  left_join(b, by = "t_stamp")

skimr::skim(aug)

```


Also looking good. Export for the main load/explore:

```{r}
aug %>% readr::write_csv("aug_final.csv")
```


## March

March had 4 rows too few (requires the main file to be in a `df` object).

```{r}
# Check manually.
df %>% 
  filter(month == "Mar") %>% 
  write_csv("mar.csv")

# Add a time difference var.
(diff_check <- df %>% 
  filter(month == "Mar") %>% 
  mutate(diff = as.numeric(t_stamp - lag(t_stamp - 1))))
  
# Check if there are longer differences between rows?
diff_check %>% 
  group_by(diff) %>% 
  count()

# Look at the case.
diff_check %>% 
  filter(diff > 16)
  
# Look at the case in its environment.
diff_check %>% 
  filter(t_stamp > lubridate::ymd_h("2021-03-14 1"),
         t_stamp < lubridate::ymd_h("2021-03-14 4"))
```

On 2021-03-14 there's an hour missing: 02:00, 02:15, 02:30 and 02:45.

These are also missing in the ra ("uncompiled") files from Rick/Caymus. So I've filled these gaps in **the base files loaded in** with an average of the previous and the next value of the gap.

## April and June

...had some duplicate rows for the first quarter hour from the previous months. Manullay removed from the D1 xls files

# F2 tank data fixes

## March

The Februray dataset had the first 288 March quarter hours with NA state data. Removed in the Feb xls data

Also, on 14-03 we had the 2am quarter hours missing which I discovered like this:

(`df` needs the be the prepped F1 frame)

```{r}
lag_check <- df %>% 
  filter(month == "Mar") %>% 
  mutate(diff = t_stamp - lag(t_stamp - 1)) 

lag_check %>% 
  group_by(diff) %>% 
  count()

lag_check %>% 
  filter(diff > 16)

lag_check %>% 
  filter(t_stamp > lubridate::ymd_h("2021-03-14 1"),
         t_stamp < lubridate::ymd_h("2021-03-14 4"))
```

And changed manually to reflect the average of the 1:45 and the 3:00 o'clock quarter hour.