---
title: "Caymus"
output: html_notebook
---

```{r eval=FALSE}
install.packages('tidyverse')
install.packages('visdat')
install.packages('skimr')
install.packages("gridExtra")
```

```{r}
library(tidyverse)
library(ggplot2)
detach("package:dplyr")
library(dplyr)
```


# D1 tank

Let's get the data in and join it for D1 first

```{r}
# Test
# t <- readxl::read_xlsx('../data/raw/D1/D1Jan2021.xlsx')

# Get file names > remove unwanted files > get file_paths
files <- list.files(path = '../data/raw/D1')

files_clean <- tibble(file_name = files) %>% 
  filter(!str_starts(file_name, '~')) %>% # remove open books
  filter(str_detect(file_name, '2021.xlsx')) %>% # keep correctly named files
  pull(file_name)

file_paths <- sapply(files_clean, function(file) { paste0('../data/raw/D1/', file)  }, USE.NAMES = FALSE)


# Read in all files > join them
df_list <- lapply(file_paths, readxl::read_xlsx)
df_orig <- bind_rows(df_list) %>% as_tibble()
names(df_orig)
```

Some recoding

```{r}
# Remove brix (not needed).
df_orig <- df_orig %>% select(-brix)

# Add date vars.
(df_orig <- df_orig %>% 
  mutate(
    year = lubridate::year(t_stamp),
    month = lubridate::month(t_stamp, label = TRUE),
    day = lubridate::mday(t_stamp),
    wday = lubridate::wday(t_stamp, label = TRUE)
  ))
```

Summary

```{r eval=FALSE}
skimr::skim(df_orig)
```

What is `pulsing`?

```{r eval=FALSE}
df_orig %>% 
  mutate(pulse_bool = !is.na(pulsing)) %>% 
  group_by(pulse_bool) %>% 
  count()

df_orig %>% 
  filter(!is.na(pulsing)) %>% 
  group_by(month, wday) %>% 
  count()

df_orig %>% 
  filter(!is.na(pulsing)) %>% 
  group_by(pulsing) %>% 
  count()
```

Got added in Nov - is 0.

Remove duplicates

```{r}
# Duplicates
(dupes <- df_orig %>% 
  group_by(t_stamp, year, month, day, wday, sp, eng, state) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)))

df <- dupes %>% 
  filter(n == 1) %>% 
  select(-n)

df_orig %>% 
  group_by(t_stamp) %>% 
  count() %>% 
  arrange(desc(n), desc(t_stamp))
```

Remove missing timestamps

```{r}
# Missings check > filter out
(df <- df %>% 
  filter(!is.na(t_stamp)))

df %>% summary()
```

Check each months has enough rows/quarter hours

```{r}

# Month distribution
df %>% 
  group_by(month) %>% 
  count() 

# Check if we have the correct number of rows (quarter hours) per month
days_per_mth <- tibble(
  month = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
  days = c(31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31)
)

days_per_mth <- days_per_mth %>% 
  mutate(quarter_hrs = days * 24 * 4)

df_check_mths <- df %>% 
  group_by(month) %>% 
  count() %>% 
  ungroup() %>% 
  left_join(days_per_mth, by = "month") %>% 
  mutate(diff = n - quarter_hrs)
  
df_check_mths

```

There were some irregularities, duplicates and holes which were fixed in the script `fix_months.Rmd`

Let's have a look

```{r eval=FALSE}
df %>% 
  ggplot(aes(x = t_stamp)) + 
    geom_line(aes(y = eng), colour = "steelblue") +
    geom_line(aes(y = sp), colour = "orange") +
    # geom_line(aes(y = brix), colour = "darkseagreen") +
    # geom_line(aes(y = state), colour = "red") +
    facet_wrap(vars(month), scales = "free")

```

```{r eval=FALSE}
visdat::vis_dat(df)
```

Name the frame for use in other scripts

```{r}
df_d1_prepped <- df
```


---

# F2 tank

Let's look at the F1 tank next

```{r}
# Get file names > get file_paths
files <- list.files(path = '../data/raw/F2')
file_paths <- sapply(files, function(file) { paste0('../data/raw/F2/', file)  }, USE.NAMES = FALSE)

# Read in all files > join them
df_list <- lapply(file_paths, readxl::read_xlsx)
df_orig <- bind_rows(df_list) %>% as_tibble()
names(df_orig)
```

Some recoding

```{r}
# Add date vars.
df_orig <- df_orig %>% 
  mutate(
    year = lubridate::year(t_stamp),
    month = lubridate::month(t_stamp, label = TRUE),
    day = lubridate::mday(t_stamp),
    wday = lubridate::wday(t_stamp, label = TRUE)
  )
```

Summary

```{r}
skimr::skim(df_orig)
```

No `brix`?

What is `pulsing`?

```{r}
df_orig %>% 
  mutate(pulse_bool = !is.na(pulsing)) %>% 
  group_by(pulse_bool) %>% 
  count()

df_orig %>% 
  filter(!is.na(pulsing)) %>% 
  group_by(month, day) %>% 
  count()

df_orig %>% 
  filter(!is.na(pulsing)) %>% 
  group_by(pulsing) %>% 
  count()
```

Let's check for duplicates (across all var's and the `t_stamp` in particular)

```{r}
# Duplicates
(dupes <- df_orig %>% 
  group_by(t_stamp, year, month, day, sp, eng, state) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)))

df <- dupes %>% 
  filter(n == 1) %>% 
  select(-n)

df_orig %>% 
  group_by(t_stamp) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  filter(n > 1)
```

...and the missings:

```{r}
# Missings check > filter out
df %>% 
  filter(is.na(t_stamp))

(df <- df %>% 
  filter(!is.na(t_stamp)))
```

Now, let's see if there are the correct number of rows per month

```{r}
# Month distribution
df %>% 
  group_by(month) %>% 
  count() 

# Check if we have the correct number of rows (quarter hours) per month
days_per_mth <- tibble(
  month = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
  days = c(31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31)
)

days_per_mth <- days_per_mth %>% 
  mutate(quarter_hrs = days * 24 * 4)

df_check_mths <- df %>% 
  group_by(month) %>% 
  count() %>% 
  ungroup() %>% 
  left_join(days_per_mth, by = "month") %>% 
  mutate(diff = n - quarter_hrs)
  
df_check_mths

```

Let's also check if all the quarter hours are actually a quarter hour apart 

```{r}

# Compare each row's t_stamp with the previous row's t_stamp value.
lag_check <- df %>% 
  mutate(diff = t_stamp - lag(t_stamp - 1)) 

lag_check %>% 
  group_by(diff) %>% 
  count()

```


There were some issues, duplicates and whole which have been explored in the fix_months.Rmd script and repaired manually in the F1 excels.

Let's have a look

```{r}

df %>% 
  ggplot(aes(x = t_stamp)) + 
    geom_line(aes(y = eng), colour = "steelblue") +
    geom_line(aes(y = sp), colour = "orange") +
    # geom_line(aes(y = state), colour = "red") +
    facet_wrap(vars(month), scales = "free")

```

```{r}
visdat::vis_dat(df)
```

