---
title: "Explore"
output: html_notebook
---

This script required the loaded and part-recoded frames from `load.Rmd` - either `df_d1` or `df_f2`

```{r eval=false}
library(tidyverse)
library(ggplot2)
detach("package:dplyr")
library(dplyr)
library(lubridate)
```

Set the data (either from tank d1 or d2)

```{r}
df <- df_d1_prepped
# df <- df_f2_prepped
```

# Discretising `state` and `sp`

Have a closer look at `state`

```{r eval=FALSE}
df %>% 
  group_by(state) %>% 
  count() %>% 
  arrange(desc(n))

df %>% 
  select(state) %>% 
  summary()
```

Lets check if this makes sense:

```{r eval=FALSE}

# Facet plot status and floored status
df %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state), colour = "orange") +
    facet_wrap(vars(month), scales = "free_x")

# Facet plot status and floored status
df %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = eng, colour = "temp")) +
    geom_line(aes(x = t_stamp, y = sp, colour = "sp")) +
    scale_colour_manual("", breaks = c("temp", "sp"), values = c("grey", "pink")) +
    facet_wrap(vars(month), scales = "free")

```

Some issues here...

The key issue is that the extraction algo extracted the discrete `state` and `sp` variables as part continuous as it iinterpolated between values. In fact `sp` is continuous but it's series should describe a step function.  An example is June:

```{r eval=FALSE}

month_selected <- "Jun"

plot1 <- df %>% 
  filter(month == month_selected) %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state, colour = "state")) +
    scale_colour_manual("", breaks = c("ceil", "state"), values = c("skyblue", "orange")) +
    scale_y_continuous(breaks = c(1,2,3)) +
    labs(y = "")

plot2 <- df %>% 
  filter(month == month_selected) %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = eng, colour = "temp")) +
    geom_line(aes(x = t_stamp, y = sp, colour = "sp")) +
    scale_colour_manual("", breaks = c("temp", "sp"), values = c("grey", "pink")) +
    labs(y = "")

gridExtra::grid.arrange(plot1, plot2, nrow = 2)

```

Question: at what points was the state 1, 2 or 3? And at what times was the setpoint (`sp`) which value?

Answer: the answer is in the file `State and SP readings from raw data.xlsx` which is an extract of all `state` and `sp` readings **when they changed** from the raw data.

Discussion: we should probably read these in and join them as a new variable. Alternatively, we could reverse-engineer the extraction algorithm, but it won't be a simple floor or ceiling exercise of the variables. We will need to distinguish previous slopes from orthogonal data changes (as in the June `sp` example) and some values don't go all the way up to integers like for example on the 29th of April at 18:30, which goes up to 2.999351:

```{r eval=FALSE}
apr <- df %>% 
  filter(t_stamp > lubridate::ymd("2021-04-29"), t_stamp < lubridate::ymd("2021-04-30"))

apr %>% View()

apr %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state), colour = "skyblue")
```

Solution: in  the file `State and SP readings from raw data.xlsx` we prepare two sheets to join into the base file here. One sheet holds all discrete `state` and the other all `sp` values from the raw data matched to the _next_ quarter hour (the values are by minute, but I assume the correct way of logging would be by the next quarter hour). Time duplicates in one quarter hour are indicated (the latest value is being taken):

```{r}
df <- df_d1_prepped

# Read in actual readings.
state_actual <- readxl::read_excel("../data/raw/D1/State and SP readings from raw data.xlsx", sheet = "state")
sp_actual <- readxl::read_excel("../data/raw/D1/State and SP readings from raw data.xlsx", sheet = "sp")

state_actual
sp_actual

# Remove duplicates (keep the latest date/value - prepped in xls)
state_actual <- state_actual %>% filter(is.na(remove_dupe)) %>% select(-remove_dupe)
sp_actual <- sp_actual %>% filter(is.na(remove_dupe)) %>% select(-remove_dupe)

# Prep df (remove all actual values that might have been added previously)
df <- df %>% 
  select(-starts_with("state_actual")) %>% 
  select(-starts_with("sp_actual"))

# Join.
(df <- df %>% 
  left_join(state_actual, by = "t_stamp") %>% 
  left_join(sp_actual, by = "t_stamp"))
  
```

All the non-set timepoints need to be filled, which happens in the following from **the value down**. The first value however is missing, which has been assumed by viewing the data:

```{r}

# Set first `state` value for `fill` to work (check with chart above)
df[1,]$state_actual = 2

# Set first `sp` value for `fill` to work (check with chart above and set the most likely value)
df[1,]$sp_actual = 55

# Fill the missing values (from top to bottom)
df <- df %>% 
  fill(state_actual, .direction = "down") %>% 
  fill(sp_actual, .direction = "down")

```

Check the solution:

```{r eval=FALSE}

df %>% 
  group_by(state_actual) %>% 
  count()

df %>% 
  group_by(sp_actual) %>% 
  count()

```

```{r eval=FALSE}

# Facet plot status and floored status
df %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = state_actual, colour = "actual")) +
    scale_colour_manual("", breaks = c("extracted", "actual"), values = c("orange", "skyblue")) +
    ylab("state") +
    facet_wrap(vars(month), scales = "free_x")

# Facet plot status and floored status
df %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = eng, colour = "temp")) +
    geom_line(aes(x = t_stamp, y = sp, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = sp_actual, colour = "actual")) +
    scale_colour_manual("", breaks = c("temp", "extracted", "actual"), values = c("grey", "pink", "skyblue")) +
    ylab("set point") +
    facet_wrap(vars(month), scales = "free_x")

```

Check period with missing actual `state` and `sp` data

There's no `state` and `sp` data from 19/10/21. `state` values from this day can go to 0.5 and there's still one `sp` value sloping up.

```{r eval=FALSE}
check <- df %>% 
  filter(t_stamp > ymd("2021-10-19"))

check %>% View()
```


## Checking `sp`:

```{r eval=FALSE}

# The whole period
check %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = sp, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = sp_actual, colour = "actual")) +
    scale_colour_manual("", breaks = c("extracted", "actual"), values = c("pink", "skyblue")) +
    ylab("setpoint") +
    facet_wrap(vars(month), scales = "free_x")

# Check which numbers are in-between possible values
check %>% 
  group_by(sp) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  View()
  
# Result: slope between 19.10 and 

# The October fun
check %>% 
  filter(t_stamp > ymd_h("2021-10-01 0"), t_stamp < ymd_h("2021-10-31 0")) %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = sp, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = sp_actual, colour = "actual")) +
    scale_colour_manual("", breaks = c("extracted", "actual"), values = c("pink", "skyblue")) +
    ylab("setpoint") +
    facet_wrap(vars(month), scales = "free_x")
	
```

The slope from 19.10 moves to **2021-10-26 18:00:00** exactly. Assuming the extraction logic to be responsible for the slope, we'll take `sp_actual` up to Oct 26 and then switch to `sp`. 

```{r eval=FALSE}

# Check all.
df %>% 
  mutate(sp_final = ifelse(t_stamp > ymd_hms("2021-10-26 18:00:00"), sp, sp_actual)) %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = sp, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = sp_actual, colour = "actual")) +
    geom_line(aes(x = t_stamp, y = sp_final, colour = "final")) +
    scale_colour_manual("", breaks = c("extracted", "actual", "final"), values = c("pink", "skyblue", "grey")) +
    ylab("setpoint") +
    facet_wrap(vars(month), scales = "free_x")

# Double check key months.
p1 <- df %>% 
  mutate(sp_final = ifelse(t_stamp > ymd_hms("2021-10-26 18:00:00"), sp, sp_actual)) %>% 
  filter(month == "Oct" | month == "Nov" | month == "Dec") %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = sp, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = sp_actual, colour = "actual")) +
    geom_line(aes(x = t_stamp, y = sp_final, colour = "final")) +
    scale_colour_manual("", breaks = c("extracted", "actual", "final"), values = c("pink", "skyblue", "grey")) +
    ylab("setpoint") +
    facet_wrap(vars(month), scales = "free_x")

p2 <- df %>% 
  mutate(sp_final = ifelse(t_stamp > ymd_hms("2021-10-26 18:00:00"), sp, sp_actual)) %>% 
  filter(month == "Oct" | month == "Nov" | month == "Dec") %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = sp, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = sp_actual, colour = "actual")) +
    scale_colour_manual("", breaks = c("extracted", "actual", "final"), values = c("pink", "skyblue", "grey")) +
    ylab("setpoint") +
    facet_wrap(vars(month), scales = "free_x")

gridExtra::grid.arrange(p1, p2, nrow = 2)

```

Yep, do it:

```{r}

(df <- df %>% 
  mutate(sp_final = ifelse(t_stamp > ymd_hms("2021-10-26 18:00:00"), sp, sp_actual)))

```




## Checking `state`:

```{r eval=FALSE}

# The whole period
check %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = state_actual, colour = "actual")) +
    scale_colour_manual("", breaks = c("extracted", "actual"), values = c("orange", "skyblue")) +
    ylab("state") +
    facet_wrap(vars(month), scales = "free_x")

# Check which numbers are in-between possible values
check %>% 
  group_by(state) %>% 
  count() %>% 
  arrange(desc(n)) 

# Result: no values between 1 and 2 but many values between 2 and 3

# The October fun
check %>% 
  filter(t_stamp > ymd_h("2021-10-23 18"), t_stamp < ymd_h("2021-10-26 18")) %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = state_actual, colour = "actual")) +
    scale_colour_manual("", breaks = c("extracted", "actual"), values = c("orange", "skyblue")) +
    ylab("state") +
    facet_wrap(vars(month), scales = "free_x")
	
```

Assuming 2.5 is not a possible `state` value, we can probably decree that 

after 19/10/21 anything higher than 2 is 3 (there are no figures between 1 and 2)

```{r eval=FALSE}

test <- df %>% 
  mutate(state_blurb = ifelse(t_stamp < ymd("2021-10-19"), state_actual, ceiling(state)))

test %>% 
  filter(t_stamp > ymd_h("2021-10-23 18"), t_stamp < ymd_h("2021-10-26 18")) %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = state_actual, colour = "actual")) +
    geom_line(aes(x = t_stamp, y = state_blurb, colour = "final")) +
    scale_colour_manual("", breaks = c("extracted", "actual", "final"), values = c("orange", "skyblue", "green"))


df %>% 
  mutate(state_actual = ifelse(t_stamp < ymd("2021-10-19"), state_actual, ceiling(state))) %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = state_actual, colour = "actual")) +
    scale_colour_manual("", breaks = c("extracted", "actual"), values = c("orange", "skyblue")) +
    facet_wrap(~month, scales = "free_x")

# All apart from mid October seems fine. Let's focus in:

oct_focus <- df %>% 
  mutate(state_actual = ifelse(t_stamp < ymd("2021-10-19"), state_actual, ceiling(state))) %>% 
  filter(t_stamp > ymd_h("2021-10-16 0"), t_stamp < ymd_h("2021-10-20 0"))

oct_focus %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = state_actual, colour = "actual")) +
    scale_colour_manual("", breaks = c("extracted", "actual"), values = c("orange", "skyblue")) +
    facet_wrap(~month, scales = "free_x")

oct_focus %>% 
  group_by(state) %>% 
  count()

oct_focus %>% View()

```

Wrt `state` it seems our ceiling function does the trick for all but the mid October values.

As the last actual `state` reading from the raw data file is on **2021-09-25 07:15:00**, I will apply the ceiling function to any extracted `state` value after this date.

Here's the final state recode:

```{r}

df <- df %>% 
  mutate(state_final = ifelse(t_stamp > ymd_hms("2021-09-25 07:15:00"), ceiling(state), state_actual))
 
df %>% 
  filter(t_stamp > ymd_h("2021-09-25 0"), t_stamp < ymd_h("2021-12-31 0")) %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state, colour = "extracted")) +
    geom_line(aes(x = t_stamp, y = state_actual, colour = "actual")) +
    geom_line(aes(x = t_stamp, y = state_final, colour = "final")) +
    scale_colour_manual("", breaks = c("extracted", "actual", "final"), values = c("orange", "skyblue", "grey"))

```


## Check congruency between extracted and actual state 

```{r}

feb_ex <- df %>% 
  filter(t_stamp > ymd("2021-02-25"), t_stamp < ymd("2021-02-27"))

# feb_ex %>% View()

p1 <- feb_ex %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = state, colour = "extracted")) +
    # geom_line(aes(x = t_stamp, y = state_actual, colour = "actual")) +
    geom_line(aes(x = t_stamp, y = state_final, colour = "final")) +
    scale_colour_manual("", breaks = c("extracted", "final"), values = c("orange", "skyblue")) +
    ylab("state") +
    facet_wrap(vars(month), scales = "free_x")

p2 <- feb_ex %>% 
  ggplot() +
    geom_line(aes(x = t_stamp, y = eng, colour = "temp")) +
    # geom_line(aes(x = t_stamp, y = sp_actual, colour = "sp" actual)) +
    geom_line(aes(x = t_stamp, y = sp_final, colour = "sp final")) +
    scale_colour_manual("", breaks = c("temp", "sp final"), values = c("grey", "pink")) +
    ylab("temp and sp") +
    facet_wrap(vars(month), scales = "free_x")

gridExtra::grid.arrange(p1, p2, nrow = 2)

  
```

There's a difference between extracted and final values in terms of when a state value got changed. Above Februrary example shows the extracted `state` to increase to 3 on Feb 25 at around 16:15pm and the final `state` to increase on Feb 26 at 0:30am. 

I'm assuming the final `state` is correct, supported by the setpoint correlation.

Rename frame and columns and remove unnecessary cols for next script...




```{r}

df_d1_recoded <- df %>% 
  select(
    t_stamp,
    year,
    month,
    day,
    wday,
    temp = eng,
    state_extracted = state,
    state_actual,
    state_final,
    sp_extracted = sp,
    sp_actual,
    sp_final
  )

```





