---
title: "Data v3 Analysis"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plotly)
detach("package:dplyr")
library(dplyr)

root_path <- "/Users/lars/Dropbox/lars/stats/projects/ZAM/caymus"
```


# Functions

```{r}

graph_central_values <- function(df, var, var_name, rounding = 0, y = "Values") {
  var_enquo = enquo(var)
  
  # Get the data range for normalised nudge_* values
  vector <- df %>% pull(!!var_enquo)
  dist <- range(vector)[2] - range(vector)[1]
  nudge <- dist / 25
  
  # Get the data
  data <- df %>% 
    group_by(ferment_period) %>% 
    summarise(
      Average = mean(!!var_enquo),
      Median = median(!!var_enquo),
      Min = min(!!var_enquo),
      Max = max(!!var_enquo),
    )
  
  # Plot
  p <- ggplot(data) +
    geom_pointrange(aes(x = ferment_period, y = Median, ymin = Min, ymax = Max, colour = ferment_period), shape = 3) +
    geom_point(aes(x = ferment_period, y = Average, colour = ferment_period), size = 3) +
    geom_text(aes(x = ferment_period, y = Median, label = paste("Median: ", round(Median, digits = rounding))),
              size = 3, hjust = "left", nudge_x = 0.05, colour = "#777777") +
    geom_text(aes(x = ferment_period, y = Average, label = paste("Avg: ", round(Average, digits = rounding)), colour = ferment_period),
              size = 3, hjust = "right", nudge_x = -0.05, fontface = "bold") +
    geom_text(aes(x = ferment_period, y = Min, label = round(Min, digits = rounding), 
                  colour = ferment_period), size = 3, vjust = "top", nudge_y = -nudge, fontface = "bold") +
    geom_text(aes(x = ferment_period, y = Max, label = round(Max, digits = rounding), 
                  colour = ferment_period),size = 3, vjust = "bottom", nudge_y = nudge, fontface = "bold") +
    labs(title = paste0(var_name, ": central values and range"),
         subtitle = "Fermentation vs non-fermentation",
         x = "", y = y) +
    theme(legend.position="none")
  
    print(p)
}

```

```{r}

# Build the scales to show `status` in the same graph as the temps
# (to_scale_max - to_scale_min) * (value - value_min) / (value_max)
linear_scale <- function(x) {
  ((17-1)*(x-0)/(130-0)) + 1
}

invert_scale <- function(x) {
  ((130-0)*(x-1)/(17-1)) + 0
}

graph_series <- function(df, title = "Tank D1 series", plotly = TRUE) {

  chart <- df %>% 
    ggplot(aes(x = time)) +
        # Specificaly set `label` to also show the non-scaled status.
        geom_line(aes(y = invert_scale(status), colour = "status", label = status), size = 0.25) +
        geom_line(aes(y = setpoint, colour = "setpoint")) +
        geom_line(aes(y = temp, colour = "temp")) +
        scale_colour_manual("", breaks = c("temp", "setpoint", "status"), values = c("orange", "lightskyblue", "grey")) +
        scale_y_continuous(
          name = "Temperatures",
          sec.axis = sec_axis(trans = ~linear_scale(.), name = "Status"),
          limits = c(0, 150)
        ) +
        scale_x_datetime(date_labels="%b %y", date_breaks  = "1 month") +
        labs(title = title, x = "Time")
    
  # Show chart
  if (plotly) {
    plotly_chart <- ggplotly(chart) %>% 
      layout(title = title)
    
    print(plotly_chart)
  } else {
    print(chart)
  }
    
}

```

```{r}

graph_scatter <- function(df, x, y, title = "title", x_lab = "x", y_lab = "y") {

  x_enquo <- enquo(x)
  y_enquo <- enquo(y)

  p <- ggplot(data = df, aes(x = !!x_enquo, y = !!y_enquo, label = pulse_group)) +
    geom_point(aes(colour = ferment_period)) +
    geom_smooth(method = "loess") +
    facet_wrap(~ ferment_period, scales = "free_x") +
    labs(title = title, x = x_lab, y = y_lab) +
    theme(legend.position = "none")

  ggplotly(p, width = 800)  
}

```

```{r}

graph_density <- function(df, x_var, fill_var, title = "Density", x = "x", y = "y", fill_values = FALSE, colour_values = FALSE) {
  x_var_enquo = enquo(x_var)
  fill_var_enquo = enquo(fill_var)

  p <- ggplot(df) +
    geom_density(aes(x = !!x_var_enquo, fill = !!fill_var_enquo, ..count..), alpha = 0.5, colour = "#555555") +
    theme(legend.title = element_blank()) +
    labs(title = title,x = x, y = y)
  
  if (fill_values != FALSE && colour_values != FALSE) {
    p <- p + scale_fill_manual(breaks = fill_values, values = colour_values)
  }
  
  print(p)
}

```


```{r}

# Read main data
df <- read_csv(paste0(root_path, "/data/v3/pulse/df-pulse.csv"), col_types = cols(pulse = "c", pulse_group = "c"))

# Set status labels
df <- df %>% 
  mutate(status_label = case_when(
    status == 0 ~ "0 - error",
    status == 1 ~ "1 - off",
    status == 2 ~ "2 - on (not cooling)",
    status == 3 ~ "3 - on (cooling)",
    status == 10 ~ "10 - low temp alarm (not cooling)",
    status == 11 ~ "11 - low low temp alarm (not cooling)",
    status == 12 ~ "12 - high temp alarm (cooling)",
    status == 13 ~ "13 - high high temp alarm (cooling)",
    status == 14 ~ "14 - low temp alarm (not cooling)",
    status == 15 ~ "15 - low low temp alarm (not cooling)",
    status == 16 ~ "16 - high temp alarm (cooling)",
    status == 17 ~ "17 - high high temp alarm (cooling)",
  ))

# Set fermentation periods
df <- df %>% 
  mutate(ferment = case_when(
    time >= ymd_hms("2021-09-16 19:44:16") & time <= ymd_hms("2021-09-23 12:10:51") ~ "Sep-09",
    time >= ymd_hms("2021-10-04 18:22:17") & time <= ymd_hms("2021-10-10 00:55:03") ~ "Oct-04",
    time >= ymd_hms("2021-10-19 08:07:31") & time <= ymd_hms("2021-10-26 18:36:02") ~ "Oct-19",
    TRUE ~ "no-ferment"
  )) %>% 
  mutate(
    ferment_bool = ifelse(ferment == "no-ferment", FALSE, TRUE),
    ferment_period = ifelse(ferment == "no-ferment", "No Fermentation", "Fermentation"),
  )

```


# The event data

Here's the full dataset as a time-series showing <span style="color:darkorange">temperature</span>, <span style="color:steelblue">setpoint</span> and  <span style="color:#777">status</span>. 

We also mark the three fermentation periods:

1. Fermentation 1 from 2021-09-16 19:44:16 to 2021-09-23 12:10:51
2. Fermentation 2 from 2021-10-04 18:22:17 to 2021-10-10 00:55:03
3. Fermentation 3 from 2021-10-19 08:07:31 to 2021-10-26 18:36:02

```{r eval=FALSE}

# Build the scales to show `status` in the same graph as the temps
# (to_scale_max - to_scale_min) * (value - value_min) / (value_max)
linear_scale <- function(x) {
  ((17-1)*(x-0)/(130-0)) + 1
}

invert_scale <- function(x) {
  ((130-0)*(x-1)/(17-1)) + 0
}

chart_full_year <- ggplot(df, aes(x = time)) +
  # Set fermentation periods.
   annotate(geom = "rect", xmin = ymd_hms("2021-09-16 19:44:16"), xmax = ymd_hms("2021-09-23 12:10:51"), ymin = 0, ymax = 150,
            fill = "tomato", size = 0, alpha = 0.15) +
   annotate(geom = "rect", xmin = ymd_hms("2021-10-04 18:22:17"), xmax = ymd_hms("2021-10-10 00:55:03"), ymin = 0, ymax = 150,
            fill = "tomato", size = 0, alpha = 0.15) +
   annotate(geom = "rect", xmin = ymd_hms("2021-10-19 08:07:31"), xmax = ymd_hms("2021-10-26 18:36:02"), ymin = 0, ymax = 150,
            fill = "tomato", size = 0, alpha = 0.15) +

  # Specificaly set `label` to also show the non-scaled status.
  geom_line(aes(y = invert_scale(status), colour = "status", label = status), size = 0.25) +
  geom_line(aes(y = setpoint, colour = "setpoint")) +
  geom_line(aes(y = temp, colour = "temp")) +
  scale_colour_manual("", breaks = c("temp", "setpoint", "status"), values = c("orange", "lightskyblue", "grey")) +
  scale_y_continuous(
    name = "Temperatures",
    sec.axis = sec_axis(trans = ~linear_scale(.), name = "Status"),
    limits = c(0, 150)
  ) +
  scale_x_datetime(date_labels="%b %y", date_breaks  = "1 month") +
  labs(title = "Tank D1 temperature, status and setpoint developments 2021",
       subtitle = "Red areas indicate fermentation periods",
       x = "Time")

# Show chart
# ggplotly(chart_full_year, width = 800) %>% 
ggplotly(chart_full_year) %>%
  layout(title = list(text = paste0('Tank D1 temperature, status and setpoint developments 2021',
                                    '<br>',
                                    '<sup>',
                                    'Red areas indicate fermentation periods',
                                    '</sup>')))

```


The time series shows the development, the peaks and troughs over time. In the following we'll describe each variable, count their categories and show their averages.

As there's a conceptual difference between **non-fermentation** and **fermentation** periods, we'll split the data profile into fermentation vs no fermentation.

## Temperature and Setpoint

```{r eval=FALSE}

graph_central_values(df, temp, "Temperature", 0, "°F")
graph_central_values(df, setpoint, "Setpoint", 0, "°F")

```

## Status

```{r eval=FALSE}

# We need this to order the table
status_order <- c(
  "0 - error",
  "1 - off",
  "2 - on (not cooling)",
  "3 - on (cooling)",
  "10 - low temp alarm (not cooling)",
  "11 - low low temp alarm (not cooling)",
  "12 - high temp alarm (cooling)",
  "13 - high high temp alarm (cooling)",
  "14 - low temp alarm (not cooling)",
  "15 - low low temp alarm (not cooling)",
  "16 - high temp alarm (cooling)",
  "17 - high high temp alarm (cooling)"
)

# Data
(f <- df %>% 
  group_by(ferment_period, status_label) %>% 
  summarise(seconds = sum(diff, na.rm = TRUE), count = n()) %>% 
  mutate(hours = seconds / 60 / 60, percent = round(seconds / sum(seconds),3)) %>% 
  ungroup())

# Plot
ggplot(f) +
  geom_bar(aes(x = status_label, y = percent, fill = ferment_period), stat = "identity") +
  geom_text(aes(x = status_label, y = percent, label = scales::percent(percent, accuracy = 1), colour = ferment_period), 
            hjust = "left", nudge_y = 0.02, size = 3, fontface = "bold") +
  scale_y_continuous(limits = c(0, max(f$percent) * 1.3), labels = scales::percent) +
  coord_flip() +
  facet_wrap(~ ferment_period) +
  labs(title = "Status: frequencies",
     subtitle = "Fermentation vs non-fermentation",
     x = "Status categories", y = "") +
  theme(legend.position="none") +
  scale_x_discrete(limits = rev(status_order))


```

## Month

```{r}

df <- df %>% 
  mutate(month_fct = factor(month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")))
  
f <- df %>% 
  group_by(ferment_period, month_fct) %>% 
  summarise(count = n()) %>% 
  mutate(percent = round(count / sum(count),3))
  
ggplot(f) +
  geom_bar(aes(x = month_fct, y = percent, fill = ferment_period), stat = "identity") +
  geom_text(aes(x = month_fct, y = percent, label = scales::percent(percent, accuracy = 1), colour = ferment_period), 
            hjust = "centre", nudge_y = 0.025, size = 3, fontface = "bold") +
  scale_y_continuous(limits = c(0, max(f$percent) * 1.25)) +
  facet_wrap(~ ferment_period) +
  labs(title = "Month: frequencies",
     subtitle = "Fermentation vs non-fermentation",
     x = "", y = "") +
  theme(legend.position="none")

```

## Hour

```{r eval=FALSE}

(f <- df %>% 
  group_by(ferment_period, hour) %>% 
  summarise(count = n()) %>% 
  mutate(percent = round(count / sum(count),3)))
  
ggplot(f) +
  geom_bar(aes(x = hour, y = percent, fill = ferment_period), stat = "identity") +
  # geom_text(aes(x = hour, y = percent, label = scales::percent(percent, accuracy = 1), colour = ferment_period), 
            # hjust = "left", nudge_y = 0.02, size = 3, fontface = "bold") +
  # coord_flip() +
  facet_wrap(~ ferment_period) +
  labs(title = "Hour: frequencies",
     subtitle = "Fermentation vs non-fermentation",
     x = "", y = "") +
  theme(legend.position="none")

```

## Days

```{r}

df <- df %>% 
  mutate(wday_fct = factor(wday, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")))
  
f <- df %>% 
  group_by(ferment_period, wday_fct) %>% 
  summarise(count = n()) %>% 
  mutate(percent = round(count / sum(count),3))
  
ggplot(f) +
  geom_bar(aes(x = reorder(wday_fct, desc(wday_fct)), y = percent, fill = ferment_period), stat = "identity") +
  geom_text(aes(x = wday_fct, y = percent, label = scales::percent(percent, accuracy = 1), colour = ferment_period), 
            hjust = "left", nudge_y = 0.01, size = 3, fontface = "bold") +
  scale_y_continuous(limits = c(0, max(f$percent) * 1.25)) +
  coord_flip() +
  facet_wrap(~ ferment_period) +
  labs(title = "Weekday: frequencies",
     subtitle = "Fermentation vs non-fermentation",
     x = "", y = "") +
  theme(legend.position="none")

```



# The pulse data

In a next step we produce a dataset that just focusses on the pulse cooling periods. Each row in this dataset is a pulse cooling period and each variable summarises a specific pulse period characteristic:

```{r}

# Number of modulations.
pulse_num <- df %>% 
  group_by(pulse_group) %>% 
  count() %>% 
  mutate(pulse_group_start = parse_date_time(pulse_group, orders = "d/m/Y H:M")) %>% 
  arrange(pulse_group_start) %>% 
  ungroup()

# Main pulse period data.
df_pulse <- df %>% 
  filter(pulse == "start" | pulse == "end") %>% 
  select(time, ferment_bool, ferment, pulse, pulse_group, temp) %>% 
  mutate(ferment_period = ifelse(ferment_bool, "Fermentation", "No Fermentation")) %>% 
  mutate(temp_diff = ifelse(pulse == "start", as.numeric(lead(temp) - temp), NA)) %>% 
  mutate(pulse_duration = ifelse(pulse == "start", as.numeric(lead(time) - time), NA)) %>% 
  filter(pulse == "start") %>% 
  select(-pulse) %>%
  left_join(pulse_num %>% select(pulse_group, pulse_number = n), by = "pulse_group") %>% 
  mutate(pulse_frequency = pulse_number / pulse_duration) %>% 
  mutate(roc = temp_diff / (pulse_duration/60))  # temperature change per hour (pulse_duration is in minutes, hence /60 is in hours)

# Remove pulse groups with low single modulation
df_pulse <- df_pulse %>% 
  filter(pulse_group != "19/09/2021 12:16" & pulse_group != "23/09/2021 10:37")

# Modulation summary variables.
df_modulation <- df %>% 
  # filter(!is.na(pulse), ferment_bool) %>% 
  filter(!is.na(pulse)) %>%
  mutate(
    status_diff_2 = ifelse(status == 2, diff, NA_real_),
    status_diff_3 = ifelse(status == 3, diff, NA_real_)
  ) %>% 
  group_by(pulse_group) %>% 
  summarise(
    pulse_dur_2_mean_sec = mean(status_diff_2, na.rm = TRUE),
    pulse_dur_3_mean_sec = mean(status_diff_3, na.rm = TRUE),
    pulse_duration_2 = sum(status_diff_2, na.rm = TRUE) / 60,
    pulse_duration_3 = sum(status_diff_3, na.rm = TRUE) / 60,
  ) %>% 
  mutate(
    cooling_percent_mean = pulse_dur_3_mean_sec / (pulse_dur_2_mean_sec + pulse_dur_3_mean_sec),
    cooling_percent = pulse_duration_3 / (pulse_duration_2 + pulse_duration_3),
  ) %>% 
  ungroup()

# Add modulation variables to main data. 
(df_pulse <- df_pulse %>%
  left_join(df_modulation, by = "pulse_group") %>% 
  select(
    time,
    ferment_bool,
    ferment_period,
    ferment,
    pulse_group,
    temp,
    temp_diff,
    pulse_duration,
    roc,
    pulse_dur_2_mean_sec,
    pulse_dur_3_mean_sec,
    cooling_percent,
    pulse_duration_2,
    pulse_duration_3,
    pulse_number,
    pulse_frequency,
  ))

```

## Number of pulses

```{r eval=FALSE}

# Get ferment v no-ferment data.
(f <- df_pulse %>% 
  group_by(ferment_period) %>% 
  count() %>% 
  ungroup())

# Chart.
ggplot(f) +
  geom_bar(aes(x = ferment_period, y = n, fill = ferment_period), stat = "identity") +
  geom_text(aes(x = ferment_period, y = n, label = n, colour = ferment_period), fontface = "bold", nudge_y = 10) +
  labs(
    title = "Number of pulse periods",
    subtitle = "Fermentation vs Non-fermentation",
    x = "", y =""
  ) +
  theme(legend.position = "none")
  

# Prep correct order of fermentation periods
period_order <- c("Sep-09", "Oct-04", "Oct-19", "no-ferment")

# Get fermentation period data.
f <- df_pulse %>% 
  group_by(ferment) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(match(ferment, period_order), ferment)

# Chart.
ggplot(f) +
  geom_bar(aes(x = ferment, y = n, fill = ifelse(ferment == "no-ferment", "no", "yes")), stat = "identity") +
  geom_text(aes(x = ferment, y = n, label = n, colour = ifelse(ferment == "no-ferment", "no", "yes")), fontface = "bold", nudge_y = 10) +
  labs(
    title = "Number of pulse periods",
    subtitle = "by fermentation periods",
    x = "", y =""
  ) +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("no" = "#CCCCCC", "yes" = "#555555")) +
  scale_colour_manual(values = c("no" = "#aaaaaa", "yes" = "#555555"))

```

```{r eval=FALSE}
summarytools::descr(df_pulse)
```

## Explanatory variables

### Pulse Duration

```{r eval=FALSE}

graph_central_values(df_pulse, pulse_duration, "Pulse Duration (in minutes)")

p <- df_pulse %>% 
  ggplot() +
    geom_bar(aes(x = reorder(pulse_group, -pulse_duration), y = pulse_duration, fill = ferment_period), stat = "identity") +
    facet_wrap(~ ferment_period) +
    labs(title="Pulse duration | non-fermentation v fermentation", x = "pulse group")

ggplotly(p, width = 800)

graph_density(df_pulse, pulse_duration, ferment_period, "Density of Pulse Duration", "Pulse Duration in Minutes", "Count")

```

### Pulse frequency

```{r eval=FALSE}

df_pulse %>% filter(ferment_bool) %>% summarytools::descr(pulse_frequency)
df_pulse %>% filter(!ferment_bool) %>% summarytools::descr(pulse_frequency)

df_pulse %>% graph_central_values(pulse_frequency, "Pulse Frequency", 2)
df_pulse %>% graph_density(pulse_frequency, ferment_period, "Density of Pulse Frequency", "Number of Pulses per Minute", "Count")

df_pulse %>% filter(ferment_bool) %>% graph_density(pulse_frequency, ferment_period, "Density of Pulse Frequency", "Number of Pulses per Minute", "Count")

```

### Cooling percent

Percentage a pulse period is cooling.

```{r eval=FALSE}

# Central values.
graph_central_values(df_pulse, cooling_percent, "Percentage a pulse period is cooling", 2, "Percent (1 = 100%)")

# Ferment vs no ferment.
df_pulse %>% 
  ggplot() +
    geom_bar(aes(x = reorder(pulse_group, -cooling_percent), y = cooling_percent, fill = ferment_period), stat = "identity") +
    facet_wrap(~ ferment_period)

# Fermentation periods.
df_pulse %>% 
  ggplot() +
    geom_bar(aes(x = reorder(pulse_group, -cooling_percent), y = cooling_percent, fill = ferment_period), stat = "identity") +
    facet_wrap(~ ferment)
  
df_pulse %>% 
  filter(ferment_bool) %>% 
  ggplot() +
    geom_density(aes(x = cooling_percent, fill = ferment_period, ..count..), alpha = 0.5, colour = "#555555") +
    theme(legend.title = element_blank()) +
    labs(title = "Hello",x = "x", y = "y")

```

#### Average durations

```{r eval=FALSE}

graph_central_values(df_pulse, pulse_dur_2_mean_sec, "Avg pulse duration in seconds - non cooling (status == 2)")
graph_central_values(df_pulse, pulse_dur_3_mean_sec, "Avg pulse duration in seconds - cooling (status == 3)")

t <- df_pulse %>% 
  filter(ferment_bool) %>% 
  select(`Status 2 duration avg` = pulse_dur_2_mean_sec, `Status 3 duration avg` = pulse_dur_3_mean_sec) %>% 
  pivot_longer(cols = everything(), names_to = "duration_mean", values_to = "values")

graph_density(t, values, duration_mean, 
              "Comparing average status durations per pulse period", "Avg. Pulse Duration in Seconds", "Count",
              fill_values = unique(t$duration_mean), colour_values = c("grey", "lightskyblue"))

```

#### Total durations

```{r eval=FALSE}

graph_central_values(df_pulse, pulse_duration_2, "Total pulse duration in minutes - non cooling (status == 2)")
graph_central_values(df_pulse, pulse_duration_3, "Total pulse duration in minutes - cooling (status == 3)")

t <- df_pulse %>% 
  filter(ferment_bool) %>% 
  select(`Status 2 duration` = pulse_duration_2, `Status 3 duration` = pulse_duration_3) %>% 
  pivot_longer(cols = everything(), names_to = "duration", values_to = "values")

graph_density(t, values, duration, 
              "Comparing total status durations per pulse period", "Total Pulse Duration in Minutes", "Count",
              fill_values = unique(t$duration), colour_values = c("grey", "lightskyblue"))

```

## Response variables

### Temperature change

Starting temperature and delta

```{r eval=FALSE}

graph_central_values(df_pulse, temp, "Temperature")
graph_central_values(df_pulse, temp_diff, "Temperature delta", 2)

graph_density(df_pulse, temp_diff, ferment_period, "Temperature Change frequencies", "change in °F", "Count")

```

### Temperature rate of change

Only during fermentation

```{r eval=FALSE}

graph_central_values(df_pulse, roc, "ROC (in °F per hour)", 2, "°F per hour")

p <- df_pulse %>% 
  filter(ferment_bool) %>% 
  ggplot() +
    geom_bar(aes(x = reorder(pulse_group, -roc), y = roc), stat = "identity") +
    labs(title="Pulse duration | non-fermentation v fermentation", x = "pulse group")

ggplotly(p, width = 800)


df_pulse %>% graph_density(roc, ferment_period, "Temperature Rate of Change frequencies", "°F change per hour", "Count")
df_pulse %>% filter(ferment_bool) %>% graph_density(roc, ferment_period, "Temperature Rate of Change frequencies", "°F change per hour", "Count")

```

Check the positives

```{r eval=FALSE}

roc_pos <- df_pulse %>% 
  filter(roc > 0, ferment_bool) %>% 
  pull(pulse_group)

df %>% 
  filter(pulse_group %in% roc_pos)
  
```

```{r eval=FALSE}

for (num in seq(length(roc_pos))) {
  df %>% 
    filter(pulse_group == roc_pos[num]) %>% 
    graph_series(paste("Pulse group", roc_pos[num]))
}

```

Turns out they are based on small differences and/or short pulses:

```{r eval=FALSE}

df_pulse %>% 
  filter(roc > 0, ferment_bool) %>% 
  select(pulse_duration, temp_diff) %>% 
  summarytools::descr()

```



# Regressions

Checking the explanatory value on temperature changes:

```{r eval=FALSE}

# Cooling percent on temp diff effects
lm_fit <- lm(temp_diff ~ cooling_percent, data = df_pulse %>% filter(ferment_bool, temp_diff < 0))
summary(lm_fit)

# Pulse duration on ROC effects
lm_fit <- lm(roc ~ pulse_duration, data = df_pulse %>% filter(ferment_bool, roc < 0))
summary(lm_fit)

# Pulse frequency on ROC effects
lm_fit <- lm(roc ~ pulse_frequency, data = df_pulse %>% filter(ferment_bool, roc < 0))
summary(lm_fit)

df_pulse %>% filter(ferment_bool, roc < 0) %>% 
  graph_scatter(pulse_frequency, roc, "Pulse Frequency vs ROC", "Pulse Frequency in Pulses per Minute", "ROC in °F per hour")

```

More detailed view at the regression showing distributions, input v output scatter, fitted v output scatter as well as residuals v output scatter:

```{r}

d <- df_pulse %>% 
  filter(ferment_bool, roc < 0)

d %>% ggplot() +
  geom_density(aes(x = roc), fill = "#555555", alpha = 0.5)

d %>% ggplot() +
  geom_density(aes(x = pulse_frequency), fill = "#555555", alpha = 0.5)
  
# Pulse frequency on ROC effects
lm_fit <- lm(roc ~ pulse_frequency, data = df_pulse %>% filter(ferment_bool, roc < 0))
summary(lm_fit)

d_residuals <- d %>% 
  mutate(
    residuals = resid(lm_fit),
    fitted = fitted(lm_fit)
  )

d_residuals %>% ggplot(aes(x = pulse_frequency, y = roc)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, size = 0.3) +
    geom_hline(yintercept = 0, size = 0.3)

d_residuals %>% ggplot(aes(x = fitted, y = roc)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, linetype = "dashed", size = 0.3) +
    geom_hline(yintercept = 0, size = 0.3)

d_residuals %>% ggplot() +
    geom_point(aes(x = roc, y = residuals)) +
    geom_hline(yintercept = 0, size = 0.3)

```

# Associations

Let's see if we can infer associations visually.

## Impact of duration on ROC

```{r eval=FALSE}

df_pulse %>% 
  graph_scatter(pulse_duration, roc,
                "Impact on pulse duration on the rate of temp change", "Pulse duration in minutes", "ROC in °F per hour")

```

## Impact of cooling intensity on temperature change

```{r eval=FALSE}

df_pulse %>% 
  graph_scatter(cooling_percent, temp_diff,
                "Impact on cooling intensity on the temp change", "% of cooling during pulse period", "Temperature change in °F")

```


## Impact of cooling intensity on ROC

```{r eval=FALSE}

df_pulse %>% 
  filter(ferment_bool) %>% 
  graph_scatter(cooling_percent, roc,
                "Impact on cooling intensity on the rate of temp change", "% of cooling during pulse period", "Rate of temperature change in °F")

```

## Pulse frequency on ROC

```{r eval=FALSE}

df_pulse %>% 
  filter(ferment_bool, roc < 0) %>% 
  graph_scatter(pulse_frequency, roc,
                "Impact on cooling intensity on the rate of temp change", "% of cooling during pulse period", "Rate of temperature change in °F")


```

